### 2019-02-01 abstract/abstract.tex
Initial abstract commit.

\documentclass{bmd2019a}

\begin{document}

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Expanded Optimization for Discovering\\}
{\fontsize{16pt}{20pt}\selectfont%
  Optimal Lateral Handling Bicycles}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{Jason K. Moore, Mont Hubbard, and Ronald A. Hess}\\}
  \textit{Mechanical and Aerospace Engineering\\
          University of California, Davis\\
          One Shields Avenue, Davis, CA, USA 95817\\
          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu}
\end{flushleft}

\section*{Abstract}
%
Physical design features of ground vehicles can affect their lateral handling
qualities. The geometry, mass, mass distribution, of the vehicle's primary
components and tire characteristics are primary contributors to poor and good
handling due to their important influence on the vehicle's dynamics. In past
work, we have presented a theoretical and computational framework for assessing
the lateral task-independent handling qualities of simplified single track
vehicle designs~\cite{Hess2012,Moore2012}. It subsequent work, we have shown
that minimizing the HQM can produce theoretically optimal handling designs with
just four geometric parameters are explored as the optimization
variables~\cite{Moore2016}. The present work's goal is to expand this
optimization problem to 24 geometry, mass, and inertial parameters as the
optimization variables so that a broader search space of optimal bicycle
designs is considered. This task is complicated by the fact that we desire to
limit the search space to physically realizable bicycle designs. Doing so
introduces 35 linear and non-linear constraint equations on the optimization
variables. We formulate a constrained optimization problem and use nonlinear
programming to discover optimal, yet realizable, bicycle designs.

Our formulation exposes 24 optimization variables
$\mathbf{p}\in\mathbb{R}^{24}$ that are both reformulations of and additions to
the parameters of the benchmark bicycle~\cite{Meijaard2007}. We separate the
rider's inertial parameters from the rear frame of the bicycle for the purposes
of rider/bicycle relative configuration but do not introduce any more than the
benchmark's four rigid bodies. The variables are as follows;
%
\begin{itemize}
  \itemsep-0.25em
  \item wheelbase, trail, steer axis tilt, wheel radii
  \item somersault angle of the person (same direction as the bicycle pitch)
  \item mass of each of the front frame, rear frame, front wheel, and rear wheel
  \item mass center of the front frame, rear frame, and rider
  \item principal central radii of gyration of both the front and rear frames
  \item principal inertia pitch angle of the front frame, rear frame, and rider
\end{itemize}

We formulate a bounds on the variables as well as linear and nonlinear
constraints among the variables to ensure a physically realizable bicycle.
Below the constraints are defined and grouped by the associated rigid body or
collection thereof:
%
\begin{description}

  \item[Vehicle] The resulting combination of the five rigid bodies.
    \begin{itemize}
      \itemsep0em
      \item The physical extents of the bicycle and rider must exist above the
        ground plane.
      \item The bicycle and rider are symmetric about the rider's sagittal
        plane.
      \item The mass of each bicycle rigid body is positive and greater than a
        minimum value and the total below a reasonably lift-able amount.
      \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
        radius.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration, i.e. -g < acceleration < g.
    \end{itemize}

  \item[Wheels] Both the front and rear wheels have the same constraints.
    \begin{itemize}
      \itemsep0em
      \item The radius must be greater than a minimum value.
      \item The bicycle wheels should be inertially wheel-like, i.e.
        $I_{rot}=mr^2, I_{rad}=I_{rot}/2$.
    \end{itemize}

  \item[Frames] Both the front and rear frames have the same constraints.
    \begin{itemize}
      \itemsep0em
      \item We limit a pair of principal directions to lie in the roll/yaw
        plane and that one of the axes must be within $\pm \pi/4$ with respect
        to the ground plane.
      \item We require the mass and inertia of the frames to be positive and of
        large enough values to be constructed from materials in a space frame
        of a specified minimal density.
      \item We enforce that the body is of a planar nature, i.e.
        $k_{lateral}^2=k_{longitudinal}^2+k_{vertical}^2$, where $k$ is the
        radius of gyration.
    \end{itemize}

  \item[Rider] A single rigid body representing the rider.
    \begin{itemize}
      \itemsep0em
      \item The mass of the rider is that of an average person.
      \item The rider's joint angles are in a fixed nominal configuration that
        is typical for upright bicycling and the resulting mass distribution of
        the rider is derived from standard body segment estimation methods.
      \item One principal axis must be parallel with the bicycle's pitch axis.
    \end{itemize}

\end{description}

The constraint statements above translate into upper and lower bounds on 12 of
the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
constraints among the variables $\mathbf{g}(\mathbf{p})$ and thus can be
expressed in the standard non-linear programming form.
%
\begin{equation}
  \begin{aligned}
    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
    & \text{subject to} & & \\
    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
  \end{aligned}
\end{equation}

We make use of the interior-point optimizer IPOPT~\cite{Wachter2006} to find
solutions to this nonlinear programming problem at variety of bicycle speeds
and will present the resulting discovered optimal bicycle designs in the full
conference paper.

\bibliographystyle{acm}
\bibliography{BMD2019}

\end{document}

### 2019-02-01 abstract/abstract.tex
Added Mont's updates.

@@
 Physical design features of ground vehicles can affect their lateral handling
-qualities. The geometry, mass, mass distribution, of the vehicle's primary
-components and tire characteristics are primary contributors to poor and good
+qualities. Geometry, mass, and mass distribution of the vehicle's primary
+components as well as tire characteristics are primary contributors to poor and good
 handling due to their important influence on the vehicle's dynamics. In past
@@
 the lateral task-independent handling qualities of simplified single track
-vehicle designs~\cite{Hess2012,Moore2012}. It subsequent work, we have shown
-that minimizing the HQM can produce theoretically optimal handling designs with
-just four geometric parameters are explored as the optimization
+vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
+that minimizing the HQM can produce theoretically optimal handling designs when
+only four geometric parameters are explored as the optimization
 variables~\cite{Moore2016}. The present work's goal is to expand this
@@
 designs is considered. This task is complicated by the fact that we desire to
-limit the search space to physically realizable bicycle designs. Doing so
+limit the search space to physically realizable designs. Doing so
 introduces 35 linear and non-linear constraint equations on the optimization
@@
   \item wheelbase, trail, steer axis tilt, wheel radii
-  \item somersault angle of the person (same direction as the bicycle pitch)
-  \item mass of each of the front frame, rear frame, front wheel, and rear wheel
-  \item mass center of the front frame, rear frame, and rider
-  \item principal central radii of gyration of both the front and rear frames
-  \item principal inertia pitch angle of the front frame, rear frame, and rider
+  \item somersault angle of the person (same direction as bicycle pitch)
+  \item masses of the front frame, rear frame, front wheel, and rear wheel
+  \item mass center locations of the front frame, rear frame, and rider
+  \item principal central radii of gyration of front and rear frames
+  \item principal inertia axis pitch angles of the front frame, rear frame, and rider
 \end{itemize}
 
-We formulate a bounds on the variables as well as linear and nonlinear
+We formulate bounds on the variables as well as linear and nonlinear
 constraints among the variables to ensure a physically realizable bicycle.
@@
         ground plane.
-      \item The bicycle and rider are symmetric about the rider's sagittal
+      \item Both bicycle and rider are symmetric about the sagittal
         plane.
@@
       \item The bicycle cannot topple forward during hard braking or backward
-        during hard acceleration, i.e. -g < acceleration < g.
+        during hard acceleration, i.e. when -g < acceleration < g.
     \end{itemize}
 
-  \item[Wheels] Both the front and rear wheels have the same constraints.
+  \item[Wheels] Both front and rear wheels have identical constraints.
     \begin{itemize}
       \itemsep0em
-      \item The radius must be greater than a minimum value.
-      \item The bicycle wheels should be inertially wheel-like, i.e.
+      \item Wheel radius must be greater than a minimum value.
+      \item Wheels are inertially wheel-like, i.e.
         $I_{rot}=mr^2, I_{rad}=I_{rot}/2$.
@@
 
-  \item[Frames] Both the front and rear frames have the same constraints.
+  \item[Frames] Both front and rear frames have identical constraints.
     \begin{itemize}
       \itemsep0em
-      \item We limit a pair of principal directions to lie in the roll/yaw
-        plane and that one of the axes must be within $\pm \pi/4$ with respect
-        to the ground plane.
-      \item We require the mass and inertia of the frames to be positive and of
-        large enough values to be constructed from materials in a space frame
-        of a specified minimal density.
-      \item We enforce that the body is of a planar nature, i.e.
+      \item A pair of principal directions lie in the roll/yaw
+        plane and one of these axes is within $\pm \pi/4$ of the ground plane.   Why?????
+      \item The mass and inertia of the frames are positive and
+        large enough to be constructed from materials in a space frame
+        of specified minimal density.
+      \item The bicycle???? is essentially of a planar nature, i.e.
         $k_{lateral}^2=k_{longitudinal}^2+k_{vertical}^2$, where $k$ is the
@@
 
-  \item[Rider] A single rigid body representing the rider.
+  \item[Rider] A single rigid body represents the rider.
     \begin{itemize}
       \itemsep0em
-      \item The mass of the rider is that of an average person.
-      \item The rider's joint angles are in a fixed nominal configuration that
-        is typical for upright bicycling and the resulting mass distribution of
-        the rider is derived from standard body segment estimation methods.
-      \item One principal axis must be parallel with the bicycle's pitch axis.
+      \item Rider mass is that of an average person.
+      \item The rider's joint angles are fixed in a nominal configuration typical of upright bicycling and the resulting mass distribution is derived from standard body segment estimation methods.
+      \item One rider principal axis is parallel with the bicycle pitch axis. Why??????
     \end{itemize}

### 2019-02-01 abstract/abstract.tex
Final edits.

@@
 qualities. Geometry, mass, and mass distribution of the vehicle's primary
-components as well as tire characteristics are primary contributors to poor and good
-handling due to their important influence on the vehicle's dynamics. In past
-work, we have presented a theoretical and computational framework for assessing
-the lateral task-independent handling qualities of simplified single track
-vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
-that minimizing the HQM can produce theoretically optimal handling designs when
-only four geometric parameters are explored as the optimization
-variables~\cite{Moore2016}. The present work's goal is to expand this
-optimization problem to 24 geometry, mass, and inertial parameters as the
-optimization variables so that a broader search space of optimal bicycle
+components as well as tire characteristics are primary contributors to poor and
+good handling due to their important influence on the vehicle's dynamics. In
+past work, we have presented a theoretical and computational framework for
+assessing the lateral task-independent handling qualities of simplified single
+track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
+that minimizing our handling quality metric (HQM) can produce theoretically
+optimal handling designs when only four geometric parameters are explored as
+the optimization variables~\cite{Moore2016}. The present work's goal is to
+expand this optimization problem to 24 geometry, mass, and inertial parameters
+as the optimization variables so that a broader search space of optimal bicycle
 designs is considered. This task is complicated by the fact that we desire to
-limit the search space to physically realizable designs. Doing so
-introduces 35 linear and non-linear constraint equations on the optimization
-variables. We formulate a constrained optimization problem and use nonlinear
-programming to discover optimal, yet realizable, bicycle designs.
+limit the search space to physically realizable designs. Doing so introduces 35
+linear and non-linear constraint equations on the optimization variables. We
+formulate a constrained optimization problem and use nonlinear programming to
+discover optimal, yet realizable, bicycle designs.
 
@@
         plane.
-      \item The mass of each bicycle rigid body is positive and greater than a
-        minimum value and the total below a reasonably lift-able amount.
+      \item The mass of each bicycle rigid body is positive, greater than a
+        minimum value, and the total mass is below a reasonably lift-able
+        amount.
       \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
@@
       \item The bicycle cannot topple forward during hard braking or backward
-        during hard acceleration, i.e. when -g < acceleration < g.
+        during hard acceleration, i.e. -1 g < acceleration < 1 g.
     \end{itemize}
@@
       \item Wheel radius must be greater than a minimum value.
-      \item Wheels are inertially wheel-like, i.e.
-        $I_{rot}=mr^2, I_{rad}=I_{rot}/2$.
+      \item Wheels are inertially wheel-like, i.e.  $I_{rot}=mr^2,
+        I_{rad}=I_{rot}/2$.
     \end{itemize}
@@
       \item A pair of principal directions lie in the roll/yaw
-        plane and one of these axes is within $\pm \pi/4$ of the ground plane.   Why?????
-      \item The mass and inertia of the frames are positive and
-        large enough to be constructed from materials in a space frame
-        of specified minimal density.
-      \item The bicycle???? is essentially of a planar nature, i.e.
+        plane to maintain symmetry and one of these axes is within $\pm \pi/4$
+        rad of the ground plane.
+      \item The mass and inertia of the frames are positive and large enough to
+        be constructed from materials in a space frame of specified minimal
+        density.
+      \item The frames are planar in nature, i.e.
         $k_{lateral}^2=k_{longitudinal}^2+k_{vertical}^2$, where $k$ is the
@@
       \item Rider mass is that of an average person.
-      \item The rider's joint angles are fixed in a nominal configuration typical of upright bicycling and the resulting mass distribution is derived from standard body segment estimation methods.
-      \item One rider principal axis is parallel with the bicycle pitch axis. Why??????
+      \item The rider's joint angles are fixed in a nominal configuration
+        typical of upright bicycling and the resulting mass distribution is
+        derived from standard body segment estimation methods.
+      \item One rider principal axis is parallel with the bicycle pitch axis to
+        enforce symmetry.
     \end{itemize}
@@
 the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
-constraints among the variables $\mathbf{g}(\mathbf{p})$ and thus can be
+constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
 expressed in the standard non-linear programming form.
@@
     & \text{subject to} & & \\
+    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
     & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
-    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
   \end{aligned}
@@
 We make use of the interior-point optimizer IPOPT~\cite{Wachter2006} to find
-solutions to this nonlinear programming problem at variety of bicycle speeds
+solutions to this nonlinear programming problem at a variety of vehicle speeds
 and will present the resulting discovered optimal bicycle designs in the full

### 2019-07-28 paper/paper.tex
Added first work on paper based on the work taken from the abandoned ICSC 2017 work.

\documentclass{icsc}

\usepackage{siunitx}

\begin{document}

\begin{center}
  \fontsize{14}{20}{\bf Optimal bicycle design to maximize handling and safety}
\end{center}

\begin{center}
  \normalsize{\bf{Jason K. Moore, Mont Hubbard,  Ronald A. Hess}}
\end{center}

\begin{center}
  \begin{tabular}{c}
    Department of Mechanical and Aerospace Engineering\\
    University of California, Davis\\
    One Shields Avenue, Davis, CA, USA 95616\\
    e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu\\
  \end{tabular}
\end{center}

\begin{keywords}
  bicycle,
  design,
  handling qualities,
  control,
  optimization
\end{keywords}

\appendix

\section{Free Parameters and Constraints}

\subsection{Rear Frame [D]}

There are six free parameters associated with the rear frame's inertial characteristics:

\begin{description}
  \item[$m_D$] Mass of the rear frame.
  \item[$k_{Dxx}$] Central radius of gyration with respect to the $X$ axis where $k_{Dxx}^2=I_{Dxx} / m_D$.
  \item[$k_{Dyy}$] Central radius of gyration with respect to the $Y$ axis where $k_{Dyy}^2=I_{Dyy} / m_D$.
  \item[$k_{Dzz}$] Central radius of gyration with respect to the $Z$ axis where $k_{Dzz}^2=I_{Dzz} / m_D$.
  \item[$\alpha_D$] Angle between the $X$ axis and the primary principal inertia axis.
  \item[$x_D$] Location along the $X$ axis of the rear frame center of mass.
  \item[$z_D$] Location along the $Z$ axis of the rear frame center of mass.
\end{description}

\begin{itemize}
  \item The body is symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} = I_{Dyz} = 0$.
  \item We require that the center of mass not penetrate the ground, i.e. $z_D < 0$.
  \item We require that the mass be positive, i.e. $m_D > 0$.
  \item We allow for any angular orientation of the $XZ$ principal directions but limit the angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
  \item We enforce that the body is planar with $k_{Dyy} = \sqrt{k_{Dxx}^2 + k_{Dyy}^2}$.
\end{itemize}

\subsection{Person [P]}

We assume that the person's joint configurations are such that they are in a nominal configuration for pedaling, i.e. an average normal everyday ???or maximum power?? riding position. The person is assumed to be symmetric about the $XZ$ plane. There are four fixed parameters that enforce this assumption:

\begin{description}
  \item[$m_P$] The mass of the person which is fixed at $XX \si{kg}$.
  \item[$k_{Pxx}$] Central radius of gyration with respect to the $X$ axis where $I_{Pxx} = m_P k_{Pxx}^2$.
  \item[$k_{Pyy}$] Central radius of gyration with respect to the $Y$ axis where $I_{Pyy} = m_P k_{Pyy}^2=$.
  \item[$k_{Pzz}$] Central radius of gyration with respect to the $Z$ axis where $I_{Pzz} = m_P k_{Pzz}^2=$.
  \item[$I_{Bxy}, I_{Byz}$] Products of inertia are zero.
  \item[$y_P$] Is zero.
\end{description}

We allow the rider to be rotated about the $Y$ axis and positioned anywhere within the plane of symmetry. Three free parameters are associated with these assumptions:

\begin{description}
  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal inertia axis of the person.
  \item[$x_P$] Location along the $X$ axis of the person's center of mass.
  \item[$z_P$] Location along the $Z$ axis of the person's center of mass.
\end{description}

To prevent the rider from being positioned and oriented such that their body is not penetrating the the ground we introduce two dimensions that define a cross whose apex is at the center of mass of the person and the cross axes are parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance along the principal axis to the tip of the toes and $w_P$ is the distance along the second principal axes to the tip of the hands.

\begin{align}
  z_P + l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P + w_P / 2 \sin{\alpha_P} & \leq 0 \\
  z_P - l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P - w_P / 2 \sin{\alpha_P} & \leq 0
\end{align}

\subsection{Rear Wheel [R]}

We enforce the assumption that the wheel should be wheel-like, i.e. $I_{Ryy} = m_R r_R^2$ and $I_{Rxx} = I_{Rzz} = I_{Ryy} / 2$ thus there are two free parameters:

\begin{description}
  \item[$r_R$] Outer radius of the rear wheel.
  \item[$m_R$] Mass of the rear wheel.
\end{description}

We also enforce that the mass and radius should be positive with:

\begin{align}
  m_R \geq 0 \\
  r_R \geq 0
\end{align}

\subsection{Front Wheel}

Same as rear wheel.

\subsection{Front Frame}

The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. There are seven free parameters:

\begin{description}
  \item[$m_H$] The mass of the rear frame.
  \item[$k_{Hxx}$] Central radius of gyration with respect to the $X$ axis where $k_{Hxx}^2=I_{Hxx} / m_H$.
  \item[$k_{Hyy}$] Central radius of gyration with respect to the $X$ axis where $k_{Hyy}^2=I_{Hyy} / m_H$.
  \item[$k_{Hzz}$] Central radius of gyration with respect to the $X$ axis where $k_{Hzz}^2=I_{Hzz} / m_H$.
  \item[$\alpha_H$] Angle between the $X$ axis and the primary principal inertia axis.
  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
\end{description}

\subsection{Total Bike}

There are three remaing free parameters associated with the fundamental geometry of the bicycle:

\begin{description}
  \item[$w$] Wheelbase
  \item[$c$] Trail
  \item[$\lambda$] Steer axis tilt
\end{description}

To prevent the wheels from overlapping we introduce:

\begin{align}
  w > r_F + r_R
\end{align}

The trail can take on any real value as well as the steer axis tilt except the bound the angle to a minimal range:

\begin{align}
  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
\end{align}

We require that the bicycle be liftable by an average person with:

\begin{align}
  m_T = m_H + m_B + m_F + m_R \leq 25 \si{kg}
\end{align}

Finally, we require that the bicycle not topple forward during hard breaking and backward during hard acceleration where:

\begin{align}
  -g < a < \frac{g}{4}
\end{align}

This translates to two constraints that bound the total center of mass in a triangle in teh $XZ$ plane:

\begin{align}
  \frac{x_T}{|z_T|} < \frac{1}{4} \\
  \frac{x_T - w}{|z_T|} < -1
\end{align}

\subsection{Summary}

The above definitions represent a constrained optimization in 23 free parameters with 17 constraints. This bounds the space of bicycle designs to be physically realizable and excludes bicycles that don't have the inherent characteristics of a bicycle.

\begin{align}
  p = & [m_D, k_{Dxx}, k_{Dzz}, \alpha_D, x_D, z_D, x_P, z_P, \alpha_P, r_R, m_R, r_F, m_H, \\
      & x_H, z_H, k_{Hxx}, k_{Hyy}, k_{Hzz}, \alpha_H, w, c, \lambda]
\end{align}

with these fixed values:

\begin{align}
  I_{Dxy} = I_{Dyz} = 0 \\
  % 
  l_P = ? \si{m}\\
  w_P = ? \si{m} \\
  m_P = 83.5 \si{kg} \\
  y_P = 0 \si{\meter} \\
  I_{Pxx} = 10.98524 \si{\kg\meter^2} \\
  I_{Pyy} = 11.2571 \si{\kg\meter^2}\\
  I_{Pzz} = 2.26922 \si{\kg\meter^2}\\
  I_{Pxy} = I_{Pyz} = 0 \si{\kg\meter^2}\\
  I_{Pxz} = -1.70211 \si{\kg\meter^2}\\
  %
  I_{Bxy} = I_{Byz} = 0 \\
  % 
  I_{Hxy} = I_{Hyz} = 0 \\
  %
  I_{Rxy} = I_{Ryz} = I_{Rxz} = 0 \\
  %
  I_{Fxy} = I_{Fyz} = I_{Fxz} = 0 \\
\end{align}

with these parameter bounds:

\begin{align}
  m_R > 0 \\
  m_F > 0 \\
  m_H > 0 \\
  m_D > 0 \\
  r_R > 0 \\
  r_F > 0 \\
  z_H < 0 \\
  z_D < 0 \\
  -\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2} \\
  -\frac{\pi}{2} \leq \alpha_P \leq \frac{\pi}{2} \\
  -\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2} \\
  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
\end{align}

with these parameter convertions:

\begin{align}
  I_{Dxx} = m_D k_{Dxx}^2 \\
  I_{Dzz} = m_D k_{Dzz}^2 \\
  I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2} \\
  I_{Dxz} = f(I_{Dxx}, I_{Dzz}, \alpha_D)  \\
  %
  m_B = m_D + m_P \\
  x_B = \frac{m_D x_D + m_P x_P}{m_B} \\
  z_B = \frac{m_D z_D + m_P z_P}{m_B} \\
  I_{Bxx} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  I_{Byy} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  I_{Bzz} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  I_{Bxy} = 0 \\
  I_{Byz} = 0 \\
  I_{Bxz} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  % 
  I_{Hxx} = m_H k_{Hxx}^2 \\
  I_{Hzz} = m_H k_{Hzz}^2 \\
  I_{Hyy} = \sqrt{I_{Hxx}^2 + I_{Hzz}^2} \\
  I_{Hxz} = f(I_{Hxx}, I_{Hzz}, \alpha_H)  \\
  %
  I_{Ryy} = m_R r_R^2 \\
  I_{Rxx} = I_{Rzz} = I_{Ryy} / 2 \\
  %
  I_{Fyy} = m_F r_F^2 \\
  I_{Fxx} = I_{Fzz} = I_{Fyy} / 2 \\
  %
  m_T = m_H + m_B + m_F + m_R \\
  x_T = \frac{m_H x_H + m_B x_B + m_F w}{m_T} \\
  z_T = \frac{m_H z_H + m_B z_B + m_F r_F + m_R r_R}{m_T} 
\end{align}

and these constraints:

\begin{align}
  I_{Hyy} \leq \sqrt{I_{Hxx}^2 + I_{Hzz}^2} \\
  z_P + l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P + w_P / 2 \sin{\alpha_P} & \leq 0 \\
  z_P - l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P - w_P / 2 \sin{\alpha_P} & \leq 0 \\
  w > r_F + r_R \\
  m_T \leq 25 \si{kg} \\
  \frac{x_T}{|z_t|} < \frac{1}{4} \\
  \frac{x_T - w}{|z_T|} < -1
\end{align}

\end{document}

### 2019-07-28 paper/paper.tex
Some tidying on the details of the parameter/constraints.

@@
-\documentclass{icsc}
+\documentclass{article}
 
 \usepackage{siunitx}
+\usepackage{amsmath}
 
@@
 \begin{center}
-  \fontsize{14}{20}{\bf Optimal bicycle design to maximize handling and safety}
+  \fontsize{14}{20}{\bf Expanded Optimization for Discovering Optimal Lateral
+  Handling Bicycles}
 \end{center}
@@
 
-\begin{keywords}
-  bicycle,
-  design,
-  handling qualities,
-  control,
-  optimization
-\end{keywords}
-
 \appendix
@@
 
-There are six free parameters associated with the rear frame's inertial characteristics:
+There are 5 free parameters associated with the rear frame's inertial
+characteristics:
 
@@
   \item[$m_D$] Mass of the rear frame.
-  \item[$k_{Dxx}$] Central radius of gyration with respect to the $X$ axis where $k_{Dxx}^2=I_{Dxx} / m_D$.
-  \item[$k_{Dyy}$] Central radius of gyration with respect to the $Y$ axis where $k_{Dyy}^2=I_{Dyy} / m_D$.
-  \item[$k_{Dzz}$] Central radius of gyration with respect to the $Z$ axis where $k_{Dzz}^2=I_{Dzz} / m_D$.
-  \item[$\alpha_D$] Angle between the $X$ axis and the primary principal inertia axis.
   \item[$x_D$] Location along the $X$ axis of the rear frame center of mass.
   \item[$z_D$] Location along the $Z$ axis of the rear frame center of mass.
+  \item[$k_{Dxx}$] Central radius of gyration with respect to the $X$ axis
+    where $k_{Dxx}^2=I_{Dxx} / m_D$.
+  \item[$k_{Dzz}$] Central radius of gyration with respect to the $Z$ axis
+    where $k_{Dzz}^2=I_{Dzz} / m_D$.
+  \item[$\alpha_D$] Angle between the $X$ axis and the primary principal
+    inertia axis.
 \end{description}
 
+Assumptions:
+
 \begin{itemize}
-  \item The body is symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} = I_{Dyz} = 0$.
-  \item We require that the center of mass not penetrate the ground, i.e. $z_D < 0$.
+  \item The body is symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} =
+    I_{Dyz} = 0$.
+  \item We require that the center of mass not penetrate the ground, i.e. $z_D
+    < 0$.
   \item We require that the mass be positive, i.e. $m_D > 0$.
-  \item We allow for any angular orientation of the $XZ$ principal directions but limit the angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
-  \item We enforce that the body is planar with $k_{Dyy} = \sqrt{k_{Dxx}^2 + k_{Dyy}^2}$.
+  \item We allow for any angular orientation of the $XZ$ principal directions
+    but limit the angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
+  \item We enforce that the body is planar with $k_{Dyy} = \sqrt{k_{Dxx}^2 +
+    k_{Dyy}^2}$.
 \end{itemize}
@@
 
-We assume that the person's joint configurations are such that they are in a nominal configuration for pedaling, i.e.
-an average normal everyday ???or maximum power??
-riding position.
-The person is assumed to be symmetric about the $XZ$ plane.
-There are four fixed parameters that enforce this assumption:
+We assume that the person's joint configurations are such that they are in a
+nominal configuration for pedaling, i.e. an average normal everyday riding
+position. The person is assumed to be symmetric about the $XZ$ plane. There are
+four fixed parameters that enforce this assumption:
 
@@
   \item[$m_P$] The mass of the person which is fixed at $XX \si{kg}$.
-  \item[$k_{Pxx}$] Central radius of gyration with respect to the $X$ axis where $I_{Pxx} = m_P k_{Pxx}^2$.
-  \item[$k_{Pyy}$] Central radius of gyration with respect to the $Y$ axis where $I_{Pyy} = m_P k_{Pyy}^2=$.
-  \item[$k_{Pzz}$] Central radius of gyration with respect to the $Z$ axis where $I_{Pzz} = m_P k_{Pzz}^2=$.
+  \item[$k_{Pxx}$] Central radius of gyration with respect to the $X$ axis
+    where $I_{Pxx} = m_P k_{Pxx}^2$.
+  \item[$k_{Pyy}$] Central radius of gyration with respect to the $Y$ axis
+    where $I_{Pyy} = m_P k_{Pyy}^2=$.
+  \item[$k_{Pzz}$] Central radius of gyration with respect to the $Z$ axis
+    where $I_{Pzz} = m_P k_{Pzz}^2=$.
   \item[$I_{Bxy}, I_{Byz}$] Products of inertia are zero.
@@
 
-We allow the rider to be rotated about the $Y$ axis and positioned anywhere within the plane of symmetry. Three free parameters are associated with these assumptions:
+We allow the rider to be rotated about the $Y$ axis and positioned anywhere
+within the plane of symmetry. Three free parameters are associated with these
+assumptions:
 
 \begin{description}
-  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal inertia axis of the person.
+  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal
+    inertia axis of the person.
   \item[$x_P$] Location along the $X$ axis of the person's center of mass.
@@
 
-To prevent the rider from being positioned and oriented such that their body is not penetrating the the ground we introduce two dimensions that define a cross whose apex is at the center of mass of the person and the cross axes are parallel to the principal axes in the $XZ$ plane.
-$l_P / 2$ is the distance along the principal axis to the tip of the toes and $w_P$ is the distance along the second principal axes to the tip of the hands.
+To prevent the rider from being positioned and oriented such that their body is
+not penetrating the the ground we introduce two dimensions that define a cross
+whose apex is at the center of mass of the person and the cross axes are
+parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
+along the principal axis to the tip of the toes and $w_P$ is the distance along
+the second principal axes to the tip of the hands.
 
@@
 
+\subsection{Front Frame}
+
+The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$.
+There are seven free parameters:
+
+\begin{description}
+  \item[$m_H$] The mass of the rear frame.
+  \item[$k_{Hxx}$] Central radius of gyration with respect to the $X$ axis
+    where $k_{Hxx}^2=I_{Hxx} / m_H$.
+  \item[$k_{Hyy}$] Central radius of gyration with respect to the $X$ axis
+    where $k_{Hyy}^2=I_{Hyy} / m_H$.
+  \item[$k_{Hzz}$] Central radius of gyration with respect to the $X$ axis
+    where $k_{Hzz}^2=I_{Hzz} / m_H$.
+  \item[$\alpha_H$] Angle between the $X$ axis and the primary principal
+    inertia axis.
+  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
+  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
+\end{description}
+
+Assumptions:
+
+\begin{itemize}
+  \item We allow for any angular orientation of the $XZ$ principal directions
+    but limit the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.
+\end{itemize}
+
 \subsection{Rear Wheel [R]}
 
-We enforce the assumption that the wheel should be wheel-like, i.e. $I_{Ryy} = m_R r_R^2$ and $I_{Rxx} = I_{Rzz} = I_{Ryy} / 2$ thus there are two free parameters:
+We enforce the assumption that the wheel should be wheel-like, i.e. $I_{Ryy} =
+m_R r_R^2$ and $I_{Rxx} = I_{Rzz} = I_{Ryy} / 2$ thus there are two free
+parameters:
 
@@
 
-\subsection{Front Frame}
-
-The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. There are seven free parameters:
-
-\begin{description}
-  \item[$m_H$] The mass of the rear frame.
-  \item[$k_{Hxx}$] Central radius of gyration with respect to the $X$ axis where $k_{Hxx}^2=I_{Hxx} / m_H$.
-  \item[$k_{Hyy}$] Central radius of gyration with respect to the $X$ axis where $k_{Hyy}^2=I_{Hyy} / m_H$.
-  \item[$k_{Hzz}$] Central radius of gyration with respect to the $X$ axis where $k_{Hzz}^2=I_{Hzz} / m_H$.
-  \item[$\alpha_H$] Angle between the $X$ axis and the primary principal inertia axis.
-  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
-  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
-\end{description}
-
 \subsection{Total Bike}
 
-There are three remaing free parameters associated with the fundamental geometry of the bicycle:
+There are three remaing free parameters associated with the fundamental
+geometry of the bicycle:
 
@@
 
-The trail can take on any real value as well as the steer axis tilt except the bound the angle to a minimal range:
+The trail can take on any real value as well as the steer axis tilt except the
+bound the angle to a minimal range:
 
@@
 \begin{align}
-  m_T = m_H + m_B + m_F + m_R \leq 25 \si{kg}
+  m_T = m_H + m_D + m_P + m_F + m_R \leq m_P + 25 \si{kg}
 \end{align}
 
-Finally, we require that the bicycle not topple forward during hard breaking and backward during hard acceleration where:
+Finally, we require that the bicycle not topple forward during hard breaking
+and backward during hard acceleration where:
 
@@
 
-\begin{align}
-  \frac{x_T}{|z_T|} < \frac{1}{4} \\
-  \frac{x_T - w}{|z_T|} < -1
+\begin{align} \frac{x_T}{|z_T|} < \frac{1}{4} \\ \frac{x_T - w}{|z_T|} < -1
 \end{align}
@@
 
-The above definitions represent a constrained optimization in 23 free parameters with 17 constraints.
-This bounds the space of bicycle designs to be physically realizable and excludes bicycles that don't have the inherent characteristics of a bicycle.
+The above definitions represent a constrained optimization in 23 free
+parameters with 17 constraints. This bounds the space of bicycle designs to be
+physically realizable and excludes bicycles that don't have the inherent
+characteristics of a bicycle.
 
 \begin{align}
-  p = & [m_D, k_{Dxx}, k_{Dzz}, \alpha_D, x_D, z_D, x_P, z_P, \alpha_P, r_R, m_R, r_F, m_H, \\
-      & x_H, z_H, k_{Hxx}, k_{Hyy}, k_{Hzz}, \alpha_H, w, c, \lambda]
+  \mathbf{p} =
+  \begin{bmatrix}
+    % front frame
+    m_D \\
+    x_D \\
+    z_D \\
+    k_{Dxx} \\
+    k_{Dzz} \\
+    \alpha_D \\
+    % person
+    x_P \\
+    z_P \\
+    \alpha_P \\
+    % front frame
+    m_H \\
+    x_H \\
+    z_H \\
+    k_{Hxx} \\
+    k_{Hyy} \\
+    k_{Hzz} \\
+    I_{Hxz} \\
+    \alpha_H \\
+    % rear wheel
+    r_R \\
+    m_R \\
+    % front wheel
+    r_F \\
+    m_F \\
+    % primary geometry
+    w \\
+    c \\
+    \lambda
+  \end{bmatrix}
 \end{align}
 
-with these fixed values:
+with these additional parameters that have fixed values:
 
@@
   I_{Dxy} = I_{Dyz} = 0 \\
+  y_D = 0 \\
   % 
@@
   % 
-  I_{Bxy} = I_{Byz} = 0 \\
-  % 
   I_{Hxy} = I_{Hyz} = 0 \\
+  y_H = 0 \\
   %
@@
 
-with these parameter bounds:
-
-\begin{align}
-  m_R > 0 \\
-  m_F > 0 \\
-  m_H > 0 \\
-  m_D > 0 \\
-  r_R > 0 \\
-  r_F > 0 \\
-  z_H < 0 \\
-  z_D < 0 \\
-  -\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2} \\
-  -\frac{\pi}{2} \leq \alpha_P \leq \frac{\pi}{2} \\
-  -\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2} \\
-  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
-\end{align}
-
-with these parameter convertions:
+with these conversions from our parameters to the benchmark parameters:
 
@@
   I_{Dxx} = m_D k_{Dxx}^2 \\
+  I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2} \\
   I_{Dzz} = m_D k_{Dzz}^2 \\
-  I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2} \\
   I_{Dxz} = f(I_{Dxx}, I_{Dzz}, \alpha_D)  \\
@@
   I_{Hxx} = m_H k_{Hxx}^2 \\
+  I_{Hyy} = m_H k_{Hyy}^2\\
   I_{Hzz} = m_H k_{Hzz}^2 \\
-  I_{Hyy} = \sqrt{I_{Hxx}^2 + I_{Hzz}^2} \\
   I_{Hxz} = f(I_{Hxx}, I_{Hzz}, \alpha_H)  \\
@@
   I_{Ryy} = m_R r_R^2 \\
-  I_{Rxx} = I_{Rzz} = I_{Ryy} / 2 \\
+  I_{Rxx} = I_{Ryy} / 2 \\
+  I_{Rzz} = I_{Ryy} / 2 \\
   %
   I_{Fyy} = m_F r_F^2 \\
-  I_{Fxx} = I_{Fzz} = I_{Fyy} / 2 \\
+  I_{Fxx} = I_{Fyy} / 2 \\
+  I_{Fzz} = I_{Fyy} / 2 \\
   %
@@
   z_T = \frac{m_H z_H + m_B z_B + m_F r_F + m_R r_R}{m_T} 
+\end{align}
+
+with these parameter bounds:
+
+\begin{align}
+  m_R > 0 \\
+  m_F > 0 \\
+  m_H > 0 \\
+  m_D > 0 \\
+  r_R > r_{min} \\
+  r_F > r_{min} \\
+  z_H < 0 \\
+  z_D < 0 \\
+  -\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2} \\
+  -\frac{\pi}{2} \leq \alpha_P \leq \frac{\pi}{2} \\
+  -\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2} \\
+  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
 \end{align}
@@
   w > r_F + r_R \\
-  m_T \leq 25 \si{kg} \\
-  \frac{x_T}{|z_t|} < \frac{1}{4} \\
-  \frac{x_T - w}{|z_T|} < -1
+  m_{Tmin} + m_P \leq m_T \leq 25 \si{kg} + m_P \\
+  \frac{x_T}{|z_T|} < \frac{1}{4} \\
+  \frac{x_T - w}{|z_T|} < -1 \\
+  real(\mathbf{s}) < 0
 \end{align}

### 2019-07-28 paper/src/convert_principal_to_benchmark.m
Framed out a function to convert our principal parameters to the benchmark parameters.

@@ -1,0 +1,106 @@
+function benchmark_par = convert_principal_to_benchmark(principal_par)
+
+    p = principal_par;
+
+    % other
+    p.g
+    p.v
+    % primary geometry
+    p.w
+    p.c
+    p.lam
+    % rear frame
+    p.mD
+    p.xD
+    p.yD
+    p.zD
+    p.kD11
+    p.kD22
+    p.kD33
+    p.alphaD
+    % person
+    p.xP
+    p.yP
+    p.zP
+    p.alphaP
+    % front frame
+    p.mH
+    p.xH
+    p.yH
+    p.zH
+    p.kH11
+    p.kH22
+    p.kH33
+    p.alphaH
+    % rear wheel
+    p.rR
+    p.mR
+    % front wheel
+    p.rF
+    p.mF
+
+
+    % other
+    b.g = p.g
+    b.v = p.v
+    % primary geometry
+    b.w = p.w;
+    b.c = p.c;
+    b.lam = p.lam;
+
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % rear frame [B]
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    b.mB = p.mD + p.mP;
+    Bcom = com_from_two(p.mD, [p.xD, p.yD, p.zD], p.mP, [p.xP, p.yD, p.zP]);
+    b.xB = Bcom(1);
+    b.yB = Bcom(2);
+    b.zB = Bcom(3);
+
+    % symmetry is assumed about the XZ plane
+    ID123 = mD*diag([kD11^2, p.kD22^2, p.kD33^2]);
+    IP123 = mP*diag([kP11^2, p.kP22^2, p.kP33^2]);
+    IP = rotate_inerita_about_y(IP123, alphaP);
+    ID = rotate_inerita_about_y(ID123, alphaD);
+    IB = sum_central_inertias(ID123, [xD, yD, zD], IP, [xP, yP, zP]);
[…]

### 2019-07-28 paper/src/convert_principal_to_benchmark.m
Started on docstring of principal_to_benchmark.

@@ -1,52 +1,63 @@
 function benchmark_par = convert_principal_to_benchmark(principal_par)
 
+
+    % primary geometry, gravity, and speed
+    % w : wheelbase [m]
+    % c : trail [m]
+    % lam : steer axis tilt [m]
+    % g : acceleration due to gravity [m/s^2]
+    % v : forward speed [m/s]
+    % rear frame
+    % mD : mass of the rear frame (without rider) [kg]
+    % xD : X CoM location [m]
+    % yD : Y CoM location [m]
+    % zD : Z CoM location [m]
+    % kD11 : Central principal radius of gyration
+    % kD22 : Central principal radius of gyration
+    % kD33 : Central principal radius of gyration
+    % alphaD : Angle between the X and 1 axes, positive about Y
+    % person
+    % mP
+    % xP
+    % yP
+    % zP
+    % kP11
+    % kP22
+    % kP33
+    % alphaP
+    % front frame
+    % mH
+    % xH
+    % yH
+    % zH
+    % kH11
+    % kH22
+    % kH33
+    % alphaH
+    % rear wheel
+    % rR
+    % mR
+    % kR11 : Maximal principal radius of gyration, aligned with Y in the
+    % case of the wheels
+    % kR22 : Intermediate principal radius of gyration
+    % kR33 : Minimal principal radius of gyration
+    % front wheel
+    % rF
+    % mF
+    % kF11
+    % kF22
+    % kF33
+
     p = principal_par;
 
-    % other
-    p.g
-    p.v
-    % primary geometry
-    p.w
-    p.c
-    p.lam
-    % rear frame
-    p.mD
-    p.xD
-    p.yD
-    p.zD
-    p.kD11
-    p.kD22
-    p.kD33
-    p.alphaD
-    % person
-    p.xP
-    p.yP
-    p.zP
-    p.alphaP
-    % front frame
-    p.mH
-    p.xH
-    p.yH
-    p.zH
-    p.kH11
-    p.kH22
-    p.kH33
-    p.alphaH
-    % rear wheel
-    p.rR
-    p.mR
[…]

### 2019-07-28 src/convert_principal_to_benchmark.m
More updates to docstring.

@@ -1,12 +1,26 @@
 function benchmark_par = convert_principal_to_benchmark(principal_par)
-
-
+    % EVAL_RHS - Returns a structure containing the benchmark bicycle
+    % parameters as defined in Meijaard2007 which are converted from the
+    % principal parameters as defined in MooreHubbard2019.
+    %
+    % Syntax: benchmark_par = convert_principal_to_benchmark(principal_par)
+    %
+    % Inputs:
+    %   principal_par - Structure containing parameter names mapped to
+    %   doubles.
+    % Outputs:
+    %   benchmark_par - Structure containing parameters names mapped to
+    %   doubls.
+    %
+    % Explanation of the principal parameters:
+    %
     % primary geometry, gravity, and speed
     % w : wheelbase [m]
     % c : trail [m]
     % lam : steer axis tilt [m]
     % g : acceleration due to gravity [m/s^2]
     % v : forward speed [m/s]
+    %
     % rear frame
     % mD : mass of the rear frame (without rider) [kg]
     % xD : X CoM location [m]
@@ -16,6 +30,7 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
     % kD22 : Central principal radius of gyration
     % kD33 : Central principal radius of gyration
     % alphaD : Angle between the X and 1 axes, positive about Y
+    %
     % person
     % mP
     % xP
@@ -25,6 +40,7 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
     % kP22
     % kP33
     % alphaP
+    %
     % front frame
     % mH
     % xH
@@ -34,6 +50,7 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
     % kH22
     % kH33
     % alphaH
+    %
     % rear wheel
     % rR
     % mR
@@ -41,6 +58,7 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
     % case of the wheels
     % kR22 : Intermediate principal radius of gyration
     % kR33 : Minimal principal radius of gyration
+    %
     % front wheel
     % rF
     % mF

### 2019-07-28 src/com_from_two.m
Added functions for com and inertial calcs.

@@ -1,0 +1,3 @@
+function com = com_from_two(m1, com1, m2, com2)
+m = m1 + m2;
+com = m1.*com1 + m2.*com2) ./ m

### 2019-07-28 src/convert_principal_to_benchmark.m

@@ -1,7 +1,8 @@
 function benchmark_par = convert_principal_to_benchmark(principal_par)
-    % EVAL_RHS - Returns a structure containing the benchmark bicycle
-    % parameters as defined in Meijaard2007 which are converted from the
-    % principal parameters as defined in MooreHubbard2019.
+    % CONVERT_PRINCIPAL_TO_BENCHMARK - Returns a structure containing the
+    % benchmark bicycle parameters as defined in Meijaard2007 which are
+    % converted from the principal parameters as defined in
+    % MooreHubbard2019.
     %
     % Syntax: benchmark_par = convert_principal_to_benchmark(principal_par)
     %
@@ -91,7 +92,7 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
     IP123 = mP*diag([kP11^2, p.kP22^2, p.kP33^2]);
     IP = rotate_inerita_about_y(IP123, alphaP);
     ID = rotate_inerita_about_y(ID123, alphaD);
-    IB = sum_central_inertias(ID123, [xD, yD, zD], IP, [xP, yP, zP]);
+    IB = sum_central_inertias(p.mD, [xD; yD; zD], ID123, p.mP, [xP; yP; zP], IP);
     b.IBxx = IB(1, 1);
     b.IByy = IB(2, 2);
     b.IBzz = IB(3, 3);

### 2019-07-28 src/parallel_axis.m

@@ -1,0 +1,20 @@
+function I = parallel_axis(Ic, m, d)
+% PARALLEL_AXIS - Returns and inertia tensor about a new point.
+% Syntax : I = parallel_axis(Ic, m, d)
+% Inputs:
+%   Ic - 3x3 double, The moment of inertia about the center of mass of the
+%        body with respect to an orthogonal coordinate system.
+%   m - 1x1 double, The mass of the body.
+%   d - 3x1 double, The distances along the three ordinates that located the new point
+%       relative to the center of mass of the body.
+%
+% Outputs:
+%   I : 3x3 double, The moment of inertia of a body about a point located by
+%   the distances in d.
+a = d(1, 1);
+b = d(2, 1);
+c = d(3, 1);
+dMat = [b**2 + c**2,      -a * b,      -a * c;
+             -a * b, c**2 + a**2,      -b * c;
+             -a * c,      -b * c, a**2 + b**2]
+I = Ic + m * dMat

### 2019-07-28 src/rotate_inertia_about_y.m

@@ -1,0 +1,18 @@
+function rotated_inertia = rotate_inertia_about_y(inertia, angle):
+
+    %Returns inertia tensor rotated through angle about the Y axis.
+%
+    %Parameters
+    %----------
+    %I : ndarray, shape(3,)
+        %An inertia tensor.
+    %angle : float
+        %Angle in radians about the positive Y axis of which to rotate the
+        %inertia tensor.
+
+    ca = cos(angle)
+    sa = sin(angle)
+    C = [ca, 0.0, -sa;
+         0.0, 1., 0.0;
+         sa, 0.0, ca];
+    rotated_inertia = C * I * C.T

### 2019-07-28 src/sum_central_inertias.m

@@ -1,0 +1,16 @@
+function I = sum_central_inertias(m1, com1, I1, m2, com2, I2)
+% SUM_CENTRAL_INERTIAS - Returns the combined inertia of two bodies.
+%
+% Syntax : I = sum_central_inertias(m1, com1, I1, m2, com2, I2)
+%
+% Inputs:
+%   m1 - 1x1 double, mass of the first body
+%   com1 - 3x1 double, center of mass of the first body
+%   I1 - 3x3 double, central inertia tensor of the first body
+%   m2 - 1x1 double, mass of the second body
+%   com2 - 3x1 double, center of mass of the second body
+%   I2 - 3x3 double, central inertia tensor of the second body
+% Outputs:
+%   I - combined central inertia tensor of both bodies
+com = com_from_two(m1, com1, m2, com2);
+I = parallel_axis(I1, m1, com-com1) +  parallel_axis(I2, m2, com-com2);

### 2019-07-28 src/bounds.m
Added file to encode the bounds on the free parameters.

@@ -1,0 +1,52 @@
+function lb, ub = bounds()
+% BOUNDS - Returns the lower and upper bounds of the free parameters.
+
+lb = [
+ nan; % w
+ nan; % c
+ -pi/2; % lam
+ 0.0; % mD
+ nan; % xD
+ nan; % zD
+ nan; % kDxx
+ nan; % kDzz
+ -pi/2; % alphaD
+ nan; % xP
+ nan; % zP
+ -pi/2; % alphaP
+ 0.0; % mH
+ nan; % xH
+ nan; % zH
+ nan; % kHxx
+ nan; % kHyy
+ nan; % kHzz
+ -pi/2; % alphaH
+ 0.127; % rR (10" OD wheel)
+ 1.0; % mR
+ 0.127; % rF (10" OD wheel)
+ 1.0]; % mF
+
+ub = [
+ nan; % w
+ nan; % c
+ pi/2; % lam
+ nan; % mD
+ nan; % xD
+ 0.0; % zD
+ nan; % kDxx
+ nan; % kDzz
+ pi/2; % alphaD
+ nan; % xP
+ 0.0; % zP
+ pi/2; % alphaP
+ nan; % mH
+ nan; % xH
+ 0.0; % zH
+ nan; % kHxx
+ nan; % kHyy
+ nan; % kHzz
+ pi/2; % alphaH
+ nan; % rR
+ nan; % mR
+ nan; % rF
+ nan]; % mF

### 2019-07-29 src/constraints.m
Added files with constraints.

@@ -1,0 +1,22 @@
+% c(x) <= 0
+% nonlinear
+Ihyy - sqrt(IHxx^2 + IHzz^2);
+zP + lP / 2 * cos(alphaP);
+zP + wP / 2 * sin(alphaP);
+zP - lP / 2 * cos(alphaP);
+zP - wP / 2 * sin(alphaP);
+4*xT/abs(zT) - 1
+abs(zT) + xT - w
+real(s)  % real part of closed loop eigenvalues must be negative
+% linear
+rF + rF - w;  % wheels can't over lap
+b.mD + mH + mR + mF - 25.0;  % bicycle mass (D,H,R,F) no more than 25 kg
+
+% ceq(x) = 0
+IDyy - sqrt(IDxx^2 + IDzz^2)  % frame is planar
+IRyy - mR*rR^2  % wheel is a ring
+IRxx - IRyy / 2  % wheel is a ring
+IRzz - IRyy / 2  % wheel is a ring
+IFyy - mF*rF^2  % wheel is a ring
+IFxx - IFyy / 2  % wheel is a ring
+IFzz - IFyy / 2  % wheel is a ring

### 2019-07-29 src/free_parameters.m

@@ -1,0 +1,28 @@
+function x = free_parameters(principal_parameters)
+% FREE_PARAMETERS - Returns the free parameter vector, x, given the full set
+% of principal parameters.
+p = principal_parameters;
+x = [
+p.w;
+p.c;
+p.lam;
+p.mD;
+p.xD;
+p.zD;
+p.kDxx;
+p.kDzz;
+p.alphaD;
+p.xP;
+p.zP;
+p.alphaP;
+p.mH;
+p.xH;
+p.zH;
+p.kHxx;
+p.kHyy;
+p.kHzz;
+p.alphaH;
+p.rR;
+p.mR;
+p.rF;
+p.mF];

### 2019-07-29 src/update_principal_parameters.m

@@ -1,0 +1,29 @@
+function p = update_principal_parameters(principal_parameters, x)
+% Updates the structure with the values stored in the free parameter vector
+% x.
+
+p = principal_parameters;
+
+p.w      = x(1, 1);
+p.c      = x(2, 1);
+p.lam    = x(3, 1);
+p.mD     = x(4, 1);
+p.xD     = x(5, 1);
+p.zD     = x(6, 1);
+p.kDxx   = x(7, 1);
+p.kDzz   = x(8, 1);
+p.alphaD = x(9, 1);
+p.xP     = x(10, 1);
+p.zP     = x(11, 1);
+p.alphaP = x(12, 1);
+p.mH     = x(13, 1);
+p.xH     = x(14, 1);
+p.zH     = x(15, 1);
+p.kHxx   = x(16, 1);
+p.kHyy   = x(17, 1);
+p.kHzz   = x(18, 1);
+p.alphaH = x(19, 1);
+p.rR     = x(20, 1);
+p.mR     = x(21, 1);
+p.rF     = x(22, 1);
+p.mF     = x(23, 1);

### 2019-07-29 src/bounds.m
paper/paper.tex
Reorder free param vector.

@@ -1,25 +1,25 @@
-function lb, ub = bounds()
+function [lb, ub] = bounds()
 % BOUNDS - Returns the lower and upper bounds of the free parameters.
 
 lb = [
- nan; % w
- nan; % c
+ -inf; % w
+ -inf; % c
  -pi/2; % lam
  0.0; % mD
- nan; % xD
- nan; % zD
- nan; % kDxx
- nan; % kDzz
+ -inf; % xD
+ -inf; % zD
+ -inf; % kDxx
+ -inf; % kDzz
  -pi/2; % alphaD
- nan; % xP
- nan; % zP
+ -inf; % xP
+ -inf; % zP
  -pi/2; % alphaP
  0.0; % mH
- nan; % xH
- nan; % zH
- nan; % kHxx
- nan; % kHyy
- nan; % kHzz
+ -inf; % xH
+ -inf; % zH
+ -inf; % kHxx
+ -inf; % kHyy
+ -inf; % kHzz
  -pi/2; % alphaH
  0.127; % rR (10" OD wheel)
  1.0; % mR
@@ -27,26 +27,26 @@ lb = [
  1.0]; % mF
 
 ub = [
- nan; % w
- nan; % c
+ inf; % w
+ inf; % c
  pi/2; % lam
- nan; % mD
- nan; % xD
+ inf; % mD
+ inf; % xD
  0.0; % zD
- nan; % kDxx
- nan; % kDzz
+ inf; % kDxx
+ inf; % kDzz
  pi/2; % alphaD
- nan; % xP
+ inf; % xP
  0.0; % zP
  pi/2; % alphaP
- nan; % mH
- nan; % xH
+ inf; % mH
+ inf; % xH
  0.0; % zH
- nan; % kHxx
- nan; % kHyy
- nan; % kHzz
+ inf; % kHxx
+ inf; % kHyy
+ inf; % kHzz
  pi/2; % alphaH
- nan; % rR
- nan; % mR
- nan; % rF
- nan]; % mF
+ inf; % rR
+ inf; % mR
+ inf; % rF
+ inf]; % mF

### 2019-07-29 src/com_from_two.m

@@ -1,3 +1,3 @@
 function com = com_from_two(m1, com1, m2, com2)
 m = m1 + m2;
-com = m1.*com1 + m2.*com2) ./ m
+com = (m1.*com1 + m2.*com2) ./ m;

### 2019-07-29 src/rotate_inertia_about_y.m

@@ -1,4 +1,4 @@
-function rotated_inertia = rotate_inertia_about_y(inertia, angle):
+function rotated_inertia = rotate_inertia_about_y(inertia, angle)
 
     %Returns inertia tensor rotated through angle about the Y axis.
 %
@@ -10,9 +10,9 @@ function rotated_inertia = rotate_inertia_about_y(inertia, angle):
         %Angle in radians about the positive Y axis of which to rotate the
         %inertia tensor.
 
-    ca = cos(angle)
-    sa = sin(angle)
+    ca = cos(angle);
+    sa = sin(angle);
     C = [ca, 0.0, -sa;
          0.0, 1., 0.0;
          sa, 0.0, ca];
-    rotated_inertia = C * I * C.T
+    rotated_inertia = C * inertia * C';

### 2019-07-29 src/bounds.m
Simplified par bounds function.

@@ -1,52 +1,35 @@
-function [lb, ub] = bounds()
-% BOUNDS - Returns the lower and upper bounds of the free parameters.
+function [lower, upper] = free_par_bounds()
+% FREE_PAR_BOUNDS - Returns the lower and upper bounds of the free
+% parameters.
+%
+% Outputs:
+%   lower - 23x1 double
+%   upper - 23x1 double
 
-lb = [
- -inf; % w
- -inf; % c
- -pi/2; % lam
- 0.0; % mD
- -inf; % xD
- -inf; % zD
- -inf; % kDxx
- -inf; % kDzz
- -pi/2; % alphaD
- -inf; % xP
- -inf; % zP
- -pi/2; % alphaP
- 0.0; % mH
- -inf; % xH
- -inf; % zH
- -inf; % kHxx
- -inf; % kHyy
- -inf; % kHzz
- -pi/2; % alphaH
- 0.127; % rR (10" OD wheel)
- 1.0; % mR
- 0.127; % rF (10" OD wheel)
- 1.0]; % mF
+bounds = [
+  -inf, inf;  % w
+  -inf, inf;  % c
+ -pi/2, pi/2; % lam
+   0.0, inf;  % mD
+  -inf, inf;  % xD
+  -inf, 0.0;  % zD
+  -inf, inf;  % kDxx
+  -inf, inf;  % kDzz
+ -pi/2, pi/2; % alphaD
+  -inf, inf;  % xP
+  -inf, 0.0;  % zP
+ -pi/2, pi/2; % alphaP
+   0.0, inf;  % mH
+  -inf, inf;  % xH
+  -inf, 0.0;  % zH
+  -inf, inf;  % kHxx
+  -inf, inf;  % kHyy
+  -inf, inf;  % kHzz
+ -pi/2, pi/2; % alphaH
+ 0.127, inf;  % rR (10" OD wheel)
+   1.0, inf;  % mR
+ 0.127, inf;  % rF (10" OD wheel)
+   1.0, inf]; % mF
 
-ub = [
- inf; % w
- inf; % c
- pi/2; % lam
- inf; % mD
- inf; % xD
- 0.0; % zD
- inf; % kDxx
- inf; % kDzz
- pi/2; % alphaD
- inf; % xP
- 0.0; % zP
- pi/2; % alphaP
- inf; % mH
- inf; % xH
- 0.0; % zH
- inf; % kHxx
- inf; % kHyy
[…]

### 2019-07-29 src/free_par_bounds.m
Renamed free par bounds function.

@@ -1,0 +1,35 @@
+function [lower, upper] = free_par_bounds()
+% FREE_PAR_BOUNDS - Returns the lower and upper bounds of the free
+% parameters.
+%
+% Outputs:
+%   lower - 23x1 double
+%   upper - 23x1 double
+
+bounds = [
+  -inf, inf;  % w
+  -inf, inf;  % c
+ -pi/2, pi/2; % lam
+   0.0, inf;  % mD
+  -inf, inf;  % xD
+  -inf, 0.0;  % zD
+  -inf, inf;  % kDxx
+  -inf, inf;  % kDzz
+ -pi/2, pi/2; % alphaD
+  -inf, inf;  % xP
+  -inf, 0.0;  % zP
+ -pi/2, pi/2; % alphaP
+   0.0, inf;  % mH
+  -inf, inf;  % xH
+  -inf, 0.0;  % zH
+  -inf, inf;  % kHxx
+  -inf, inf;  % kHyy
+  -inf, inf;  % kHzz
+ -pi/2, pi/2; % alphaH
+ 0.127, inf;  % rR (10" OD wheel)
+   1.0, inf;  % mR
+ 0.127, inf;  % rF (10" OD wheel)
+   1.0, inf]; % mF
+
+lower = bounds(:, 1);
+upper = bounds(:, 1);

### 2019-07-29 src/com_from_two.m
Docstring for com combine.

@@ -1,3 +1,15 @@
 function com = com_from_two(m1, com1, m2, com2)
+% COM_FROM_TWO - Returns a vector giving the center of mass of two combined
+% bodies.
+%
+% Inputs:
+%   m1 - 1x1 double, mass of body 1
+%   com1 - 3x1 double, Cartesian coordinates of body 1's center of mass
+%   m2 - 1x1 double, mass of body 2
+%   com2 - 3x1 double, Cartesian coordinates of body 2's center of mass
+%
+% Outputs:
+%   com - 3x1 double, Cartesian coordinates of the combined bodies center of
+%         mass
 m = m1 + m2;
-com = (m1.*com1 + m2.*com2) ./ m;
+com = (m1.*com1 + m2.*com2)./m;

### 2019-07-29 src/bounds.m
Deleted bounds.m

@@ -1,35 +1,0 @@
-function [lower, upper] = free_par_bounds()
-% FREE_PAR_BOUNDS - Returns the lower and upper bounds of the free
-% parameters.
-%
-% Outputs:
-%   lower - 23x1 double
-%   upper - 23x1 double
-
-bounds = [
-  -inf, inf;  % w
-  -inf, inf;  % c
- -pi/2, pi/2; % lam
-   0.0, inf;  % mD
-  -inf, inf;  % xD
-  -inf, 0.0;  % zD
-  -inf, inf;  % kDxx
-  -inf, inf;  % kDzz
- -pi/2, pi/2; % alphaD
-  -inf, inf;  % xP
-  -inf, 0.0;  % zP
- -pi/2, pi/2; % alphaP
-   0.0, inf;  % mH
-  -inf, inf;  % xH
-  -inf, 0.0;  % zH
-  -inf, inf;  % kHxx
-  -inf, inf;  % kHyy
-  -inf, inf;  % kHzz
- -pi/2, pi/2; % alphaH
- 0.127, inf;  % rR (10" OD wheel)
-   1.0, inf;  % mR
- 0.127, inf;  % rF (10" OD wheel)
-   1.0, inf]; % mF
-
-lower = bounds(:, 1);
-upper = bounds(:, 1);

### 2019-07-29 src/constraints.m
Fleshed out constraint function more.

@@ -1,22 +1,60 @@
+function [c, ceq] = compute_constraints(x, p)
+% COMPUTE_CONSTRAINTS - Returns the nonlinear inequality and equality
+% constraints.
+%
+% Inputs:
+%   x - 23x1 double, free parameters
+%   p - structure, default principal parameters
+%
+% Outputs:
+%   c - 15x1 double, inequality constraint values
+%   ceq - 7x1 double, equality constraint values
+
+p = update_principal_parameters(p, x);
+b = convert_principal_to_benchmark(p);
+
+% Inequality constraints
 % c(x) <= 0
 % nonlinear
-Ihyy - sqrt(IHxx^2 + IHzz^2);
-zP + lP / 2 * cos(alphaP);
-zP + wP / 2 * sin(alphaP);
-zP - lP / 2 * cos(alphaP);
-zP - wP / 2 * sin(alphaP);
-4*xT/abs(zT) - 1
-abs(zT) + xT - w
-real(s)  % real part of closed loop eigenvalues must be negative
+% if this was equality the handelbar would be XZ planar, but if we want
+% some IHzz and IHxx (protruding handle bars in Y direction) then this:
+% TODO : need better explanation of this.
+c(1, 1) = b.IHyy - sqrt(b.IHxx^2 + b.IHzz^2);
+% the person does not penetrate the ground
+c(1, 2) = p.zP + p.lP / 2 * cos(p.alphaP);
+c(1, 3) = p.zP + p.wP / 2 * sin(p.alphaP);
+c(1, 4) = p.zP - p.lP / 2 * cos(p.alphaP);
+c(1, 5) = p.zP - p.wP / 2 * sin(p.alphaP);
+% the bike doesn't tip forward and backward when braking and accelerating
+comT = com_from_two(b.mB + b.mH, ...
+    com_from_two(b.mB, [b.xB; b.yB; b.zB], b.mH, [b.xH; b.yH; b.zH]), ...
+    b.mR + b.mF, ...
+    com_from_two(b.mR, [b.xR; b.yR; b.zR], b.mF, [b.xF; b.yF; b.zF]))
+xT = comT(1, 1);
+zT = comT(3, 1);
+c(1, 6) = 4*xT/abs(zT) - 1
+c(1, 7) = (xT - p.w) / abs(zT) + 1
[…]

### 2019-07-29 src/convert_principal_to_benchmark.m
Formatting of prin to bench convert func.

@@ -1,140 +1,137 @@
 function benchmark_par = convert_principal_to_benchmark(principal_par)
-    % CONVERT_PRINCIPAL_TO_BENCHMARK - Returns a structure containing the
-    % benchmark bicycle parameters as defined in Meijaard2007 which are
-    % converted from the principal parameters as defined in
-    % MooreHubbard2019.
-    %
-    % Syntax: benchmark_par = convert_principal_to_benchmark(principal_par)
-    %
-    % Inputs:
-    %   principal_par - Structure containing parameter names mapped to
-    %   doubles.
-    % Outputs:
-    %   benchmark_par - Structure containing parameters names mapped to
-    %   doubls.
-    %
-    % Explanation of the principal parameters:
-    %
-    % primary geometry, gravity, and speed
-    % w : wheelbase [m]
-    % c : trail [m]
-    % lam : steer axis tilt [m]
-    % g : acceleration due to gravity [m/s^2]
-    % v : forward speed [m/s]
-    %
-    % rear frame
-    % mD : mass of the rear frame (without rider) [kg]
-    % xD : X CoM location [m]
-    % yD : Y CoM location [m]
-    % zD : Z CoM location [m]
-    % kD11 : Central principal radius of gyration
-    % kD22 : Central principal radius of gyration
-    % kD33 : Central principal radius of gyration
-    % alphaD : Angle between the X and 1 axes, positive about Y
-    %
-    % person
-    % mP
-    % xP
-    % yP
-    % zP
-    % kP11
-    % kP22
-    % kP33
-    % alphaP
-    %
-    % front frame
-    % mH
-    % xH
-    % yH
-    % zH
-    % kH11
-    % kH22
-    % kH33
-    % alphaH
-    %
[…]

### 2019-07-29 src/free_parameters.m
Formatting, indentation.

@@ -1,28 +1,29 @@
 function x = free_parameters(principal_parameters)
 % FREE_PARAMETERS - Returns the free parameter vector, x, given the full set
 % of principal parameters.
+
 p = principal_parameters;
-x = [
-p.w;
-p.c;
-p.lam;
-p.mD;
-p.xD;
-p.zD;
-p.kDxx;
-p.kDzz;
-p.alphaD;
-p.xP;
-p.zP;
-p.alphaP;
-p.mH;
-p.xH;
-p.zH;
-p.kHxx;
-p.kHyy;
-p.kHzz;
-p.alphaH;
-p.rR;
-p.mR;
-p.rF;
-p.mF];
+
+x = [p.w;
+     p.c;
+     p.lam;
+     p.mD;
+     p.xD;
+     p.zD;
+     p.kDxx;
+     p.kDzz;
+     p.alphaD;
+     p.xP;
+     p.zP;
+     p.alphaP;
+     p.mH;
+     p.xH;
+     p.zH;
+     p.kHxx;
+     p.kHyy;
+     p.kHzz;
+     p.alphaH;
+     p.rR;
+     p.mR;
+     p.rF;
+     p.mF];

### 2019-07-29 src/parallel_axis.m
Improved documentation.

@@ -16,5 +16,5 @@ b = d(2, 1);
 c = d(3, 1);
 dMat = [b**2 + c**2,      -a * b,      -a * c;
              -a * b, c**2 + a**2,      -b * c;
-             -a * c,      -b * c, a**2 + b**2]
-I = Ic + m * dMat
+             -a * c,      -b * c, a**2 + b**2];
+I = Ic + m * dMat;

### 2019-07-29 src/rotate_inertia_about_y.m

@@ -1,18 +1,14 @@
 function rotated_inertia = rotate_inertia_about_y(inertia, angle)
-
-    %Returns inertia tensor rotated through angle about the Y axis.
+% ROTATE_INERTIA_ABOUT_Y - Returns inertia tensor rotated through angle
+% about the Y axis.
 %
-    %Parameters
-    %----------
-    %I : ndarray, shape(3,)
-        %An inertia tensor.
-    %angle : float
-        %Angle in radians about the positive Y axis of which to rotate the
-        %inertia tensor.
-
-    ca = cos(angle);
-    sa = sin(angle);
-    C = [ca, 0.0, -sa;
-         0.0, 1., 0.0;
-         sa, 0.0, ca];
-    rotated_inertia = C * inertia * C';
+% Inputs:
+%   inertia - 3x3 double, inertia tensor
+%   angle - 1x1 double, Angle in radians about the positive Y axis of which
+%           to rotate the inertia tensor.
+ca = cos(angle);
+sa = sin(angle);
+C = [ca, 0.0, -sa;
+     0.0, 1., 0.0;
+     sa, 0.0, ca];
+rotated_inertia = C * inertia * C';

### 2019-07-29 src/default_principal_parameters.m
Added a file with some default parameters based on the pista.

@@ -1,0 +1,42 @@
+function p = default_principal_parameters()
+
+% based on Jason sitting on the Pista
+
+p.w = 0.989
+p.c = 0.0617300113263
+p.g = 9.81
+p.v = 5.0
+p.lam = 0.275762021815
+p.mH = 2.27
+p.mR = 1.38
+p.mF = 1.58
+p.rR = 0.332088156971
+p.rF = 0.333838861345
+p.xH = 0.906321313298
+p.zH = -0.732376356772
+% TODO : pick some reasonable values for these two.
+p.lP = nan;
+p.wP = nan;
+
+% TODO : Put into principal radius of gyration form.
+IFxx = 0.0553476418952
+IFyy = 0.106185287785
+IRxx = 0.0552268228143
+IRyy = 0.0764034890242
+
+% TODO : Put into principal radius of gyration form.
+IHxx = 0.0979884309288
+IHxz = -0.00440828467861
+IHyy = 0.0691907586693
+IHzz = 0.0396107377431
+
+% TODO: Break into separate bicycle and rider and put in principal radius of
+% gyration form
+mB = 4.49
+xB = 0.382963782674
+yB = 0.0
+zB = -0.476798507146
+IBxx = 0.289968529291
+IBxz = 0.0502317583637
+IByy = 0.475817094274
+IBzz = 0.249418014295

### 2019-07-29 src/convert_principal_to_benchmark.m
test

@@ -57,7 +57,7 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
 % kR22 : Intermediate principal radius of gyration
 % kR33 : Minimal principal radius of gyration
 %
-% front wheel
+% front  wheel
 % rF
 % mF
 % kF11

### 2019-07-29 src/constraints.m
Added a todo about expressing in principal parameters.

@@ -19,6 +19,7 @@ b = convert_principal_to_benchmark(p);
 % if this was equality the handelbar would be XZ planar, but if we want
 % some IHzz and IHxx (protruding handle bars in Y direction) then this:
 % TODO : need better explanation of this.
+% TODO : Should we express in terms of kH11, kH22, kH33?
 c(1, 1) = b.IHyy - sqrt(b.IHxx^2 + b.IHzz^2);
 % the person does not penetrate the ground
 c(1, 2) = p.zP + p.lP / 2 * cos(p.alphaP);

### 2019-07-30 src/combine_mass_centers.m
Added a function to compute mass centers of an arbitrary number of bodies.

@@ -1,0 +1,13 @@
+function mass_center = combine_mass_centers(masses, mass_centers)
+% COMBINE_MASS_CENTERS - Returns a vector giving the mass center the
+% combined bodies.
+%
+% Inputs:
+%   masses - 1xn double, masses of the n bodies
+%   mass_centers - 3xn double, Cartesian coordinates of the n bodies' mass
+%   centers
+%
+% Outputs:
+%   com - 3x1 double, Cartesian coordinates of the combined bodies mass
+%   center
+mass_center = sum(masses .* mass_centers, 2) ./ sum(masses);

### 2019-07-30 src/com_from_two.m
Switched to more versatile com combination function.

@@ -1,15 +1,0 @@
-function com = com_from_two(m1, com1, m2, com2)
-% COM_FROM_TWO - Returns a vector giving the center of mass of two combined
-% bodies.
-%
-% Inputs:
-%   m1 - 1x1 double, mass of body 1
-%   com1 - 3x1 double, Cartesian coordinates of body 1's center of mass
-%   m2 - 1x1 double, mass of body 2
-%   com2 - 3x1 double, Cartesian coordinates of body 2's center of mass
-%
-% Outputs:
-%   com - 3x1 double, Cartesian coordinates of the combined bodies center of
-%         mass
-m = m1 + m2;
-com = (m1.*com1 + m2.*com2)./m;

### 2019-07-30 src/compute_constraints.m

@@ -27,10 +27,11 @@ c(1, 3) = p.zP + p.wP / 2 * sin(p.alphaP);
 c(1, 4) = p.zP - p.lP / 2 * cos(p.alphaP);
 c(1, 5) = p.zP - p.wP / 2 * sin(p.alphaP);
 % the bike doesn't tip forward and backward when braking and accelerating
-comT = com_from_two(b.mB + b.mH, ...
-    com_from_two(b.mB, [b.xB; b.yB; b.zB], b.mH, [b.xH; b.yH; b.zH]), ...
-    b.mR + b.mF, ...
-    com_from_two(b.mR, [b.xR; b.yR; b.zR], b.mF, [b.xF; b.yF; b.zF]))
+comT = combine_mass_centers([b.mB, b.mH, b.mR, b.mF],
+                            [[b.xB; b.yB; b.zB],
+                             [b.xH; b.yH; b.zH],
+                             [b.xR; b.yR; b.zR],
+                             [b.xF; b.yF; b.zF]]);
 xT = comT(1, 1);
 zT = comT(3, 1);
 c(1, 6) = 4*xT/abs(zT) - 1

### 2019-07-30 src/convert_principal_to_benchmark.m

@@ -79,7 +79,8 @@ b.v = p.v
 % rear frame [B]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 b.mB = p.mD + p.mP;
-Bcom = com_from_two(p.mD, [p.xD, p.yD, p.zD], p.mP, [p.xP, p.yD, p.zP]);
+Bcom = combine_mass_centers([p.mD, p.mP],
+                            [[p.xD; p.yD; p.zD], [p.xP; p.yD; p.zP]]);
 b.xB = Bcom(1);
 b.yB = Bcom(2);
 b.zB = Bcom(3);

### 2019-07-30 src/sum_central_inertias.m

@@ -12,5 +12,5 @@ function I = sum_central_inertias(m1, com1, I1, m2, com2, I2)
 %   I2 - 3x3 double, central inertia tensor of the second body
 % Outputs:
 %   I - combined central inertia tensor of both bodies
-com = com_from_two(m1, com1, m2, com2);
+com = combine_mass_centers([m1, m2], [com1, com2]);
 I = parallel_axis(I1, m1, com-com1) +  parallel_axis(I2, m2, com-com2);

### 2019-07-30 src/compute_constraints.m
Added correct calls to generate_data and made frame planar constraint based on principal params.

@@ -37,9 +37,16 @@ zT = comT(3, 1);
 c(1, 6) = 4*xT/abs(zT) - 1
 c(1, 7) = (xT - p.w) / abs(zT) + 1
 % real part of closed loop eigenvalues must be negative
-% TODO : Correct the two following lines to use the proper syntax.
-stuff = generate_data(b);
-real_evals = real(eig(stuff.something.A));
+[A, B, C, D] = whipple_pull_force_abcd(b, b.v);
+data = generate_data('Browser', b.v, ...
+                     'simulate', false, ...
+                     'forceTransfer', {}, ...
+                     'fullSystem', false, ...
+                     'stateSpace', {A, B, C, D}, ...
+                     'display', 0);
+lateral_dev_loop = minreal(tf(data.closedLoops.Y.num,
+                              data.closedLoops.Y.den));
+real_evals = real(pole(lateral_dev_loop));
 c(1, 8) = real_evals(1);
 c(1, 9) = real_evals(2);
 c(1, 10) = real_evals(3);
@@ -53,7 +60,7 @@ c(1, 15) = p.mD + p.mH + p.mR + p.mF - 25.0;  % bicycle mass (D,H,R,F) no more t
 
 % Equality constraints
 % ceq(x) = 0
-ceq(1, 1) = p.kDyy - sqrt(p.kDxx^2 + p.kDzz^2)  % frame is planar
+ceq(1, 1) = p.kD11 - sqrt(p.kD22^2 + p.kD33^2)  % frame is planar
 ceq(2, 1) = b.IRyy - b.mR*b.rR^2  % wheel is a ring
 ceq(3, 1) = b.IRxx - b.IRyy / 2  % wheel is a ring
 ceq(4, 1) = b.IRzz - b.IRyy / 2  % wheel is a ring

### 2019-07-30 src/convert_principal_to_benchmark.m
Added more description of parameters.

@@ -30,28 +30,30 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
 % alphaD : Angle between the X and 1 axes, positive about Y
 %
 % person
-% mP
-% xP
-% yP
-% zP
-% kP11
-% kP22
-% kP33
-% alphaP
+% lP : length of person [m]
+% wP : longitudinal width of person [m]
+% mP : mass of person [kg]
+% xP : mass center [m]
+% yP : mass center [m]
+% zP : mass center [m]
+% kP11 : maximal central principal radius of gyration of person [m]
+% kP22 : intermediate central principal radius of gyration of person [m]
+% kP33 : minimal central principal radius of gyration of person [m]
+% alphaP : Angle between the X axis and the maximal principal axis [rad]
 %
 % front frame
-% mH
-% xH
-% yH
-% zH
+% mH : mass of front frame (handlebar/fork)
+% xH : X mass center [m]
+% yH : Y mass center [m]
+% zH : Z mass center [m]
 % kH11
 % kH22
 % kH33
 % alphaH
 %
 % rear wheel
-% rR
-% mR
+% rR : radius of rear wheel [m]
+% mR : mass of rear wheel [kg]
 % kR11 : Maximal principal radius of gyration, aligned with Y in the
 % case of the wheels
 % kR22 : Intermediate principal radius of gyration

### 2019-07-30 src/free_par_bounds.m
Corrected indexing to extract upper bounds.

@@ -32,4 +32,4 @@ bounds = [
    1.0, inf]; % mF
 
 lower = bounds(:, 1);
-upper = bounds(:, 1);
+upper = bounds(:, 2);

### 2019-07-30 src/free_parameters.m
Improved docstring of free_params.

@@ -1,6 +1,12 @@
 function x = free_parameters(principal_parameters)
 % FREE_PARAMETERS - Returns the free parameter vector, x, given the full set
 % of principal parameters.
+%
+% Inputs
+%   principal_parameters - Structure containing all of the principal
+%   parameters.
+% Outputs
+%   x - 23x1 double, vector of free parameters
 
 p = principal_parameters;
 

### 2019-07-30 src/default_principal_parameters.m
Transferred over Jason on Browser parameters from BicycleParameters.

@@ -1,42 +1,57 @@
 function p = default_principal_parameters()
 
-% based on Jason sitting on the Pista
+% based on Jason sitting on the Browser (parameters taken from
+% BicycleParameters/data)
 
-p.w = 0.989
-p.c = 0.0617300113263
+p.c = 0.0685808540382
+p.w = 1.121
+p.lam = 0.399680398707
 p.g = 9.81
-p.v = 5.0
-p.lam = 0.275762021815
-p.mH = 2.27
-p.mR = 1.38
-p.mF = 1.58
-p.rR = 0.332088156971
-p.rF = 0.333838861345
-p.xH = 0.906321313298
-p.zH = -0.732376356772
+p.v = 3.0
+
+p.mR = 3.11
+p.rR = 0.340958858855
+% TODO : convert to kR11, kR22, kR33
+p.IRxx = 0.0904114316323
+p.IRyy = 0.152391250767
+p.IRzz = p.IRxx;
+
+p.mF = 2.02
+p.rF = 0.34352982332
+% TODO : convert to kF11, kF22, kF33
+p.IFxx = 0.0883826870796
+p.IFyy = 0.149221207336
+p.IFzz = p.IFxx;
+
 % TODO : pick some reasonable values for these two.
-p.lP = nan;
-p.wP = nan;
+p.lP = 1.728;  % estimated from the BicycleParameters figure
+p.wP = 0.483;  % estimated from the BicycleParameters figure
+mP = 72.0
+xP = 0.2909
+yP = 0.0
+zP = -1.1091
+IPxx = 7.9985
+IPxz = -1.9272
+IPyy = 8.0689
+IPzz = 2.3624
 
 % TODO : Put into principal radius of gyration form.
-IFxx = 0.0553476418952
-IFyy = 0.106185287785
-IRxx = 0.0552268228143
-IRyy = 0.0764034890242
+p.mH = 3.22
+p.xH = 0.866949640247
+p.yH = 0.0
+p.zH = -0.748236400835
+p.IHxx = 0.25335539588
+p.IHxz = -0.0720217263293
+p.IHyy = 0.245827908036
+p.IHzz = 0.0955686343473
 
 % TODO : Put into principal radius of gyration form.
-IHxx = 0.0979884309288
-IHxz = -0.00440828467861
-IHyy = 0.0691907586693
-IHzz = 0.0396107377431
-
-% TODO: Break into separate bicycle and rider and put in principal radius of
[…]

### 2019-07-30 src/default_principal_parameters.m
Updated person parametes to reflect the most recent BicycleParameters calcs (prior ones were from earlier parameters).

@@ -26,14 +26,14 @@ p.IFzz = p.IFxx;
 % TODO : pick some reasonable values for these two.
 p.lP = 1.728;  % estimated from the BicycleParameters figure
 p.wP = 0.483;  % estimated from the BicycleParameters figure
-mP = 72.0
-xP = 0.2909
-yP = 0.0
-zP = -1.1091
-IPxx = 7.9985
-IPxz = -1.9272
-IPyy = 8.0689
-IPzz = 2.3624
+p.mP = 83.50000000000001
+p.xP = 0.3157679154070924
+p.yP = 0.0
+p.zP = -1.0989885659819278
+p.IPxx = 10.985244511152501
+p.IPyy = 11.257100519570363
+p.IPzz = 2.2692214245606905
+p.IPxz = -1.7021081054955438
 
 % TODO : Put into principal radius of gyration form.
 p.mH = 3.22

### 2019-07-31 src/free_parameters.m
Corrected radii of gyration in free_parameters.

@@ -16,8 +16,8 @@ x = [p.w;
      p.mD;
      p.xD;
      p.zD;
-     p.kDxx;
-     p.kDzz;
+     p.kD11;
+     p.kD33;
      p.alphaD;
      p.xP;
      p.zP;
@@ -25,9 +25,9 @@ x = [p.w;
      p.mH;
      p.xH;
      p.zH;
-     p.kHxx;
-     p.kHyy;
-     p.kHzz;
+     p.kH11;
+     p.kH22;
+     p.kH33;
      p.alphaH;
      p.rR;
      p.mR;

### 2019-07-31 src/default_principal_parameters.m
Swapped out the xyx inertias for the radii of gyration principal versions in default parameters.

@@ -3,55 +3,67 @@ function p = default_principal_parameters()
 % based on Jason sitting on the Browser (parameters taken from
 % BicycleParameters/data)
 
-p.c = 0.0685808540382
-p.w = 1.121
-p.lam = 0.399680398707
-p.g = 9.81
-p.v = 3.0
+% geometry and other
+p.c = 0.0685808540382;
+p.w = 1.121;
+p.lam = 0.399680398707;
+p.g = 9.81;
+p.v = 3.0;
 
-p.mR = 3.11
-p.rR = 0.340958858855
-% TODO : convert to kR11, kR22, kR33
-p.IRxx = 0.0904114316323
-p.IRyy = 0.152391250767
-p.IRzz = p.IRxx;
+% rear wheel
+p.mR = 3.11;
+p.rR = 0.340958858855;
+p.xR = 0.0;
+p.yR = 0.0;
+p.zR = -p.rR;
+IRxx = 0.0904114316323;
+IRyy = 0.152391250767;
+IRzz = IRxx;
+p.kR11 = sqrt(IRyy/p.mR);
+p.kR22 = sqrt(IRxx/p.mR);
+p.kR33 = sqrt(IRzz/p.mR);
 
-p.mF = 2.02
-p.rF = 0.34352982332
-% TODO : convert to kF11, kF22, kF33
-p.IFxx = 0.0883826870796
-p.IFyy = 0.149221207336
-p.IFzz = p.IFxx;
+% front wheel
+p.mF = 2.02;
+p.rF = 0.34352982332;
+p.xF = p.w;
+p.yF = 0.0;
+p.zF = -p.rF;
+IFxx = 0.0883826870796;
+IFyy = 0.149221207336;
+IFzz = IFxx;
+p.kF11 = sqrt(IFyy/p.mF);
+p.kF22 = sqrt(IFxx/p.mF);
+p.kF33 = sqrt(IFzz/p.mF);
 
-% TODO : pick some reasonable values for these two.
+% person
 p.lP = 1.728;  % estimated from the BicycleParameters figure
 p.wP = 0.483;  % estimated from the BicycleParameters figure
-p.mP = 83.50000000000001
-p.xP = 0.3157679154070924
-p.yP = 0.0
-p.zP = -1.0989885659819278
-p.IPxx = 10.985244511152501
-p.IPyy = 11.257100519570363
-p.IPzz = 2.2692214245606905
-p.IPxz = -1.7021081054955438
[…]

### 2019-07-31 src/convert_principal_to_benchmark.m
Fixed typos in conversion function.

@@ -74,8 +74,8 @@ p = principal_par;
 b.w = p.w;
 b.c = p.c;
 b.lam = p.lam;
-b.g = p.g
-b.v = p.v
+b.g = p.g;
+b.v = p.v;
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % rear frame [B]
@@ -88,11 +88,12 @@ b.yB = Bcom(2);
 b.zB = Bcom(3);
 
 % symmetry is assumed about the XZ plane
-ID123 = mD*diag([kD11^2, p.kD22^2, p.kD33^2]);
-IP123 = mP*diag([kP11^2, p.kP22^2, p.kP33^2]);
-IP = rotate_inerita_about_y(IP123, alphaP);
-ID = rotate_inerita_about_y(ID123, alphaD);
-IB = sum_central_inertias(p.mD, [xD; yD; zD], ID123, p.mP, [xP; yP; zP], IP);
+ID123 = p.mD*diag([p.kD11^2, p.kD22^2, p.kD33^2]);
+IP123 = p.mP*diag([p.kP11^2, p.kP22^2, p.kP33^2]);
+IP = rotate_inertia_about_y(IP123, -p.alphaP);
+ID = rotate_inertia_about_y(ID123, -p.alphaD);
+IB = sum_central_inertias(p.mD, [p.xD; p.yD; p.zD], ID123, ...
+                          p.mP, [p.xP; p.yP; p.zP], IP);
 b.IBxx = IB(1, 1);
 b.IByy = IB(2, 2);
 b.IBzz = IB(3, 3);
@@ -102,13 +103,13 @@ b.IBxz = IB(1, 3);
 % front frame [H]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 b.mH = p.mH;
-b.xH = p.xH:
+b.xH = p.xH;
 b.yH = p.yH;
-b.zH = p.zH:
+b.zH = p.zH;
 
 % symmetry is assumed about the XZ plane
-IH123 = mH*diag([kH11^2, p.kH22^2, p.kH33^2]);
[…]

### 2019-07-31 src/compute_constraints.m
Fixed typos in constraints function.

@@ -27,15 +27,16 @@ c(1, 3) = p.zP + p.wP / 2 * sin(p.alphaP);
 c(1, 4) = p.zP - p.lP / 2 * cos(p.alphaP);
 c(1, 5) = p.zP - p.wP / 2 * sin(p.alphaP);
 % the bike doesn't tip forward and backward when braking and accelerating
-comT = combine_mass_centers([b.mB, b.mH, b.mR, b.mF],
-                            [[b.xB; b.yB; b.zB],
-                             [b.xH; b.yH; b.zH],
-                             [b.xR; b.yR; b.zR],
-                             [b.xF; b.yF; b.zF]]);
+comT = combine_mass_centers([p.mD, p.mP, p.mH, p.mR, p.mF], ...
+                            [[p.xD; p.yD; p.zD], ...
+                             [p.xP; p.yP; p.zP], ...
+                             [p.xH; p.yH; p.zH], ...
+                             [p.xR; p.yR; p.zR], ...
+                             [p.xF; p.yF; p.zF]]);
 xT = comT(1, 1);
 zT = comT(3, 1);
-c(1, 6) = 4*xT/abs(zT) - 1
-c(1, 7) = (xT - p.w) / abs(zT) + 1
+c(1, 6) = 4*xT/abs(zT) - 1;
+c(1, 7) = (xT - p.w) / abs(zT) + 1;
 % real part of closed loop eigenvalues must be negative
 [A, B, C, D] = whipple_pull_force_abcd(b, b.v);
 data = generate_data('Browser', b.v, ...
@@ -60,10 +61,10 @@ c(1, 15) = p.mD + p.mH + p.mR + p.mF - 25.0;  % bicycle mass (D,H,R,F) no more t
 
 % Equality constraints
 % ceq(x) = 0
-ceq(1, 1) = p.kD11 - sqrt(p.kD22^2 + p.kD33^2)  % frame is planar
-ceq(2, 1) = b.IRyy - b.mR*b.rR^2  % wheel is a ring
-ceq(3, 1) = b.IRxx - b.IRyy / 2  % wheel is a ring
-ceq(4, 1) = b.IRzz - b.IRyy / 2  % wheel is a ring
-ceq(5, 1) = b.IFyy - b.mF*b.rF^2  % wheel is a ring
[…]

### 2019-07-31 src/test_functions.m
Added a test script for the functions.

@@ -1,0 +1,11 @@
+p = default_principal_parameters();
+
+b = convert_principal_to_benchmark(p);
+
+x = free_parameters(p);
+
+p_updated = update_principal_parameters(p, nan*ones(23,1));
+
+[lb, ub] = free_par_bounds();
+
+[c, ceq] = compute_constraints(x, p);

### 2019-07-31 src/compute_constraints.m
Fixed coding typos, now runs in matlab with HumanControl code.

@@ -45,7 +45,7 @@ data = generate_data('Browser', b.v, ...
                      'fullSystem', false, ...
                      'stateSpace', {A, B, C, D}, ...
                      'display', 0);
-lateral_dev_loop = minreal(tf(data.closedLoops.Y.num,
+lateral_dev_loop = minreal(tf(data.closedLoops.Y.num, ...
                               data.closedLoops.Y.den));
 real_evals = real(pole(lateral_dev_loop));
 c(1, 8) = real_evals(1);

### 2019-07-31 src/convert_principal_to_benchmark.m

@@ -81,7 +81,7 @@ b.v = p.v;
 % rear frame [B]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 b.mB = p.mD + p.mP;
-Bcom = combine_mass_centers([p.mD, p.mP],
+Bcom = combine_mass_centers([p.mD, p.mP], ...
                             [[p.xD; p.yD; p.zD], [p.xP; p.yD; p.zP]]);
 b.xB = Bcom(1);
 b.yB = Bcom(2);

### 2019-07-31 src/parallel_axis.m

@@ -14,7 +14,7 @@ function I = parallel_axis(Ic, m, d)
 a = d(1, 1);
 b = d(2, 1);
 c = d(3, 1);
-dMat = [b**2 + c**2,      -a * b,      -a * c;
-             -a * b, c**2 + a**2,      -b * c;
-             -a * c,      -b * c, a**2 + b**2];
+dMat = [b^2 + c^2,    -a * b,    -a * c;
+           -a * b, c^2 + a^2,    -b * c;
+           -a * c,    -b * c, a^2 + b^2];
 I = Ic + m * dMat;

### 2019-07-31 src/update_principal_parameters.m

@@ -10,8 +10,8 @@ p.lam    = x(3, 1);
 p.mD     = x(4, 1);
 p.xD     = x(5, 1);
 p.zD     = x(6, 1);
-p.kDxx   = x(7, 1);
-p.kDzz   = x(8, 1);
+p.kD11   = x(7, 1);
+p.kD33   = x(8, 1);
 p.alphaD = x(9, 1);
 p.xP     = x(10, 1);
 p.zP     = x(11, 1);
@@ -19,11 +19,17 @@ p.alphaP = x(12, 1);
 p.mH     = x(13, 1);
 p.xH     = x(14, 1);
 p.zH     = x(15, 1);
-p.kHxx   = x(16, 1);
-p.kHyy   = x(17, 1);
-p.kHzz   = x(18, 1);
+p.kH11   = x(16, 1);
+p.kH22   = x(17, 1);
+p.kH33   = x(18, 1);
 p.alphaH = x(19, 1);
 p.rR     = x(20, 1);
 p.mR     = x(21, 1);
 p.rF     = x(22, 1);
 p.mF     = x(23, 1);
+
+% TODO : These can be added as constraints, no? Or these shouldn't be
+% principal parameters.
+p.zR = -p.rR;
+p.xF = p.w;
+p.zF = -p.rF;

### 2019-07-31 src/convert_principal_to_benchmark.m
Correct the wheel inertia conversions.

@@ -120,11 +120,11 @@ b.IHxz = IH(1, 3);
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 b.rR = p.rR;
 b.mR = p.mR;
-% wheel is symmetric about XY, YZ, and XZ planes, thus no prodcuts of
+% wheel is symmetric about XY, YZ, and XZ planes, thus no products of
 % inertia, wheel is ring or disc like, thus Ixx=Izz and Iyy > Ixx
-b.IRxx = p.mR * p.kD22^2;
-b.IRyy = p.mR * p.kD11^2;
-b.IRzz = p.mR * p.kD33^2;
+b.IRxx = p.mR * p.kR22^2;
+b.IRyy = p.mR * p.kR11^2;
+b.IRzz = p.mR * p.kR33^2;
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % front wheel [F]
@@ -133,9 +133,9 @@ b.rF = p.rR;
 b.mF = p.mR;
 % wheel is symmetric about XY, YZ, and XZ planes, thus no prodcuts of
 % inertia, wheel is always ring or disc like, thus Ixx=Izz and Iyy > Ixx
-b.IFxx = p.mF * p.kD22^2;
-b.IFyy = p.mF * p.kD11^2;
-b.IFzz = p.mF * p.kD33^2;
+b.IFxx = p.mF * p.kF22^2;
+b.IFyy = p.mF * p.kF11^2;
+b.IFzz = p.mF * p.kF33^2;
 
 % return the benchmark parameters
 benchmark_par = b;

### 2019-07-31 src/convert_principal_to_benchmark.m
Fixed more conversion errors.

@@ -82,17 +82,17 @@ b.v = p.v;
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 b.mB = p.mD + p.mP;
 Bcom = combine_mass_centers([p.mD, p.mP], ...
-                            [[p.xD; p.yD; p.zD], [p.xP; p.yD; p.zP]]);
-b.xB = Bcom(1);
-b.yB = Bcom(2);
-b.zB = Bcom(3);
+                            [[p.xD; p.yD; p.zD], [p.xP; p.yP; p.zP]]);
+b.xB = Bcom(1, 1);
+b.yB = Bcom(2, 1);
+b.zB = Bcom(3, 1);
 
 % symmetry is assumed about the XZ plane
 ID123 = p.mD*diag([p.kD11^2, p.kD22^2, p.kD33^2]);
 IP123 = p.mP*diag([p.kP11^2, p.kP22^2, p.kP33^2]);
 IP = rotate_inertia_about_y(IP123, -p.alphaP);
 ID = rotate_inertia_about_y(ID123, -p.alphaD);
-IB = sum_central_inertias(p.mD, [p.xD; p.yD; p.zD], ID123, ...
+IB = sum_central_inertias(p.mD, [p.xD; p.yD; p.zD], ID, ...
                           p.mP, [p.xP; p.yP; p.zP], IP);
 b.IBxx = IB(1, 1);
 b.IByy = IB(2, 2);
@@ -129,8 +129,8 @@ b.IRzz = p.mR * p.kR33^2;
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % front wheel [F]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-b.rF = p.rR;
-b.mF = p.mR;
+b.rF = p.rF;
+b.mF = p.mF;
 % wheel is symmetric about XY, YZ, and XZ planes, thus no prodcuts of
 % inertia, wheel is always ring or disc like, thus Ixx=Izz and Iyy > Ixx
 b.IFxx = p.mF * p.kF22^2;

### 2019-07-31 src/is_equal.m
Added equality checks to the test_functions.m.

@@ -1,0 +1,5 @@
+function is_equal(a, b)
+if (abs(a-b) < 1e4*eps(min(abs(a),abs(b))))
+else
+    fprintf("%f is not equal to %f\n", a, b)
+end

### 2019-07-31 src/test_functions.m

@@ -2,6 +2,34 @@ p = default_principal_parameters();
 
 b = convert_principal_to_benchmark(p);
 
+is_equal(b.w, 1.121)
+is_equal(b.c, 0.0685808540382)
+is_equal(b.lam, 0.399680398707)
+is_equal(b.g, 9.81)
+is_equal(b.mB, 93.36)
+is_equal(b.xB, 0.3115627743494798)
+is_equal(b.yB, 0.0)
+is_equal(b.zB, -1.0397244169145061)
+is_equal(b.IBxx, 14.291736774539832)
+is_equal(b.IByy, 15.364340673709581)
+is_equal(b.IBxz, -1.621358873619717)
+is_equal(b.IBzz, 3.0399891044245795)
+is_equal(b.mH, 3.22)
+is_equal(b.xH, 0.866949640247)
+is_equal(b.zH, -0.748236400835)
+is_equal(b.IHxx, 0.25335539588)
+is_equal(b.IHyy, 0.245827908036)
+is_equal(b.IHzz, 0.0955686343473)
+is_equal(b.IHxz, -0.0720217263293)
+is_equal(b.mF, 2.02)
+is_equal(b.rF, 0.34352982332)
+is_equal(b.IFxx, 0.0883826870796)
+is_equal(b.IFyy, 0.149221207336)
+is_equal(b.mR, 3.11)
+is_equal(b.rR, 0.340958858855)
+is_equal(b.IRxx, 0.0904114316323)
+is_equal(b.IRyy, 0.152391250767)
+
 x = free_parameters(p);
 
 p_updated = update_principal_parameters(p, nan*ones(23,1));

### 2019-07-31 src/parallel_axis.m
dMat to D

@@ -14,7 +14,7 @@ function I = parallel_axis(Ic, m, d)
 a = d(1, 1);
 b = d(2, 1);
 c = d(3, 1);
-dMat = [b^2 + c^2,    -a * b,    -a * c;
-           -a * b, c^2 + a^2,    -b * c;
-           -a * c,    -b * c, a^2 + b^2];
-I = Ic + m * dMat;
+D = [b^2 + c^2,    -a * b,    -a * c;
+        -a * b, c^2 + a^2,    -b * c;
+        -a * c,    -b * c, a^2 + b^2];
+I = Ic + m * D;

### 2019-07-31 src/is_equal.m
Print a name in is_equal.

@@ -1,5 +1,5 @@
-function is_equal(a, b)
+function is_equal(a, b, n)
 if (abs(a-b) < 1e4*eps(min(abs(a),abs(b))))
 else
-    fprintf("%f is not equal to %f\n", a, b)
+    fprintf("%s: %f is not equal to expected %f\n", n, a, b)
 end

### 2019-07-31 src/test_functions.m
Use a loop for equality testing.

@@ -2,33 +2,40 @@ p = default_principal_parameters();
 
 b = convert_principal_to_benchmark(p);
 
-is_equal(b.w, 1.121)
-is_equal(b.c, 0.0685808540382)
-is_equal(b.lam, 0.399680398707)
-is_equal(b.g, 9.81)
-is_equal(b.mB, 93.36)
-is_equal(b.xB, 0.3115627743494798)
-is_equal(b.yB, 0.0)
-is_equal(b.zB, -1.0397244169145061)
-is_equal(b.IBxx, 14.291736774539832)
-is_equal(b.IByy, 15.364340673709581)
-is_equal(b.IBxz, -1.621358873619717)
-is_equal(b.IBzz, 3.0399891044245795)
-is_equal(b.mH, 3.22)
-is_equal(b.xH, 0.866949640247)
-is_equal(b.zH, -0.748236400835)
-is_equal(b.IHxx, 0.25335539588)
-is_equal(b.IHyy, 0.245827908036)
-is_equal(b.IHzz, 0.0955686343473)
-is_equal(b.IHxz, -0.0720217263293)
-is_equal(b.mF, 2.02)
-is_equal(b.rF, 0.34352982332)
-is_equal(b.IFxx, 0.0883826870796)
-is_equal(b.IFyy, 0.149221207336)
-is_equal(b.mR, 3.11)
-is_equal(b.rR, 0.340958858855)
-is_equal(b.IRxx, 0.0904114316323)
-is_equal(b.IRyy, 0.152391250767)
+expd_b.w = 1.121;
+expd_b.c = 0.0685808540382;
+expd_b.lam = 0.399680398707;
+expd_b.g = 9.81;
+expd_b.mB = 93.36;
+expd_b.xB = 0.3115627743494798;
+expd_b.yB = 0.0;
+expd_b.zB = -1.0397244169145061;
+expd_b.IBxx = 14.291736774539832;
+expd_b.IByy = 15.364340673709581;
+expd_b.IBxz = -1.621358873619717;
+expd_b.IBzz = 3.0399891044245795;
+expd_b.mH = 3.22;
+expd_b.xH = 0.866949640247;
+expd_b.yH = 0.0;
+expd_b.zH = -0.748236400835;
+expd_b.IHxx = 0.25335539588;
+expd_b.IHyy = 0.245827908036;
+expd_b.IHzz = 0.0955686343473;
+expd_b.IHxz = -0.0720217263293;
[…]

### 2019-07-31 src/sum_central_inertias.m
Whitespace.

@@ -13,4 +13,4 @@ function I = sum_central_inertias(m1, com1, I1, m2, com2, I2)
 % Outputs:
 %   I - combined central inertia tensor of both bodies
 com = combine_mass_centers([m1, m2], [com1, com2]);
-I = parallel_axis(I1, m1, com-com1) +  parallel_axis(I2, m2, com-com2);
+I = parallel_axis(I1, m1, com-com1) + parallel_axis(I2, m2, com-com2);

### 2019-07-31 src/compute_constraints.m
Changed constraint to use xyz coords for planarness.

@@ -61,7 +61,7 @@ c(1, 15) = p.mD + p.mH + p.mR + p.mF - 25.0;  % bicycle mass (D,H,R,F) no more t
 
 % Equality constraints
 % ceq(x) = 0
-ceq(1, 1) = p.kD11 - sqrt(p.kD22^2 + p.kD33^2);  % frame is planar
+ceq(1, 1) = b.IByy - sqrt(b.IBxx^2 + b.IBzz^2);  % frame is planar
 ceq(2, 1) = b.IRyy - b.mR*b.rR^2;  % wheel is a ring
 ceq(3, 1) = b.IRxx - b.IRyy / 2;  % wheel is a ring
 ceq(4, 1) = b.IRzz - b.IRyy / 2;  % wheel is a ring

### 2019-07-31 src/default_principal_parameters.m
Noted which default inertias are aligned with the Y axis.

@@ -19,7 +19,7 @@ p.zR = -p.rR;
 IRxx = 0.0904114316323;
 IRyy = 0.152391250767;
 IRzz = IRxx;
-p.kR11 = sqrt(IRyy/p.mR);
+p.kR11 = sqrt(IRyy/p.mR);  % aligned with Y axis
 p.kR22 = sqrt(IRxx/p.mR);
 p.kR33 = sqrt(IRzz/p.mR);
 
@@ -32,7 +32,7 @@ p.zF = -p.rF;
 IFxx = 0.0883826870796;
 IFyy = 0.149221207336;
 IFzz = IFxx;
-p.kF11 = sqrt(IFyy/p.mF);
+p.kF11 = sqrt(IFyy/p.mF);  % aligned with Y axis
 p.kF22 = sqrt(IFxx/p.mF);
 p.kF33 = sqrt(IFzz/p.mF);
 
@@ -44,7 +44,7 @@ p.xP = 0.3157679154070924;
 p.yP = 0.0;
 p.zP = -1.0989885659819278;
 p.kP11 = 0.36796654;
-p.kP22 = 0.36717241;
+p.kPyy = 0.36717241;  % aligned with the Y axis
 p.kP33 = 0.15276369;
 p.alphaP = -0.18617531745505142;
 
@@ -54,7 +54,7 @@ p.xH = 0.866949640247;
 p.yH = 0.0;
 p.zH = -0.748236400835;
 p.kH11 = 0.29556015;
-p.kH22 = 0.27630431;
+p.kH22 = 0.27630431;  % aligned with the Y axis
 p.kH33 = 0.14493343;
 p.alphaH = -0.3699481738478786;
 
@@ -63,7 +63,7 @@ p.mD = 9.86;
 p.xD = 0.275951285677;
 p.yD = 0.0;
 p.zD = -0.537842424305;
-p.kD11 = 0.36538846;
+p.kD11 = 0.36538846;  % this is aligned with the Y axis
 p.kD22 = 0.28586689;
 p.kD33 = 0.22079007;
 p.alphaD = -1.1722101094171953;

### 2019-07-31 src/convert_principal_to_benchmark.m
Noting this issue with the order of the principal inertias.

@@ -88,12 +88,16 @@ b.yB = Bcom(2, 1);
 b.zB = Bcom(3, 1);
 
 % symmetry is assumed about the XZ plane
-ID123 = p.mD*diag([p.kD11^2, p.kD22^2, p.kD33^2]);
+% NOTE : with the order of ID123 we have to ensure that the largest inertia
+% is pointing along the Y direction
+ID123 = p.mD*diag([p.kD22^2, p.kD11^2, p.kD33^2]);
+ID = rotate_inertia_about_y(ID123, -p.alphaD);
+
 IP123 = p.mP*diag([p.kP11^2, p.kP22^2, p.kP33^2]);
 IP = rotate_inertia_about_y(IP123, -p.alphaP);
-ID = rotate_inertia_about_y(ID123, -p.alphaD);
+
 IB = sum_central_inertias(p.mD, [p.xD; p.yD; p.zD], ID, ...
-                          p.mP, [p.xP; p.yP; p.zP], IP);
+                          p.mP, [p.xP; p.yP; p.zP], IP)
 b.IBxx = IB(1, 1);
 b.IByy = IB(2, 2);
 b.IBzz = IB(3, 3);

### 2019-07-31 src/convert_principal_to_benchmark.m
Changed to the handle the arbitrariness of the principal inertias in the xz plane.

@@ -24,9 +24,9 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
 % xD : X CoM location [m]
 % yD : Y CoM location [m]
 % zD : Z CoM location [m]
-% kD11 : Central principal radius of gyration
-% kD22 : Central principal radius of gyration
-% kD33 : Central principal radius of gyration
+% kDaa : Central principal radius of gyration
+% kDbb : Central principal radius of gyration
+% kDyy : Central principal radius of gyration
 % alphaD : Angle between the X and 1 axes, positive about Y
 %
 % person
@@ -36,9 +36,9 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
 % xP : mass center [m]
 % yP : mass center [m]
 % zP : mass center [m]
-% kP11 : maximal central principal radius of gyration of person [m]
-% kP22 : intermediate central principal radius of gyration of person [m]
-% kP33 : minimal central principal radius of gyration of person [m]
+% kPaa : maximal central principal radius of gyration of person [m]
+% kPbb : intermediate central principal radius of gyration of person [m]
+% kPyy : minimal central principal radius of gyration of person [m]
 % alphaP : Angle between the X axis and the maximal principal axis [rad]
 %
 % front frame
@@ -46,9 +46,9 @@ function benchmark_par = convert_principal_to_benchmark(principal_par)
 % xH : X mass center [m]
 % yH : Y mass center [m]
 % zH : Z mass center [m]
-% kH11
-% kH22
-% kH33
+% kHaa
+% kHbb
+% kHyy
 % alphaH
 %
 % rear wheel
@@ -88,16 +88,17 @@ b.yB = Bcom(2, 1);
 b.zB = Bcom(3, 1);
 
 % symmetry is assumed about the XZ plane
-% NOTE : with the order of ID123 we have to ensure that the largest inertia
-% is pointing along the Y direction
-ID123 = p.mD*diag([p.kD22^2, p.kD11^2, p.kD33^2]);
-ID = rotate_inertia_about_y(ID123, -p.alphaD);
[…]

### 2019-07-31 src/default_principal_parameters.m

@@ -43,19 +43,19 @@ p.mP = 83.50000000000001;
 p.xP = 0.3157679154070924;
 p.yP = 0.0;
 p.zP = -1.0989885659819278;
-p.kP11 = 0.36796654;
+p.kPaa = 0.36796654;
+p.kPbb = 0.15276369;
 p.kPyy = 0.36717241;  % aligned with the Y axis
-p.kP33 = 0.15276369;
-p.alphaP = -0.18617531745505142;
+p.alphaP = -0.18617531745505142;  % angle between X and the max of kPa and kpB
 
 % front frame
 p.mH = 3.22;
 p.xH = 0.866949640247;
 p.yH = 0.0;
 p.zH = -0.748236400835;
-p.kH11 = 0.29556015;
-p.kH22 = 0.27630431;  % aligned with the Y axis
-p.kH33 = 0.14493343;
+p.kHaa = 0.29556015;
+p.kHbb = 0.14493343;
+p.kHyy = 0.27630431;  % aligned with the Y axis
 p.alphaH = -0.3699481738478786;
 
 % rear frame (only) parameters
@@ -63,7 +63,7 @@ p.mD = 9.86;
 p.xD = 0.275951285677;
 p.yD = 0.0;
 p.zD = -0.537842424305;
-p.kD11 = 0.36538846;  % this is aligned with the Y axis
-p.kD22 = 0.28586689;
-p.kD33 = 0.22079007;
+p.kDaa = 0.28586689;
+p.kDbb = 0.22079007;
+p.kDyy = 0.36538846;  % this is aligned with the Y axis
 p.alphaD = -1.1722101094171953;

### 2019-07-31 src/test_functions.m
Added expected values for IFzz, IRzz.

@@ -26,10 +26,12 @@ expd_b.mF = 2.02;
 expd_b.rF = 0.34352982332;
 expd_b.IFxx = 0.0883826870796;
 expd_b.IFyy = 0.149221207336;
+expd_b.IFzz = 0.0883826870796;
 expd_b.mR = 3.11;
 expd_b.rR = 0.340958858855;
 expd_b.IRxx = 0.0904114316323;
 expd_b.IRyy = 0.152391250767;
+expd_b.IRzz = 0.0904114316323;
 expd_b.v = 3.0;
 
 fields = fieldnames(b);

### 2019-07-31 src/compute_constraints.m
Added kDyy as a free parameter.

@@ -61,7 +61,7 @@ c(1, 15) = p.mD + p.mH + p.mR + p.mF - 25.0;  % bicycle mass (D,H,R,F) no more t
 
 % Equality constraints
 % ceq(x) = 0
-ceq(1, 1) = b.IByy - sqrt(b.IBxx^2 + b.IBzz^2);  % frame is planar
+ceq(1, 1) = p.IDyy - sqrt(p.IDaa^2 + p.IDbb^2);  % frame is planar
 ceq(2, 1) = b.IRyy - b.mR*b.rR^2;  % wheel is a ring
 ceq(3, 1) = b.IRxx - b.IRyy / 2;  % wheel is a ring
 ceq(4, 1) = b.IRzz - b.IRyy / 2;  % wheel is a ring

### 2019-07-31 src/free_par_bounds.m

@@ -3,8 +3,8 @@ function [lower, upper] = free_par_bounds()
 % parameters.
 %
 % Outputs:
-%   lower - 23x1 double
-%   upper - 23x1 double
+%   lower - 24x1 double
+%   upper - 24x1 double
 
 bounds = [
   -inf, inf;  % w
@@ -13,8 +13,9 @@ bounds = [
    0.0, inf;  % mD
   -inf, inf;  % xD
   -inf, 0.0;  % zD
-  -inf, inf;  % kDxx
-  -inf, inf;  % kDzz
+  -inf, inf;  % kDaa
+  -inf, inf;  % kDbb
+  -inf, inf;  % kDyy
  -pi/2, pi/2; % alphaD
   -inf, inf;  % xP
   -inf, 0.0;  % zP
@@ -22,9 +23,9 @@ bounds = [
    0.0, inf;  % mH
   -inf, inf;  % xH
   -inf, 0.0;  % zH
-  -inf, inf;  % kHxx
+  -inf, inf;  % kHaa
+  -inf, inf;  % kHbb
   -inf, inf;  % kHyy
-  -inf, inf;  % kHzz
  -pi/2, pi/2; % alphaH
  0.127, inf;  % rR (10" OD wheel)
    1.0, inf;  % mR

### 2019-07-31 src/free_parameters.m

@@ -6,7 +6,7 @@ function x = free_parameters(principal_parameters)
 %   principal_parameters - Structure containing all of the principal
 %   parameters.
 % Outputs
-%   x - 23x1 double, vector of free parameters
+%   x - 24x1 double, vector of free parameters
 
 p = principal_parameters;
 
@@ -16,8 +16,9 @@ x = [p.w;
      p.mD;
      p.xD;
      p.zD;
-     p.kD11;
-     p.kD33;
+     p.kDaa;
+     p.kDbb;
+     p.kDyy;
      p.alphaD;
      p.xP;
      p.zP;
@@ -25,9 +26,9 @@ x = [p.w;
      p.mH;
      p.xH;
      p.zH;
-     p.kH11;
-     p.kH22;
-     p.kH33;
+     p.kHaa;
+     p.kHbb;
+     p.kHyy;
      p.alphaH;
      p.rR;
      p.mR;

### 2019-08-20 paper/paper.tex
paper/paper.tex
Reorder free param vector.
Most table drafts in and text rearranged.

@@
-\documentclass{article}
+\documentclass{bmd2019a}
 
@@
 \usepackage{amsmath}
+\usepackage{booktabs}
 
@@
 
-\begin{center}
-  \fontsize{14}{20}{\bf Expanded Optimization for Discovering Optimal Lateral
-  Handling Bicycles}
-\end{center}
+\begin{flushleft}
+{\fontsize{16pt}{20pt}\selectfont%
+  Expanded Optimization for Discovering\\}
+{\fontsize{16pt}{20pt}\selectfont%
+  Optimal Lateral Handling Bicycles}
+\end{flushleft}
 
-\begin{center}
-  \normalsize{\bf{Jason K. Moore, Mont Hubbard,  Ronald A. Hess}}
-\end{center}
+%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
+\begin{flushleft}
+  {\fontsize{12}{14}{Jason K. Moore, Mont Hubbard, and Ronald A. Hess}\\}
+  \textit{Mechanical and Aerospace Engineering\\
+          University of California, Davis\\
+          One Shields Avenue, Davis, CA, USA 95817\\
+          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu}
+\end{flushleft}
 
-\begin{center}
-  \begin{tabular}{c}
-    Department of Mechanical and Aerospace Engineering\\
-    University of California, Davis\\
-    One Shields Avenue, Davis, CA, USA 95616\\
-    e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu\\
+\section*{Abstract}
+%
+Physical design features of ground vehicles can affect their lateral handling
+qualities. Geometry, mass, and mass distribution of the vehicle's primary
+components as well as tire characteristics are primary contributors to poor and
+good handling due to their important influence on the vehicle's dynamics. In
+past work, we have presented a theoretical and computational framework for
+assessing the lateral task-independent handling qualities of simplified single
+track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
+that minimizing our handling quality metric (HQM) can produce theoretically
+optimal handling designs when only four geometric parameters are explored as
+the optimization variables~\cite{Moore2016}. The present work's goal is to
+expand this optimization problem to 24 geometry, mass, and inertial parameters
+as the optimization variables so that a broader search space of optimal bicycle
+designs is considered. This task is complicated by the fact that we desire to
+limit the search space to physically realizable designs. Doing so introduces 35
+linear and non-linear constraint equations on the optimization variables. We
+formulate a constrained optimization problem and use nonlinear programming to
+discover optimal, yet realizable, bicycle designs.
+
+\section*{Introduction}
+%
+We introduced a method of optimizing four geometric parameters of a bicycle's
+design to minimize the so called Handling Quality Metric
+(HQM)~\cite{Moore2016}. Here we expand that method to optimize all X geometric
+and inertial parameters of the benchmark parameterimization of the linear
+Whipple-Carvallo bicycle model under constraints that guarantee a physically
+realizable bicycle. This improves over the prior work by expanding the search
+space with many more parameters and the guarntee of realizability.
+
+\section*{Problem Definition}
+
+We search for 24 optimization variables that minimize the Handling Quality
+Metric defined in ~\cite{Hess2012}.
+
+Our formulation exposes 24 optimization variables
+$\mathbf{p}\in\mathbb{R}^{24}$ that are both reformulations of and additions to
+the parameters of the benchmark bicycle~\cite{Meijaard2007}. We separate the
+rider's inertial parameters from the rear frame of the bicycle for the purposes
+of rider/bicycle relative configuration but do not introduce any more than the
+benchmark's four rigid bodies. The variables are as follows;
+%
+\begin{itemize}
+  \itemsep-0.25em
+  \item wheelbase, trail, steer axis tilt, wheel radii
+  \item somersault angle of the person (same direction as bicycle pitch)
+  \item masses of the front frame, rear frame, front wheel, and rear wheel
+  \item mass center locations of the front frame, rear frame, and rider
+  \item principal central radii of gyration of front and rear frames
+  \item principal inertia axis pitch angles of the front frame, rear frame, and rider
+\end{itemize}
+
+We formulate bounds on the variables as well as equality and inequality
+constraints among the variables to ensure a physically realizable bicycle.
+Below the constraints are defined and grouped by the associated rigid body or
+collection thereof:
+%
+\begin{description}
+
+  \item[Vehicle] The resulting combination of the five rigid bodies.
+    \begin{itemize}
+      \itemsep0em
+      \item The physical extents of the bicycle and rider must exist above the
+        ground plane.
+      \item Both bicycle and rider are symmetric about the sagittal
+        plane.
+      \item The mass of each bicycle rigid body is positive, greater than a
+        minimum value, and the total mass is below a reasonably lift-able
+        amount.
+      \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
+        radius.
+      \item The bicycle cannot topple forward during hard braking or backward
+        during hard acceleration, i.e. -1 g < acceleration < 1 g.
+    \end{itemize}
+
+  \item[Wheels] Both front and rear wheels have identical constraints.
+    \begin{itemize}
+      \itemsep0em
+      \item Wheel radius must be greater than a minimum value.
+      \item Wheels are inertially wheel-like, i.e.  $I_{rot}=mr^2,
+        I_{rad}=I_{rot}/2$.
+    \end{itemize}
+
+  \item[Frames] Both front and rear frames have identical constraints.
+    \begin{itemize}
+      \itemsep0em
+      \item A pair of principal directions lie in the roll/yaw
+        plane to maintain symmetry and one of these axes is within $\pm \pi/4$
+        rad of the ground plane.
+      \item The mass and inertia of the frames are positive and large enough to
+        be constructed from materials in a space frame of specified minimal
+        density.
+      \item The frames are planar in nature, i.e.
+        $k_{lateral}^2=k_{longitudinal}^2+k_{vertical}^2$, where $k$ is the
+        radius of gyration.
+    \end{itemize}
+
+  \item[Rider] A single rigid body represents the rider.
+    \begin{itemize}
+      \itemsep0em
+      \item Rider mass is that of an average person.
+      \item The rider's joint angles are fixed in a nominal configuration
+        typical of upright bicycling and the resulting mass distribution is
+        derived from standard body segment estimation methods.
+      \item One rider principal axis is parallel with the bicycle pitch axis to
+        enforce symmetry.
+    \end{itemize}
+
+\end{description}
+
+The constraint statements above translate into upper and lower bounds on 12 of
+the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
+constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
+expressed in the standard non-linear programming form.
+%
+\begin{equation}
+  \begin{aligned}
+    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
+    & \text{subject to} & & \\
+    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
+    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
+  \end{aligned}
+\end{equation}
+
+\subsection*{Parameters}
+%
+We introduce a new parameterization of the linear Whipple-Carvallo bicycle
+model. This parameterization differs from the benchmark
+parameterization~\cite{Meijaard2007} in that the person and rear frame are
+separate rigid bodies each with their on inertial parameters and that the
+inertial parameters of all bodies are expressed in terms of principal central
+radii of gyration so that the mass is decoupled from the inertia. We also
+introduce two simple dimensional parameters that define the geometric extents
+of the person. Table~\ref{tab:parameters} provides the parameters and the
+initial values which are derived from.
+
+% TODO : Create a table of the parameters and values.
+\begin{table}
+  \caption{Parameters and their default values.}
+  \begin{tabular}{llll}
+    \toprule
+    Variable & Value & Units & Description \\
+    \midrule
+    $c$ &  0.068581 & \si{\meter} & Trail \\
+    $w$ &  1.1210 & \si{\meter} & Wheelbase \\
+    $\lambda$ &  0.39968 & \si{\radian} & Steer axis tilt \\
+    $g$ &  9.81 & \si{\meter\per\second\squared} & Acceleration due to gravity \\
+    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
+    Rear Wheel [R] & & \\
+    \midrule
+    $m_R$ &  3.11 & \si{\kilogram} & mass \\
+    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
+    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
+    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
+    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
+    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
+    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
+    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
+    Front Wheel [F] & & \\
+    \midrule
+    $m_F$ &  2.02 & \si{\kilogram} & Mass \\
+    $r_F$ &  0.34353 & \si{\meter} & Radius \\
+    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
+    $y_F$ & 0.0 & \si{\meter} & Mass center \\
+    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
+    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
+    \midrule
+    $l_P$ &  1.7280 & \si{\meter} & Body length \\
+    $w_P$ &  0.48300 & \si{\meter} & Body width \\
+    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
+    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
+    $y_P$ & 0.0 & \si{\meter} & Mass center \\
+    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
+    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
+    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
+    \midrule
+    $m_H$ & 3.2200 & \si{\kilogram} & Mass \\
+    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
+    $y_H$ & 0.0 & \si{\meter} & Mass center \\
+    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
+    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
+    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
+    \midrule
+    $m_D$ &  9.8600 & \si{\kilogram} & Mass \\
+    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
+    $y_D$ & 0 & \si{\meter} & Mass center \\
+    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
+    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
+    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
+    \bottomrule
   \end{tabular}
-\end{center}
+\end{table}
 
-\appendix
+These parameters can be transformed into the benchmark parameterization
+r adily, but not vice versa.
 
-\section{Free Parameters and Constraints}
+We search for the optimal values of a subset of 23 of the principal parameters.
 
-\subsection{Rear Frame [D]}
+\begin{align}
+  \mathbf{p} =
+    [w \quad
+     c \quad
+     \lambda \quad
+     m_D \quad
+     x_D \quad
+     z_D \quad
+     k_{Daa} \quad
+     k_{Dbb} \quad
+     \alpha_D \quad
+     x_P \quad
+     z_P \quad
+     \alpha_P \quad
+     m_H \quad
+     x_H \quad
+     z_H \quad
+     k_{Haa} \quad
+     k_{Hbb} \quad
+     k_{Hyy} \quad
+     \alpha_H \quad
+     r_R \quad
+     m_R \quad
+     r_F \quad
+     m_F]
+\end{align}
 
-There are 5 free parameters associated with the rear frame's inertial
+\subsection*{Constraints}
+%
+The optimal principal parameters are subject to a set of constraints designed
+to ensure a physically realizable bicycle is obtained.
+
+\begin{table}
+  \caption{Inequality constraints}
+  \begin{tabular}{lll}
+    \toprule
+    Constraint & Equation & Description \\
+    \midrule
+    $c_1$    & $\sqrt{I_{Hxx}^2+I_{Hzz}^2} \geq I_{Hyy}$           & Narrow front frame. \\
+    $c_2$    & $0 \geq z_P + l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_3$    & $0 \geq z_P + w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_4$    & $0 \geq z_P - l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_5$    & $0 \geq z_P - w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_6$    & $x_T \geq |z_T|/4$                                  & Maximum acceleration of $1/4g$. \\
+    $c_7$    & $w-x_T \geq 3/4|z_T|$                               & Maximum deceleration of $3/4g$. \\
+    $c_8$    & $2k_{Hyy} \geq \sqrt{(x_H - w)^2 + (z_H + r_F)^2}$  & Minimal inertial spread. \\
+    $c_9$    & $2k_{Dyy} \geq \sqrt{(x_D - 0)^2 + (z_D + r_R)^2}$  & Minimal inertial spread. \\
+    $c_{10}$ & $w \geq r_F +r_R$                                   & Non-overlapping wheels. \\
+    $c_{11}$ & $25\si{\kg} \geq m_D + m_H + m_R + m_F$             & Maximum bicycle mass. \\
+    $c_{12}$ & $-z_D \geq 1.4 k_{Dyy}$                             & Rear frame cannot penetrate ground. \\
+    $c_{13}$ & $-z_H \geq 1.4 k_{Hyy}$                             & Front frame cannot penetrate ground. \\
+    $c_{14,\ldots,21}$ & $0 \geq s_1,\ldots,s_8$                   & Closed loop stability. \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Equality constraints}
+  \begin{tabular}{lll}
+    \toprule
+    Constraint & Equation & Description \\
+    \midrule
+    $g_1$ & $I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2}$ & Rear frame is planar. \\
+    $g_2$ & $k_{Ryy} = r_R$  & Rear wheel is a ring \\
+    $g_3$ & $k_{Raa} = k_{Ryy}/2$ & Rear wheel is a ring \\
+    $g_4$ & $k_{Rbb} = k_{Ryy}/2$ & Rear wheel is a ring \\
+    $g_5$ & $k_{Fyy} = r_F$  & Front wheel is a ring \\
+    $g_6$ & $k_{Faa} = k_{Fyy}/2$ & Front wheel is a ring \\
+    $g_7$ & $k_{Fbb} = k_{Fyy}/2$ & Front wheel is a ring \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \caption{Parameter Bounds}
+  \begin{tabular}{ccccc}
+    \toprule
+    Min & & Parmeter & & Max\\
+    \midrule
+    $-\infty$ & $\leq$ & $w       $  &  $\leq$  &  $\infty$ \\
+    $-\infty$ & $\leq$ & $c       $  &  $\leq$  &  $\infty$ \\
+    $-\pi/2 $ & $\leq$ & $\lambda $  &  $\leq$  &  $\pi/2$ \\
+    $1.0    $ & $\leq$ & $m_D     $  &  $\leq$  &  $\infty$ \\
+    $-\infty$ & $\leq$ & $x_D     $  &  $\leq$  &  $\infty$ \\
+    $-\infty$ & $\leq$ & $z_D     $  &  $\leq$  &  $0.0$    \\
+    $0.0    $ & $\leq$ & $k_{Daa} $  &  $\leq$  &  $\infty$ \\
+    $0.0    $ & $\leq$ & $k_{Dbb} $  &  $\leq$  &  $\infty$ \\
+    $-\pi/2 $ & $\leq$ & $\alpha_D$  &  $\leq$  &  $\pi/2$  \\
+    $-\infty$ & $\leq$ & $x_P     $  &  $\leq$  &  $\infty$ \\
+    $-\infty$ & $\leq$ & $z_P     $  &  $\leq$  &  $0.0$    \\
+    $-\pi/2 $ & $\leq$ & $\alpha_P$  &  $\leq$  &  $\pi/2$ \\
+    $0.25   $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\
+    $-\infty$ & $\leq$ & $x_H     $  &  $\leq$  &  $\infty$ \\
+    $-\infty$ & $\leq$ & $z_H     $  &  $\leq$  &  $0.0$ \\
+    $0.0    $ & $\leq$ & $k_{Haa} $  &  $\leq$  &  $\infty$ \\
+    $0.0    $ & $\leq$ & $k_{Hbb} $  &  $\leq$  &  $\infty$ \\
+    $0.0    $ & $\leq$ & $k_{Hyy} $  &  $\leq$  &  $\infty$ \\
+    $-\pi/2 $ & $\leq$ & $\alpha_H$  &  $\leq$  &  $\pi/2$ \\
+    $0.127  $ & $\leq$ & $r_R     $  &  $\leq$  &  $\infty$ \\
+    $1.0    $ & $\leq$ & $m_R     $  &  $\leq$  &  $\infty$ \\
+    $0.127  $ & $\leq$ & $r_F     $  &  $\leq$  &  $\infty$ \\
+    $1.0    $ & $\leq$ & $m_F     $  &  $\leq$  &  $\infty$ \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\subsection*{Rear Frame [D]}
+
+There are five free parameters associated with the rear frame's inertial
 characteristics:
@@
   \item[$z_D$] Location along the $Z$ axis of the rear frame center of mass.
-  \item[$k_{Dxx}$] Central radius of gyration with respect to the $X$ axis
-    where $k_{Dxx}^2=I_{Dxx} / m_D$.
-  \item[$k_{Dzz}$] Central radius of gyration with respect to the $Z$ axis
-    where $k_{Dzz}^2=I_{Dzz} / m_D$.
-  \item[$\alpha_D$] Angle between the $X$ axis and the primary principal
-    inertia axis.
+  \item[$k_{Daa}$] Maximum principal central radius of gyration in the $XZ$
+    plane.
+  \item[$k_{Dbb}$] Minimum principal central radius of gyration in the $XZ$
+    plane.
+  \item[$\alpha_D$] Angle between the $X$ axis and the maximum principal axis
+    that lies in the $XZ$ plane.
 \end{description}
 
-Assumptions:
-
-\begin{itemize}
-  \item The body is symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} =
-    I_{Dyz} = 0$.
-  \item We require that the center of mass not penetrate the ground, i.e. $z_D
-    < 0$.
-  \item We require that the mass be positive, i.e. $m_D > 0$.
-  \item We allow for any angular orientation of the $XZ$ principal directions
-    but limit the angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
-  \item We enforce that the body is planar with $k_{Dyy} = \sqrt{k_{Dxx}^2 +
-    k_{Dyy}^2}$.
-\end{itemize}
+Several constraints are set for the rear frame. We constrain the rear frame to
+be planar, $g_1$ and symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} =
+I_{Dyz} = 0$. We prevent the rear frame from penetrating the ground by limiting
+the inertial spread with respect to its mass center, $c_{12}$, but also set a
+minimum inertial spread to ensure a frame can span from the rear wheel to the
+mass center of the rear frame, $c_9$. The spread factor in $c_{12}$ of 1.4 is
+based on the ratio of geometrical spread of a typical bicycle frame and its
+radius of gyration. Several parameters are bounded. We require the rear frame
+mass to be positive, the center of mass not penetrate the ground, and we allow
+for any angular orientation of the $XZ$ principal directions but limit the
+angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
 
@@
 nominal configuration for pedaling, i.e. an average normal everyday riding
-position. The person is assumed to be symmetric about the $XZ$ plane. There are
-four fixed parameters that enforce this assumption:
+position on a typical bicycle. The person is assumed to be symmetric about the
+$XZ$ plane. We allow the rider to be rotated about the $Y$ axis and positioned
+anywhere within the plane of symmetry above the ground. Three free parameters
+are associated with these assumptions:
+
+\begin{description}
+  \item[$x_P$] Location along the $X$ axis of the person's center of mass.
+  \item[$z_P$] Location along the $Z$ axis of the person's center of mass.
+  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal
+    inertia axis of the person.
+\end{description}
+
+There are four fixed parameters that enforce this assumption:
 
@@
 
-We allow the rider to be rotated about the $Y$ axis and positioned anywhere
-within the plane of symmetry. Three free parameters are associated with these
-assumptions:
-
-\begin{description}
-  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal
-    inertia axis of the person.
-  \item[$x_P$] Location along the $X$ axis of the person's center of mass.
-  \item[$z_P$] Location along the $Z$ axis of the person's center of mass.
-\end{description}
-
 To prevent the rider from being positioned and oriented such that their body is
@@
 along the principal axis to the tip of the toes and $w_P$ is the distance along
-the second principal axes to the tip of the hands.
-
-\begin{align}
-  z_P + l_P / 2 \cos{\alpha_P} & \leq 0 \\
-  z_P + w_P / 2 \sin{\alpha_P} & \leq 0 \\
-  z_P - l_P / 2 \cos{\alpha_P} & \leq 0 \\
-  z_P - w_P / 2 \sin{\alpha_P} & \leq 0
-\end{align}
+the second principal axes to the tip of the hands. Constraints $c_2,\ldots,c_5$.
 
@@
   \item[$m_H$] The mass of the rear frame.
-  \item[$k_{Hxx}$] Central radius of gyration with respect to the $X$ axis
+  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
+  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
+  \item[$k_{Haa}$] Central radius of gyration with respect to the $X$ axis
     where $k_{Hxx}^2=I_{Hxx} / m_H$.
@@
     where $k_{Hyy}^2=I_{Hyy} / m_H$.
-  \item[$k_{Hzz}$] Central radius of gyration with respect to the $X$ axis
+  \item[$k_{Hbb}$] Central radius of gyration with respect to the $X$ axis
     where $k_{Hzz}^2=I_{Hzz} / m_H$.
@@
     inertia axis.
-  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
-  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
 \end{description}
 
-Assumptions:
+We allow for any angular orientation of the $XZ$ principal directions but limit
+the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.  We prevent the
+rear frame from penetrating the ground by limiting the inertial spread with
+respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
+ensure a frame can span from the rear wheel to the mass center of the rear
+frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
+geometrical spread of a typical bicycle frame and its radius of gyration. The
+front frame is not planar but is narrow with respect to the $XZ$ plane, which
+is enforced by constraint $c_1$.
 
-\begin{itemize}
-  \item We allow for any angular orientation of the $XZ$ principal directions
-    but limit the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.
-\end{itemize}
+\subsection{Front [F] and Rear [R] Wheels}
 
-\subsection{Rear Wheel [R]}
+We enforce the assumption that both wheels have moments of inertia of that of a
+simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than
+some minimal size based on small bicycle wheel.
 
-We enforce the assumption that the wheel should be wheel-like, i.e. $I_{Ryy} =
-m_R r_R^2$ and $I_{Rxx} = I_{Rzz} = I_{Ryy} / 2$ thus there are two free
-parameters:
+\subsection{Total Bike [T]}
 
-\begin{description}
-  \item[$r_R$] Outer radius of the rear wheel.
-  \item[$m_R$] Mass of the rear wheel.
-\end{description}
+\subsubsection{Parameters}
 
-We also enforce that the mass and radius should be positive with:
+There are three remaing parameters associated with the fundamental geometry of
+the bicycle: wheelbase $w$, trail $c$, and steer axis tilt $\lambda$ all the
+same as defined in~\cite{Meijaard2007}.
 
-\begin{align}
-  m_R \geq 0 \\
-  r_R \geq 0
-\end{align}
-
-\subsection{Front Wheel}
-
-Same as rear wheel.
-
-\subsection{Total Bike}
-
-There are three remaing free parameters associated with the fundamental
-geometry of the bicycle:
-
-\begin{description}
-  \item[$w$] Wheelbase
-  \item[$c$] Trail
-  \item[$\lambda$] Steer axis tilt
-\end{description}
-
-To prevent the wheels from overlapping we introduce:
-
-\begin{align}
-  w > r_F + r_R
-\end{align}
+\subsubsection{Bounds}
 
@@
 
-We require that the bicycle be liftable by an average person with:
+\subsubsection{Constraints}
+
+We introduce a constraint $c_{10}$ that prevents the wheels from physically
+overlapping and require that the bicycle be liftable by an average person,
+$c_{11}$. Finally, we require that the bicycle not topple forward during hard
+breaking or backward during hard acceleration.
 
 \begin{align}
-  m_T = m_H + m_D + m_P + m_F + m_R \leq m_P + 25 \si{kg}
+  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}
 \end{align}
 
-Finally, we require that the bicycle not topple forward during hard breaking
-and backward during hard acceleration where:
+This translates to two constraints, $c_6,c_7$ that bound the total center of
+mass $(x_T,z_T)$ in a triangle in the $XZ$ plane.
 
-\begin{align}
-  -g < a < \frac{g}{4}
-\end{align}
-
-This translates to two constraints that bound the total center of mass in a triangle in teh $XZ$ plane:
-
-\begin{align} \frac{x_T}{|z_T|} < \frac{1}{4} \\ \frac{x_T - w}{|z_T|} < -1
-\end{align}
+Lastly, we constrain the eight closed loop eigenvalues associated with the
+controller in ~\cite{Hess2012} to be stable, i.e. have negative real parts.
+These are expressed in constraints $c_{14},\ldots,c_{21}$. Closed loop
+stability is required for the HQM to provide a meaningful result.
 
@@
   \begin{bmatrix}
+    % primary geometry
+    w \\
+    c \\
+    \lambda
     % front frame
@@
     m_F \\
-    % primary geometry
-    w \\
-    c \\
-    \lambda
   \end{bmatrix}
@@
 
+\bibliographystyle{acm}
+\bibliography{BMD2019}
+
+\section*{Solving}
+\section*{Results}
+\section*{Discussion and Conclusion}
+
 \end{document}

### 2019-08-21 paper/paper.tex
Improved parameter section opening.

@@
 \end{flushleft}
+
+% - Abstract
+% - Introduction
+% - Problem Definition
+%   - Optimization
+%   - Parameterization
+%   - Bounds
+%   - Constraints
+% - Solutions
+%   - Parameters
+%   - Eigenvalues
+%   - HQM
+% - Discussion and Conclusion
 
@@
 \section*{Problem Definition}
+%
+We search for 24 optimization variables that minimize the Handling Quality
+Metric defined in~\cite{Hess2012}.
 
-We search for 24 optimization variables that minimize the Handling Quality
-Metric defined in ~\cite{Hess2012}.
+\subsection*{Optimization}
+%
+The constraint statements above translate into upper and lower bounds on 12 of
+the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
+constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
+expressed in the standard non-linear programming form.
+%
+\begin{equation}
+  \begin{aligned}
+    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
+    & \text{subject to} & & \\
+    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
+    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
+  \end{aligned}
+\end{equation}
 
-Our formulation exposes 24 optimization variables
-$\mathbf{p}\in\mathbb{R}^{24}$ that are both reformulations of and additions to
-the parameters of the benchmark bicycle~\cite{Meijaard2007}. We separate the
-rider's inertial parameters from the rear frame of the bicycle for the purposes
-of rider/bicycle relative configuration but do not introduce any more than the
-benchmark's four rigid bodies. The variables are as follows;
+\subsection*{Bicycle Model Parameters}
+%
+Our problem formulation relies on a new parameterization that reflects both a
+reformulation of and addition to the benchmark parameterization of the linear
+Whipple-Carvallo bicycle model~\cite{Meijaard2007}. This parameterization
+differs from the benchmark parameterization in that the person and rear frame
+are treated as separate but fixed rigid bodies each with their on inertial
+parameters and that the inertial parameters of all bodies are expressed in
+terms of central principal radii of gyration so that the mass is decoupled from
+the inertia. We also introduce two simple dimensional parameters that define
+the geometric extents of the person. Table~\ref{tab:parameters} provides the
+parameters and the initial values which are derived from the measurements of a
+Batuvus Browser Bicycle and the rider, Jason, presented in~\cite{Moore2012}.
+This allows the position and orientation of the rider to be specified
+independtly of the rear frame but does not introduce any more than the
+benchmark's four rigid bodies.
+
+The variables are as follows:
 %
 \begin{itemize}
-  \itemsep-0.25em
   \item wheelbase, trail, steer axis tilt, wheel radii
@@
 
+\begin{table}
+  \caption{Parameters and their default values.}
+  \begin{tabular}{llll}
+    \toprule
+    Variable & Value & Units & Description \\
+    \midrule
+    $c$ &  0.068581 & \si{\meter} & Trail \\
+    $w$ &  1.1210 & \si{\meter} & Wheelbase \\
+    $\lambda$ &  0.39968 & \si{\radian} & Steer axis tilt \\
+    $g$ &  9.81 & \si{\meter\per\second\squared} & Acceleration due to gravity \\
+    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
+    Rear Wheel [R] & & \\
+    \midrule
+    $m_R$ &  3.11 & \si{\kilogram} & mass \\
+    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
+    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
+    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
+    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
+    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
+    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
+    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
+    Front Wheel [F] & & \\
+    \midrule
+    $m_F$ &  2.02 & \si{\kilogram} & Mass \\
+    $r_F$ &  0.34353 & \si{\meter} & Radius \\
+    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
+    $y_F$ & 0.0 & \si{\meter} & Mass center \\
+    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
+    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
+    Person [P] & & \\
+    \midrule
+    $l_P$ &  1.7280 & \si{\meter} & Body length \\
+    $w_P$ &  0.48300 & \si{\meter} & Body width \\
+    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
+    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
+    $y_P$ & 0.0 & \si{\meter} & Mass center \\
+    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
+    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
+    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
+    Front Frame [H] & & \\
+    \midrule
+    $m_H$ & 3.2200 & \si{\kilogram} & Mass \\
+    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
+    $y_H$ & 0.0 & \si{\meter} & Mass center \\
+    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
+    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
+    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
+    Rear Frame [D] & & \\
+    \midrule
+    $m_D$ &  9.8600 & \si{\kilogram} & Mass \\
+    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
+    $y_D$ & 0 & \si{\meter} & Mass center \\
+    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
+    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
+    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
+    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+$\mathbf{p}\in\mathbb{R}^{24}$These parameters can be transformed into the benchmark parameterization
+r adily, but not vice versa.
+We search for the optimal values of a subset of 23 of the principal parameters.
+
+\begin{align}
+  \mathbf{p} = &
+    [w \quad
+     c \quad
+     \lambda \quad
+     m_D \quad
+     x_D \quad
+     z_D \quad
+     k_{Daa} \quad
+     k_{Dbb} \quad
+     \alpha_D \quad
+     x_P \quad
+     z_P \quad
+     \alpha_P \quad \\
+ &    m_H \quad
+     x_H \quad
+     z_H \quad
+     k_{Haa} \quad
+     k_{Hbb} \quad
+     k_{Hyy} \quad
+     \alpha_H \quad
+     r_R \quad
+     m_R \quad
+     r_F \quad
+     m_F]
+\end{align}
+
+\subsection*{Bounds and Constraints}
+%
+The optimal principal parameters are subject to a set of constraints designed
+to ensure a physically realizable bicycle is obtained.
 We formulate bounds on the variables as well as equality and inequality
@@
 \begin{description}
-
   \item[Vehicle] The resulting combination of the five rigid bodies.
@@
 \end{description}
-
-The constraint statements above translate into upper and lower bounds on 12 of
-the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
-constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
-expressed in the standard non-linear programming form.
-%
-\begin{equation}
-  \begin{aligned}
-    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
-    & \text{subject to} & & \\
-    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
-    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
-  \end{aligned}
-\end{equation}
-
-\subsection*{Parameters}
-%
-We introduce a new parameterization of the linear Whipple-Carvallo bicycle
-model. This parameterization differs from the benchmark
-parameterization~\cite{Meijaard2007} in that the person and rear frame are
-separate rigid bodies each with their on inertial parameters and that the
-inertial parameters of all bodies are expressed in terms of principal central
-radii of gyration so that the mass is decoupled from the inertia. We also
-introduce two simple dimensional parameters that define the geometric extents
-of the person. Table~\ref{tab:parameters} provides the parameters and the
-initial values which are derived from.
-
-% TODO : Create a table of the parameters and values.
-\begin{table}
-  \caption{Parameters and their default values.}
-  \begin{tabular}{llll}
-    \toprule
-    Variable & Value & Units & Description \\
-    \midrule
-    $c$ &  0.068581 & \si{\meter} & Trail \\
-    $w$ &  1.1210 & \si{\meter} & Wheelbase \\
-    $\lambda$ &  0.39968 & \si{\radian} & Steer axis tilt \\
-    $g$ &  9.81 & \si{\meter\per\second\squared} & Acceleration due to gravity \\
-    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
-    Rear Wheel [R] & & \\
-    \midrule
-    $m_R$ &  3.11 & \si{\kilogram} & mass \\
-    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
-    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
-    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
-    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
-    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
-    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
-    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
-    Front Wheel [F] & & \\
-    \midrule
-    $m_F$ &  2.02 & \si{\kilogram} & Mass \\
-    $r_F$ &  0.34353 & \si{\meter} & Radius \\
-    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
-    $y_F$ & 0.0 & \si{\meter} & Mass center \\
-    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
-    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
-    \midrule
-    $l_P$ &  1.7280 & \si{\meter} & Body length \\
-    $w_P$ &  0.48300 & \si{\meter} & Body width \\
-    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
-    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
-    $y_P$ & 0.0 & \si{\meter} & Mass center \\
-    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
-    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
-    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
-    \midrule
-    $m_H$ & 3.2200 & \si{\kilogram} & Mass \\
-    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
-    $y_H$ & 0.0 & \si{\meter} & Mass center \\
-    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
-    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
-    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
-    \midrule
-    $m_D$ &  9.8600 & \si{\kilogram} & Mass \\
-    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
-    $y_D$ & 0 & \si{\meter} & Mass center \\
-    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
-    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
-    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-These parameters can be transformed into the benchmark parameterization
-r adily, but not vice versa.
-
-We search for the optimal values of a subset of 23 of the principal parameters.
-
-\begin{align}
-  \mathbf{p} =
-    [w \quad
-     c \quad
-     \lambda \quad
-     m_D \quad
-     x_D \quad
-     z_D \quad
-     k_{Daa} \quad
-     k_{Dbb} \quad
-     \alpha_D \quad
-     x_P \quad
-     z_P \quad
-     \alpha_P \quad
-     m_H \quad
-     x_H \quad
-     z_H \quad
-     k_{Haa} \quad
-     k_{Hbb} \quad
-     k_{Hyy} \quad
-     \alpha_H \quad
-     r_R \quad
-     m_R \quad
-     r_F \quad
-     m_F]
-\end{align}
-
-\subsection*{Constraints}
-%
-The optimal principal parameters are subject to a set of constraints designed
-to ensure a physically realizable bicycle is obtained.
 
@@
 
+\section*{Rigid Body Definitions}
+
 \subsection*{Rear Frame [D]}
-
+%
 There are five free parameters associated with the rear frame's inertial
@@
 
-with these additional parameters that have fixed values:
-
-\begin{align}
-  I_{Dxy} = I_{Dyz} = 0 \\
-  y_D = 0 \\
-  % 
-  l_P = ? \si{m}\\
-  w_P = ? \si{m} \\
-  m_P = 83.5 \si{kg} \\
-  y_P = 0 \si{\meter} \\
-  I_{Pxx} = 10.98524 \si{\kg\meter^2} \\
-  I_{Pyy} = 11.2571 \si{\kg\meter^2}\\
-  I_{Pzz} = 2.26922 \si{\kg\meter^2}\\
-  I_{Pxy} = I_{Pyz} = 0 \si{\kg\meter^2}\\
-  I_{Pxz} = -1.70211 \si{\kg\meter^2}\\
-  % 
-  I_{Hxy} = I_{Hyz} = 0 \\
-  y_H = 0 \\
-  %
-  I_{Rxy} = I_{Ryz} = I_{Rxz} = 0 \\
-  %
-  I_{Fxy} = I_{Fyz} = I_{Fxz} = 0 \\
-\end{align}
-
-with these conversions from our parameters to the benchmark parameters:
-
-\begin{align}
-  I_{Dxx} = m_D k_{Dxx}^2 \\
-  I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2} \\
-  I_{Dzz} = m_D k_{Dzz}^2 \\
-  I_{Dxz} = f(I_{Dxx}, I_{Dzz}, \alpha_D)  \\
-  %
-  m_B = m_D + m_P \\
-  x_B = \frac{m_D x_D + m_P x_P}{m_B} \\
-  z_B = \frac{m_D z_D + m_P z_P}{m_B} \\
-  I_{Bxx} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
-  I_{Byy} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
-  I_{Bzz} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
-  I_{Bxy} = 0 \\
-  I_{Byz} = 0 \\
-  I_{Bxz} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
-  % 
-  I_{Hxx} = m_H k_{Hxx}^2 \\
-  I_{Hyy} = m_H k_{Hyy}^2\\
-  I_{Hzz} = m_H k_{Hzz}^2 \\
-  I_{Hxz} = f(I_{Hxx}, I_{Hzz}, \alpha_H)  \\
-  %
-  I_{Ryy} = m_R r_R^2 \\
-  I_{Rxx} = I_{Ryy} / 2 \\
-  I_{Rzz} = I_{Ryy} / 2 \\
-  %
-  I_{Fyy} = m_F r_F^2 \\
-  I_{Fxx} = I_{Fyy} / 2 \\
-  I_{Fzz} = I_{Fyy} / 2 \\
-  %
-  m_T = m_H + m_B + m_F + m_R \\
-  x_T = \frac{m_H x_H + m_B x_B + m_F w}{m_T} \\
-  z_T = \frac{m_H z_H + m_B z_B + m_F r_F + m_R r_R}{m_T} 
-\end{align}
-
-with these parameter bounds:
-
-\begin{align}
-  m_R > 0 \\
-  m_F > 0 \\
-  m_H > 0 \\
-  m_D > 0 \\
-  r_R > r_{min} \\
-  r_F > r_{min} \\
-  z_H < 0 \\
-  z_D < 0 \\
-  -\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2} \\
-  -\frac{\pi}{2} \leq \alpha_P \leq \frac{\pi}{2} \\
-  -\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2} \\
-  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
-\end{align}
-
-and these constraints:
-
-\begin{align}
-  I_{Hyy} \leq \sqrt{I_{Hxx}^2 + I_{Hzz}^2} \\
-  z_P + l_P / 2 \cos{\alpha_P} & \leq 0 \\
-  z_P + w_P / 2 \sin{\alpha_P} & \leq 0 \\
-  z_P - l_P / 2 \cos{\alpha_P} & \leq 0 \\
-  z_P - w_P / 2 \sin{\alpha_P} & \leq 0 \\
-  w > r_F + r_R \\
-  m_{Tmin} + m_P \leq m_T \leq 25 \si{kg} + m_P \\
-  \frac{x_T}{|z_T|} < \frac{1}{4} \\
-  \frac{x_T - w}{|z_T|} < -1 \\
-  real(\mathbf{s}) < 0
-\end{align}
-
 \bibliographystyle{acm}

### 2019-08-21 paper/paper.tex
Simplified constraints section.

@@
 % - Problem Definition
+%   - Parameterization
+%   - Bounds and Constraints
+%   - Free Parameters
 %   - Optimization
-%   - Parameterization
-%   - Bounds
-%   - Constraints
 % - Solutions
@@
 
-\section*{Introduction}
+\section{Introduction}
 %
@@
 
-\section*{Problem Definition}
+\section{Problem Definition}
 %
@@
 
-\subsection*{Optimization}
-%
-The constraint statements above translate into upper and lower bounds on 12 of
-the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
-constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
-expressed in the standard non-linear programming form.
-%
-\begin{equation}
-  \begin{aligned}
-    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
-    & \text{subject to} & & \\
-    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
-    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
-  \end{aligned}
-\end{equation}
-
-\subsection*{Bicycle Model Parameters}
+\subsection{Bicycle Model Parameters}
 %
@@
 benchmark's four rigid bodies.
-
-The variables are as follows:
 %
-\begin{itemize}
-  \item wheelbase, trail, steer axis tilt, wheel radii
-  \item somersault angle of the person (same direction as bicycle pitch)
-  \item masses of the front frame, rear frame, front wheel, and rear wheel
-  \item mass center locations of the front frame, rear frame, and rider
-  \item principal central radii of gyration of front and rear frames
-  \item principal inertia axis pitch angles of the front frame, rear frame, and rider
-\end{itemize}
-
 \begin{table}
-  \caption{Parameters and their default values.}
+  \caption{Full set of 47 parameters and their default values derived from the
+  measurements in~\cite{Moore2012}.}
+  \label{tab:parameters}
   \begin{tabular}{llll}
@@
 
-$\mathbf{p}\in\mathbb{R}^{24}$These parameters can be transformed into the benchmark parameterization
-r adily, but not vice versa.
-We search for the optimal values of a subset of 23 of the principal parameters.
-
-\begin{align}
-  \mathbf{p} = &
-    [w \quad
-     c \quad
-     \lambda \quad
-     m_D \quad
-     x_D \quad
-     z_D \quad
-     k_{Daa} \quad
-     k_{Dbb} \quad
-     \alpha_D \quad
-     x_P \quad
-     z_P \quad
-     \alpha_P \quad \\
- &    m_H \quad
-     x_H \quad
-     z_H \quad
-     k_{Haa} \quad
-     k_{Hbb} \quad
-     k_{Hyy} \quad
-     \alpha_H \quad
-     r_R \quad
-     m_R \quad
-     r_F \quad
-     m_F]
-\end{align}
-
-\subsection*{Bounds and Constraints}
+\subsection{Bounds and Constraints}
 %
 The optimal principal parameters are subject to a set of constraints designed
-to ensure a physically realizable bicycle is obtained.
-We formulate bounds on the variables as well as equality and inequality
-constraints among the variables to ensure a physically realizable bicycle.
-Below the constraints are defined and grouped by the associated rigid body or
-collection thereof:
+to ensure a physically realizable bicycle is obtained. We formulate bounds on
+the variables as well as equality and inequality constraints among the
+variables to ensure a physically realizable bicycle. Below the constraints are
+defined and grouped by the associated rigid body or collection thereof:
 %
@@
 
-\section*{Rigid Body Definitions}
-
 \subsection*{Rear Frame [D]}
 %
+Several constraints are set for the rear frame. We constrain the rear frame to
+be planar, $g_1$, and symmetric with respect to the $XZ$ plane. We prevent the
+rear frame from penetrating the ground by limiting the inertial spread with
+respect to its mass center, $c_{12}$, but also set a minimum inertial spread to
+ensure a frame can span from the rear wheel to the mass center of the rear
+frame, $c_9$. The spread factor in $c_{12}$ of 1.4 is based on the ratio of
+geometrical spread of a typical bicycle frame and its radius of gyration.
+Several parameters are bounded. We require the rear frame mass to be positive,
+the center of mass not penetrate the ground, and we allow for any angular
+orientation of the $XZ$ principal directions but limit the angle to
+$-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$ as angle beyond that are
+redundant.
+
+\subsection{Person [P]}
+
+We assume that the person's joint configurations are such that they are in a
+nominal configuration for pedaling, i.e. an average normal everyday riding
+position on a typical bicycle, i.e. they stayed in the same configuration as
+they were seated on the Batavus Browser bicycle. The person is assumed to be
+symmetric about the $XZ$ plane. We allow the rider to be rotated about the $Y$
+axis and positioned anywhere within the plane of symmetry above the ground.
+
+To prevent the rider from being positioned and oriented such that their body is
+not penetrating the the ground we introduce two dimensions that define a cross
+whose apex is at the center of mass of the person and the cross axes are
+parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
+along the principal axis to the tip of the toes and $w_P$ is the distance along
+the second principal axes to the tip of the hands. Constraints
+$c_2,\ldots,c_5$.
+
+\subsection{Front Frame}
+
+The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. We
+allow for any angular orientation of the $XZ$ principal directions but limit
+the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.  We prevent the
+rear frame from penetrating the ground by limiting the inertial spread with
+respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
+ensure a frame can span from the rear wheel to the mass center of the rear
+frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
+geometrical spread of a typical bicycle frame and its radius of gyration. The
+front frame is not planar but is narrow with respect to the $XZ$ plane, which
+is enforced by constraint $c_1$.
+
+\subsection{Front [F] and Rear [R] Wheels}
+
+We enforce the assumption that both wheels have moments of inertia of that of a
+simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than a
+minimal size based on small purchable spoked wheel with tire.
+
+\subsection{Total Bike [T]}
+
+The trail and wheelbase can take on any real values. The steer axis tilt is
+limted to 180 degrees. We introduce a constraint $c_{10}$ that prevents the
+wheels from physically overlapping and require that the bicycle be liftable by
+an average person, $c_{11}$. Finally, we require that the bicycle not topple
+forward during hard breaking or backward during hard acceleration.
+
+\begin{align}
+  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}
+\end{align}
+
+This translates to two constraints, $c_6,c_7$ that bound the total center of
+mass $(x_T,z_T)$ in a triangle in the $XZ$ plane. Lastly, we constrain the
+eight closed loop eigenvalues associated with the controller in
+~\cite{Hess2012} to be stable, i.e. have negative real parts.  These are
+expressed in constraints $c_{14},\ldots,c_{21}$. Closed loop stability is
+required for the HQM to provide a meaningful result.
+
+\subsection{Free parameters}
+%
+These constraints leave 23 free parameters for the optimizer to These
+parameters can be transformed into the benchmark parameterization r adily, but
+not vice versa.
+We search for the optimal values of a subset of 23 of the principal parameters.
+
+$\mathbf{p}\in\mathbb{R}^{23}$
+
 There are five free parameters associated with the rear frame's inertial
@@
 \end{description}
-
-Several constraints are set for the rear frame. We constrain the rear frame to
-be planar, $g_1$ and symmetric with respect to the $XZ$ plane, i.e. $I_{Dxy} =
-I_{Dyz} = 0$. We prevent the rear frame from penetrating the ground by limiting
-the inertial spread with respect to its mass center, $c_{12}$, but also set a
-minimum inertial spread to ensure a frame can span from the rear wheel to the
-mass center of the rear frame, $c_9$. The spread factor in $c_{12}$ of 1.4 is
-based on the ratio of geometrical spread of a typical bicycle frame and its
-radius of gyration. Several parameters are bounded. We require the rear frame
-mass to be positive, the center of mass not penetrate the ground, and we allow
-for any angular orientation of the $XZ$ principal directions but limit the
-angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$.
-
-\subsection{Person [P]}
-
-We assume that the person's joint configurations are such that they are in a
-nominal configuration for pedaling, i.e. an average normal everyday riding
-position on a typical bicycle. The person is assumed to be symmetric about the
-$XZ$ plane. We allow the rider to be rotated about the $Y$ axis and positioned
-anywhere within the plane of symmetry above the ground. Three free parameters
+ Three free parameters
 are associated with these assumptions:
@@
 
-There are four fixed parameters that enforce this assumption:
-
-\begin{description}
-  \item[$m_P$] The mass of the person which is fixed at $XX \si{kg}$.
-  \item[$k_{Pxx}$] Central radius of gyration with respect to the $X$ axis
-    where $I_{Pxx} = m_P k_{Pxx}^2$.
-  \item[$k_{Pyy}$] Central radius of gyration with respect to the $Y$ axis
-    where $I_{Pyy} = m_P k_{Pyy}^2=$.
-  \item[$k_{Pzz}$] Central radius of gyration with respect to the $Z$ axis
-    where $I_{Pzz} = m_P k_{Pzz}^2=$.
-  \item[$I_{Bxy}, I_{Byz}$] Products of inertia are zero.
-  \item[$y_P$] Is zero.
-\end{description}
-
-To prevent the rider from being positioned and oriented such that their body is
-not penetrating the the ground we introduce two dimensions that define a cross
-whose apex is at the center of mass of the person and the cross axes are
-parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
-along the principal axis to the tip of the toes and $w_P$ is the distance along
-the second principal axes to the tip of the hands. Constraints $c_2,\ldots,c_5$.
-
-\subsection{Front Frame}
-
-The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$.
 There are seven free parameters:
@@
 
-We allow for any angular orientation of the $XZ$ principal directions but limit
-the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.  We prevent the
-rear frame from penetrating the ground by limiting the inertial spread with
-respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
-ensure a frame can span from the rear wheel to the mass center of the rear
-frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
-geometrical spread of a typical bicycle frame and its radius of gyration. The
-front frame is not planar but is narrow with respect to the $XZ$ plane, which
-is enforced by constraint $c_1$.
-
-\subsection{Front [F] and Rear [R] Wheels}
-
-We enforce the assumption that both wheels have moments of inertia of that of a
-simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than
-some minimal size based on small bicycle wheel.
-
-\subsection{Total Bike [T]}
-
-\subsubsection{Parameters}
-
 There are three remaing parameters associated with the fundamental geometry of
@@
 
-\subsubsection{Bounds}
-
-The trail can take on any real value as well as the steer axis tilt except the
-bound the angle to a minimal range:
-
 \begin{align}
-  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
+  \begin{split}
+    \mathbf{p} = [ &
+       w        \quad
+       c        \quad
+       \lambda  \quad
+       m_D      \quad
+       x_D      \quad
+       z_D      \quad
+       k_{Daa}  \quad
+       k_{Dbb}  \quad
+       \alpha_D \quad
+       x_P      \quad
+       z_P      \quad
+       \alpha_P \\
+     & m_H      \quad
+       x_H      \quad
+       z_H      \quad
+       k_{Haa}  \quad
+       k_{Hbb}  \quad
+       k_{Hyy}  \quad
+       \alpha_H \quad
+       r_R      \quad
+       m_R      \quad
+       r_F      \quad
+       m_F]
+  \end{split}
 \end{align}
 
-\subsubsection{Constraints}
+There are four fixed parameters that enforce this assumption:
 
-We introduce a constraint $c_{10}$ that prevents the wheels from physically
-overlapping and require that the bicycle be liftable by an average person,
-$c_{11}$. Finally, we require that the bicycle not topple forward during hard
-breaking or backward during hard acceleration.
-
-\begin{align}
-  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}
-\end{align}
-
-This translates to two constraints, $c_6,c_7$ that bound the total center of
-mass $(x_T,z_T)$ in a triangle in the $XZ$ plane.
-
-Lastly, we constrain the eight closed loop eigenvalues associated with the
-controller in ~\cite{Hess2012} to be stable, i.e. have negative real parts.
-These are expressed in constraints $c_{14},\ldots,c_{21}$. Closed loop
-stability is required for the HQM to provide a meaningful result.
-
+\begin{description}
+  \item[$m_P$] The mass of the person which is fixed at $XX \si{kg}$.
+  \item[$k_{Pxx}$] Central radius of gyration with respect to the $X$ axis
+    where $I_{Pxx} = m_P k_{Pxx}^2$.
+  \item[$k_{Pyy}$] Central radius of gyration with respect to the $Y$ axis
+    where $I_{Pyy} = m_P k_{Pyy}^2=$.
+  \item[$k_{Pzz}$] Central radius of gyration with respect to the $Z$ axis
+    where $I_{Pzz} = m_P k_{Pzz}^2=$.
+  \item[$I_{Bxy}, I_{Byz}$] Products of inertia are zero.
+  \item[$y_P$] Is zero.
+\end{description}
 \subsection{Summary}
@@
 
+\subsection*{Optimization}
+%
+The constraint statements above translate into upper and lower bounds on 12 of
+the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
+constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
+expressed in the standard non-linear programming form.
+%
+\begin{equation}
+  \begin{aligned}
+    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
+    & \text{subject to} & & \\
+    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
+    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
+  \end{aligned}
+\end{equation}
+
 \section*{Solving}

### 2019-08-21 paper/paper.tex
More formatting and start of results text.

@@
 \begin{flushleft}
-  {\fontsize{12}{14}{Jason K. Moore, Mont Hubbard, and Ronald A. Hess}\\}
+  {\fontsize{12}{14}{Jason K. Moore and Mont Hubbard}\\}
   \textit{Mechanical and Aerospace Engineering\\
@@
           One Shields Avenue, Davis, CA, USA 95817\\
-          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu, rahess@ucdavis.edu}
+          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu}
 \end{flushleft}
@@
 % - Introduction
-% - Problem Definition
 % - Parameterization
@@
 %
-Physical design features of ground vehicles can affect their lateral handling
-qualities. Geometry, mass, and mass distribution of the vehicle's primary
-components as well as tire characteristics are primary contributors to poor and
-good handling due to their important influence on the vehicle's dynamics. In
-past work, we have presented a theoretical and computational framework for
-assessing the lateral task-independent handling qualities of simplified single
-track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
-that minimizing our handling quality metric (HQM) can produce theoretically
-optimal handling designs when only four geometric parameters are explored as
-the optimization variables~\cite{Moore2016}. The present work's goal is to
-expand this optimization problem to 24 geometry, mass, and inertial parameters
-as the optimization variables so that a broader search space of optimal bicycle
-designs is considered. This task is complicated by the fact that we desire to
-limit the search space to physically realizable designs. Doing so introduces 35
-linear and non-linear constraint equations on the optimization variables. We
-formulate a constrained optimization problem and use nonlinear programming to
-discover optimal, yet realizable, bicycle designs.
-
-\section{Introduction}
-%
 We introduced a method of optimizing four geometric parameters of a bicycle's
@@
 
-\section{Problem Definition}
+\section{Introduction}
 %
-We search for 24 optimization variables that minimize the Handling Quality
-Metric defined in~\cite{Hess2012}.
+Physical design features of ground vehicles can affect their lateral handling
+qualities. Geometry, mass, and mass distribution of the vehicle's primary
+components as well as tire characteristics are primary contributors to poor and
+good handling due to their important influence on the vehicle's dynamics. In
+past work, we have presented a theoretical and computational framework for
+assessing the lateral task-independent handling qualities of simplified single
+track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
+that minimizing our proposed handling quality metric (HQM) can produce theoretically
+optimal handling designs when only four geometric parameters are explored as
+the optimization variables~\cite{Moore2016}. The present work's goal is to
+expand this optimization problem to all of the geometry, mass, and inertial
+parameters present in the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}.
+This broadens the search space considerably but we constrain it so that only
+realizable optimal bicycle designs are discovered. To do so, we formulate a
+constrained optimization problem and use derivative-free optimization to
+discover optimal, yet realizable, bicycle designs.
 
-\subsection{Bicycle Model Parameters}
+\section{Bicycle Model Parameterization}
 %
-Our problem formulation relies on a new parameterization that reflects both a
-reformulation of and addition to the benchmark parameterization of the linear
-Whipple-Carvallo bicycle model~\cite{Meijaard2007}. This parameterization
-differs from the benchmark parameterization in that the person and rear frame
-are treated as separate but fixed rigid bodies each with their on inertial
-parameters and that the inertial parameters of all bodies are expressed in
-terms of central principal radii of gyration so that the mass is decoupled from
-the inertia. We also introduce two simple dimensional parameters that define
-the geometric extents of the person. Table~\ref{tab:parameters} provides the
-parameters and the initial values which are derived from the measurements of a
-Batuvus Browser Bicycle and the rider, Jason, presented in~\cite{Moore2012}.
-This allows the position and orientation of the rider to be specified
-independtly of the rear frame but does not introduce any more than the
-benchmark's four rigid bodies.
+Our problem formulation relies on a new bicycle model parameterization that
+reflects both a reformulation of and addition to the benchmark parameterization
+of the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}. We call this
+the ``principal parameterizatoin'' as opposed to the ``benchmark
+parameterization''. This parameterization differs from the benchmark
+parameterization in three ways.  Firstly, the person and rear frame are treated
+as separate rigid bodies each with their on inertial parameters. Secondly, we
+express the inertial parameters of each rigid body in terms of central
+principal radii of gyration to decouple the mass from the inertia terms.
+Lastly, we introduce two simple dimensional parameters that define the
+geometric extents of the person which are used to constrain the location of the
+person's body. Table~\ref{tab:parameters} provides the parameters and the
+reference values which are derived from the measurements of a Batavus Browser
+Bicycle and the rider Jason presented in~\cite{Moore2012}. This
+parameterization can be transformed into the benchmark parameterization
+readily, but not vice versa.
 %
 \begin{table}
-  \caption{Full set of 47 parameters and their default values derived from the
-  measurements in~\cite{Moore2012}.}
+  \caption{Full set of 47 principal parameters and their default values derived
+  from the measurements in~\cite{Moore2012}.}
   \label{tab:parameters}
+  \centering
   \begin{tabular}{llll}
@@
 
-\subsection{Bounds and Constraints}
+\section{Bounds and Constraints}
 %
 The optimal principal parameters are subject to a set of constraints designed
-to ensure a physically realizable bicycle is obtained. We formulate bounds on
-the variables as well as equality and inequality constraints among the
-variables to ensure a physically realizable bicycle. Below the constraints are
-defined and grouped by the associated rigid body or collection thereof:
+to ensure that a physically realizable bicycle is obtained form the opimization
+procedure. These constraints are made up of bounds on the free parameters and
+both equality and inequality constraints among the parameters. Below the basic
+constraint concepts presented and grouped by the associated rigid body or
+collection thereof:
 %
 \begin{description}
-  \item[Vehicle] The resulting combination of the five rigid bodies.
+  \item[Total] The resulting combination of the five rigid bodies.
     \begin{itemize}
-      \itemsep0em
-      \item The physical extents of the bicycle and rider must exist above the
+      \item The physical extents of the rigid bodies must exist above the
         ground plane.
-      \item Both bicycle and rider are symmetric about the sagittal
+      \item Both bicycle and rider are symmetric about the rider's sagittal
         plane.
@@
         amount.
-      \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
-        radius.
+      \item The wheels cannot overlap.
       \item The bicycle cannot topple forward during hard braking or backward
-        during hard acceleration, i.e. -1 g < acceleration < 1 g.
+        during hard acceleration.
     \end{itemize}
@@
     \begin{itemize}
-      \itemsep0em
-      \item Wheel radius must be greater than a minimum value.
-      \item Wheels are inertially wheel-like, i.e.  $I_{rot}=mr^2,
-        I_{rad}=I_{rot}/2$.
+      \item Wheel radius and mass must be greater than a minimum value.
+      \item Wheels are inertially wheel-like, i.e. symmetric about each plane
+        and most of the mass is at the rim.
     \end{itemize}
 
-  \item[Frames] Both front and rear frames have identical constraints.
+  \item[Frames] Front frame (handlebar + fork) and rear frame
     \begin{itemize}
-      \itemsep0em
-      \item A pair of principal directions lie in the roll/yaw
-        plane to maintain symmetry and one of these axes is within $\pm \pi/4$
-        rad of the ground plane.
       \item The mass and inertia of the frames are positive and large enough to
@@
         density.
-      \item The frames are planar in nature, i.e.
-        $k_{lateral}^2=k_{longitudinal}^2+k_{vertical}^2$, where $k$ is the
-        radius of gyration.
+      \item The rear frame is planar in nature and the front frame's moments of
+        inertia are consitently dependent.
     \end{itemize}
@@
     \begin{itemize}
-      \itemsep0em
       \item Rider mass is that of an average person.
@@
         derived from standard body segment estimation methods.
-      \item One rider principal axis is parallel with the bicycle pitch axis to
-        enforce symmetry.
     \end{itemize}
@@
 
+These bounds, equality, and inequality constraints are presented mathematically
+in tables \ref{tab:bounds}, \ref{tab:equality-constraints},
+\ref{tab:inequality-constraints} respectively and explained in more detail in
+the following sections.
+%
 \begin{table}
-  \caption{Inequality constraints}
-  \begin{tabular}{lll}
-    \toprule
-    Constraint & Equation & Description \\
-    \midrule
-    $c_1$    & $\sqrt{I_{Hxx}^2+I_{Hzz}^2} \geq I_{Hyy}$           & Narrow front frame. \\
-    $c_2$    & $0 \geq z_P + l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
-    $c_3$    & $0 \geq z_P + w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
-    $c_4$    & $0 \geq z_P - l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
-    $c_5$    & $0 \geq z_P - w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
-    $c_6$    & $x_T \geq |z_T|/4$                                  & Maximum acceleration of $1/4g$. \\
-    $c_7$    & $w-x_T \geq 3/4|z_T|$                               & Maximum deceleration of $3/4g$. \\
-    $c_8$    & $2k_{Hyy} \geq \sqrt{(x_H - w)^2 + (z_H + r_F)^2}$  & Minimal inertial spread. \\
-    $c_9$    & $2k_{Dyy} \geq \sqrt{(x_D - 0)^2 + (z_D + r_R)^2}$  & Minimal inertial spread. \\
-    $c_{10}$ & $w \geq r_F +r_R$                                   & Non-overlapping wheels. \\
-    $c_{11}$ & $25\si{\kg} \geq m_D + m_H + m_R + m_F$             & Maximum bicycle mass. \\
-    $c_{12}$ & $-z_D \geq 1.4 k_{Dyy}$                             & Rear frame cannot penetrate ground. \\
-    $c_{13}$ & $-z_H \geq 1.4 k_{Hyy}$                             & Front frame cannot penetrate ground. \\
-    $c_{14,\ldots,21}$ & $0 \geq s_1,\ldots,s_8$                   & Closed loop stability. \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{Equality constraints}
-  \begin{tabular}{lll}
-    \toprule
-    Constraint & Equation & Description \\
-    \midrule
-    $g_1$ & $I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2}$ & Rear frame is planar. \\
-    $g_2$ & $k_{Ryy} = r_R$  & Rear wheel is a ring \\
-    $g_3$ & $k_{Raa} = k_{Ryy}/2$ & Rear wheel is a ring \\
-    $g_4$ & $k_{Rbb} = k_{Ryy}/2$ & Rear wheel is a ring \\
-    $g_5$ & $k_{Fyy} = r_F$  & Front wheel is a ring \\
-    $g_6$ & $k_{Faa} = k_{Fyy}/2$ & Front wheel is a ring \\
-    $g_7$ & $k_{Fbb} = k_{Fyy}/2$ & Front wheel is a ring \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\begin{table}
-  \caption{Parameter Bounds}
+  \caption{Free parameter upper and lower bounds}
+  \label{tab:bounds}
+  \centering
   \begin{tabular}{ccccc}
@@
 \end{table}
+%
+\begin{table}
+  \caption{Equality constraints}
+  \label{tab:equality-constraints}
+  \centering
+  \begin{tabular}{lll}
+    \toprule
+    Constraint & Equation & Description \\
+    \midrule
+    $g_1$ & $I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2}$ & Rear frame is planar. \\
+    $g_2$ & $k_{Ryy} = r_R$  & Rear wheel is a ring \\
+    $g_3$ & $k_{Raa} = k_{Ryy}/2$ & Rear wheel is a ring \\
+    $g_4$ & $k_{Rbb} = k_{Ryy}/2$ & Rear wheel is a ring \\
+    $g_5$ & $k_{Fyy} = r_F$  & Front wheel is a ring \\
+    $g_6$ & $k_{Faa} = k_{Fyy}/2$ & Front wheel is a ring \\
+    $g_7$ & $k_{Fbb} = k_{Fyy}/2$ & Front wheel is a ring \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+%
+\begin{table}
+  \caption{Inequality constraints}
+  \label{tab:inequality-constraints}
+  \centering
+  \begin{tabular}{lll}
+    \toprule
+    Constraint & Equation & Description \\
+    \midrule
+    $c_1$    & $\sqrt{I_{Hxx}^2+I_{Hzz}^2} \geq I_{Hyy}$           & Consistent moments of inertia. \\
+    $c_2$    & $0 \geq z_P + l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_3$    & $0 \geq z_P + w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_4$    & $0 \geq z_P - l_P/2 \cos{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_5$    & $0 \geq z_P - w_P/2 \sin{\alpha_P}$                 & Person cannot penetrate ground. \\
+    $c_6$    & $x_T \geq |z_T|/4$                                  & Maximum acceleration of $1/4g$. \\
+    $c_7$    & $w-x_T \geq 3/4|z_T|$                               & Maximum deceleration of $3/4g$. \\
+    $c_8$    & $2k_{Hyy} \geq \sqrt{(x_H - w)^2 + (z_H + r_F)^2}$  & Minimal inertial spread. \\
+    $c_9$    & $2k_{Dyy} \geq \sqrt{(x_D - 0)^2 + (z_D + r_R)^2}$  & Minimal inertial spread. \\
+    $c_{10}$ & $w \geq r_F +r_R$                                   & Non-overlapping wheels. \\
+    $c_{11}$ & $25\si{\kg} \geq m_D + m_H + m_R + m_F$             & Maximum bicycle mass. \\
+    $c_{12}$ & $-z_D \geq 1.4 k_{Dyy}$                             & Rear frame cannot penetrate ground. \\
+    $c_{13}$ & $-z_H \geq 1.4 k_{Hyy}$                             & Front frame cannot penetrate ground. \\
+    $c_{14,\ldots,21}$ & $0 \geq s_1,\ldots,s_8$                   & Closed loop stability. \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
-\subsection*{Rear Frame [D]}
+\subsubsection*{Rear Frame [D]}
 %
@@
 \subsection{Front Frame}
-
+%
 The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. We
@@
 \subsection{Front [F] and Rear [R] Wheels}
-
+%
 We enforce the assumption that both wheels have moments of inertia of that of a
@@
 \subsection{Total Bike [T]}
-
+%
 The trail and wheelbase can take on any real values. The steer axis tilt is
@@
 
-\subsection{Free parameters}
+\section{Free parameters}
 %
-These constraints leave 23 free parameters for the optimizer to These
-parameters can be transformed into the benchmark parameterization r adily, but
-not vice versa.
-We search for the optimal values of a subset of 23 of the principal parameters.
-
-$\mathbf{p}\in\mathbb{R}^{23}$
-
-There are five free parameters associated with the rear frame's inertial
-characteristics:
-
-\begin{description}
-  \item[$m_D$] Mass of the rear frame.
-  \item[$x_D$] Location along the $X$ axis of the rear frame center of mass.
-  \item[$z_D$] Location along the $Z$ axis of the rear frame center of mass.
-  \item[$k_{Daa}$] Maximum principal central radius of gyration in the $XZ$
-    plane.
-  \item[$k_{Dbb}$] Minimum principal central radius of gyration in the $XZ$
-    plane.
-  \item[$\alpha_D$] Angle between the $X$ axis and the maximum principal axis
-    that lies in the $XZ$ plane.
-\end{description}
- Three free parameters
-are associated with these assumptions:
-
-\begin{description}
-  \item[$x_P$] Location along the $X$ axis of the person's center of mass.
-  \item[$z_P$] Location along the $Z$ axis of the person's center of mass.
-  \item[$\alpha_P$] Angle between the $X$ axis and the primary principal
-    inertia axis of the person.
-\end{description}
-
-There are seven free parameters:
-
-\begin{description}
-  \item[$m_H$] The mass of the rear frame.
-  \item[$x_B$] Location along the $X$ axis of the rear frame's center of mass.
-  \item[$z_H$] Location along the $Z$ axis of the rear frame's center of mass.
-  \item[$k_{Haa}$] Central radius of gyration with respect to the $X$ axis
-    where $k_{Hxx}^2=I_{Hxx} / m_H$.
-  \item[$k_{Hyy}$] Central radius of gyration with respect to the $X$ axis
-    where $k_{Hyy}^2=I_{Hyy} / m_H$.
-  \item[$k_{Hbb}$] Central radius of gyration with respect to the $X$ axis
-    where $k_{Hzz}^2=I_{Hzz} / m_H$.
-  \item[$\alpha_H$] Angle between the $X$ axis and the primary principal
-    inertia axis.
-\end{description}
-
-There are three remaing parameters associated with the fundamental geometry of
-the bicycle: wheelbase $w$, trail $c$, and steer axis tilt $\lambda$ all the
-same as defined in~\cite{Meijaard2007}.
-
+The above constraints leaves 23 of the 47 parameters free for optimizing which
+we collect in the vector $\mathbf{p}\in\mathbb{R}^{23}$ and define as:
+%
 \begin{align}
@@
 
-There are four fixed parameters that enforce this assumption:
+\section{Optimization}
+%
+Our objective in the optimization is to minimize the peak HQM value subject to
+the bounds, $\mathbf{p}^L,\mathbf{p}^U$, and the constraints
+$\mathbf{g}(\mathbf{p}),\mathbf{c}(\mathbf{p})$ . Given a set of bicycle model
+parameter values we generate a bandwidth limited human-like controller using
+the methods in ~\cite{Moore2012}. Once the closed loop stable controller is
+constructed, the HQM can be computed as per the definition in \cite{Hess2012}
+and the scaler peak value returned as the objective $J$. This problem is
+presented as a non-linear programming problem in the following equation.
+%
+\begin{equation}
+  \begin{aligned}
+    & \underset{\mathbf{p}}{\text{minimize}} & & J(\mathbf{p})=max(\textrm{HQM}(\mathbf{p})) \\
+    & \text{subject to} & & \\
+    & & & \mathbf{g}(\mathbf{p}) \leq \mathbf{0} \\
+    & & & \mathbf{c}(\mathbf{p}) = \mathbf{0} \\
+    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U
+  \end{aligned}
+\end{equation}
 
-\begin{description}
-  \item[$m_P$] The mass of the person which is fixed at $XX \si{kg}$.
-  \item[$k_{Pxx}$] Central radius of gyration with respect to the $X$ axis
-    where $I_{Pxx} = m_P k_{Pxx}^2$.
-  \item[$k_{Pyy}$] Central radius of gyration with respect to the $Y$ axis
-    where $I_{Pyy} = m_P k_{Pyy}^2=$.
-  \item[$k_{Pzz}$] Central radius of gyration with respect to the $Z$ axis
-    where $I_{Pzz} = m_P k_{Pzz}^2=$.
-  \item[$I_{Bxy}, I_{Byz}$] Products of inertia are zero.
-  \item[$y_P$] Is zero.
-\end{description}
-\subsection{Summary}
+We make use of the derivative-free optimizer CMA-ES~\cite{Hansen1996} to find
+solutions to this problem. The optimization supports parameter bounds and
+equality constraints but does not support inequality constraints. To get around
+this limitation we move the inequality constraints into the objective function
+an penalize the objective if the constraints are violated with the following
+rules:
+%
+\begin{align}
+  J(\mathbf{p}) =
+  \begin{cases}
+    max(\textrm{HQM}(\mathbf{p})) & \textrm{if} \quad any(\mathbf{g}(\mathbf{p})) < 0 \\
+    30 + ||\mathbf{g}_{+}(\mathbf{p})||/10 & \textrm{if} \quad
+      any(\mathbf{g}(\mathbf{p})) \geq 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| < 30 \\
+    ||\mathbf{g}_{+}(\mathbf{p})|| & \textrm{if} \quad
+      any(\mathbf{g}(\mathbf{p})) \geq 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| \geq 30
+  \end{cases}
+\end{align}
+where $||\mathbf{g}_{+}||$ is the norm of the positive elements of
+$\mathbf{g}$.
 
-The above definitions represent a constrained optimization in 23 free
-parameters with 17 constraints. This bounds the space of bicycle designs to be
-physically realizable and excludes bicycles that don't have the inherent
-characteristics of a bicycle.
+This creates a discontinous objective function but in practice the CMS-ES
+algorthim is able to move into the parameter space where all the constraints
+are satisfied and find a (local) minima. For our purposes, this sufficiently
+finds parameter values that produce an optimally handling design.
 
-\begin{align}
-  \mathbf{p} =
-  \begin{bmatrix}
-    % primary geometry
-    w \\
-    c \\
-    \lambda
-    % front frame
-    m_D \\
-    x_D \\
-    z_D \\
-    k_{Dxx} \\
-    k_{Dzz} \\
-    \alpha_D \\
-    % person
-    x_P \\
-    z_P \\
-    \alpha_P \\
-    % front frame
-    m_H \\
-    x_H \\
-    z_H \\
-    k_{Hxx} \\
-    k_{Hyy} \\
-    k_{Hzz} \\
-    I_{Hxz} \\
-    \alpha_H \\
-    % rear wheel
-    r_R \\
-    m_R \\
-    % front wheel
-    r_F \\
-    m_F \\
-  \end{bmatrix}
-\end{align}
+\section{Results}
+%
+We discover four bicycles for four different design speeds (3, 5, 7, and 9 m/s)
+that have an optimally low HQM, see Table~\ref{tab:hqm} and satisify all
+constraints and parameter boundaries. We belive these bicycles to be physically
+realizable with minor differences. The most striking feature is that all of the
+bicycle are larger that the reference bicycle in some way.  The 3 m/s bicycle
+places both the rider's mass center and the rear frame's mass center close to 3
+meters above the ground plane. The 5 and 7 m/s bicycles have wheelbase values
+that are simlar to the reference bicycle, but 3 and 9 m/s have very large wheel
+bases. The 9 m/s bicycle is very large overall. Only the 5 m/s bicycle is of a
+scale close to a typical bicycle. It is noteable that the 9 m/s bicycle has
+signficant negative trail.
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=6in]{../figures/optimal-geometries.png}
+  \caption{Depcitions of the bicycle geometry and geometric representations of
+    the inertial quanties for the reference bicycle and four optimal solutions
+    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
+    wheel (oragne), rear wheel (purple), rear frame (blue), front frame
+    (green), and person (red). The solid black lines represent the essential
+    bicycle geometry. The dotted black line represents the steer axis. The
+    solid colored curves represent the contours of solid ellipsoids with
+    equivalent inertia as the principal inertia of the associated rigid body.
+    The dotted colored lines represent the extents of the centerial radii of
+    gyration of each rigid body.}
+\end{figure}
+%
+\begin{table}
+  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
+    at each speed.}
+  \label{tab:hqm}
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
+    \midrule
+    3 & 13.0753 & 2.0118 & 85\% \\
+    5 & 4.5213  & 0.0115 & 100\% \\
+    7 & 3.0434  & 0.0220 &  99\% \\
+    9 & 2.3377  & 0.8386 &  64\% \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\section{Discussion and Conclusion}
+%
 
@@
 
-\subsection*{Optimization}
-%
-The constraint statements above translate into upper and lower bounds on 12 of
-the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
-constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
-expressed in the standard non-linear programming form.
-%
-\begin{equation}
-  \begin{aligned}
-    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
-    & \text{subject to} & & \\
-    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
-    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
-  \end{aligned}
-\end{equation}
-
-\section*{Solving}
-\section*{Results}
-\section*{Discussion and Conclusion}
-
 \end{document}

### 2019-08-21 src/geometry_plot.py
Stacked the geometry plots.

@@ -5,16 +5,15 @@ from bicycleparameters.parameter_sets import PrincipalParameterSet
 
 speeds = [3, 5, 7, 9]
 
-fig, axes = plt.subplots(nrows=3, ncols=len(speeds)+1, sharey=True,
-                         gridspec_kw={'width_ratios': [1, 2, 1.5, 1.5, 4]})
+fig, axes = plt.subplots(nrows=3, ncols=2)
 
 with open('data/principal-browserjason.yml', 'r') as f:
     par_dict = yaml.load(f)['values']
     par_set = PrincipalParameterSet(par_dict)
-    par_set.plot_all(ax=axes[0])
-    axes[0].set_title('Reference')
+    par_set.plot_all(ax=axes[0, 0])
+    axes[0, 0].set_title('Reference')
 
-for ax, speed in zip(axes[1:], speeds):
+for ax, speed in zip(axes.flatten()[2:], speeds):
 
     tail = 'a' if speed == 3 else 'best'
 
@@ -22,13 +21,11 @@ for ax, speed in zip(axes[1:], speeds):
         par_dict = yaml.load(f)['values']
 
     par_set = PrincipalParameterSet(par_dict)
-    #par_set.plot_geometry(ax=ax)
     par_set.plot_all(ax=ax)
-    #ax.set_aspect('equal')
     ax.set_title('{} m/s'.format(speed))
     ax.set_ylabel('')
 
-width = 12.0
-fig.set_size_inches((width, 500/1900*width))
+width = 8.0
+fig.set_size_inches((width, 4/3*width))
 fig.tight_layout()
 fig.savefig('figures/optimal-geometries.png', dpi=300)

### 2019-08-21 src/eigenvalue_plots.py
Added code to make plots of eigenvalues versus speed.

@@ -1,0 +1,41 @@
+import yaml
+import numpy as np
+import matplotlib.pyplot as plt
+
+from bicycleparameters.parameter_sets import PrincipalParameterSet
+from bicycleparameters.models import MinimalLinearWhippleCarvalloModel
+
+speeds = [3, 5, 7, 9]
+speed_range = np.linspace(0, 10, num=500)
+
+fig, axes = plt.subplots(nrows=3, ncols=2)
+
+with open('data/principal-browserjason.yml', 'r') as f:
+    par_dict = yaml.load(f)['values']
+    par_set = PrincipalParameterSet(par_dict)
+    ben_par_set = par_set.to_benchmark()
+    model = MinimalLinearWhippleCarvalloModel(ben_par_set)
+    model.plot_eigenvalue_parts(v=speed_range, ax=axes[0, 0])
+    axes[0, 0].set_ylim((-10, 10))
+    axes[0, 0].set_title('Reference')
+
+for ax, speed in zip(axes.flatten()[2:], speeds):
+
+    tail = 'a' if speed == 3 else 'best'
+
+    with open('data/principal-extended{}{}.yml'.format(speed, tail), 'r') as f:
+        par_dict = yaml.load(f)['values']
+
+    par_set = PrincipalParameterSet(par_dict)
+    ben_par_set = par_set.to_benchmark()
+    model = MinimalLinearWhippleCarvalloModel(ben_par_set)
+    model.plot_eigenvalue_parts(v=speed_range, ax=ax)
+    ax.axvline(speed, linestyle='--')
+    ax.set_title('{} m/s'.format(speed))
+    ax.set_ylim((-10, 10))
+    ax.set_ylabel('')
+
+width = 8.0
+fig.set_size_inches((width, 4/3*width))
+fig.tight_layout()
+fig.savefig('figures/optimal-eigenvalues.png', dpi=300)

### 2019-08-22 src/eigenvalue_plots.py
Made speed line black.

@@ -30,7 +30,7 @@ for ax, speed in zip(axes.flatten()[2:], speeds):
     ben_par_set = par_set.to_benchmark()
     model = MinimalLinearWhippleCarvalloModel(ben_par_set)
     model.plot_eigenvalue_parts(v=speed_range, ax=ax)
-    ax.axvline(speed, linestyle='--')
+    ax.axvline(speed, color='black', linestyle='--')
     ax.set_title('{} m/s'.format(speed))
     ax.set_ylim((-10, 10))
     ax.set_ylabel('')

### 2019-08-22 paper/paper.tex
Added eigenvalue figure to paper.
Added table of optimal values.

@@
 signficant negative trail.
+
+3
+- Rider is over 3 m in the air and leaned forward.
+- 3 m wheel base with 0.5 meter positive trail
+- wheels are over double the diameter of browser wheels
+
+5 m/s
+- rider is laid back a 0.5 meter off ground
+- < 1 m wheelbase
+
+7 m/s
+- miminal rotatoinal inertia about the steer axis
+- tall slender fork/handlebar
+- no trail
+- half sized wheels (~12 inches)
+- normal wheelbase
+- rider tipped forward and low (0.6 m)
+- vertical steer axis
+- 3 m high frame center of mass
 %
@@
 \end{table}
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
+  \caption{}
+\end{figure}
+%
+\begin{table}
+  \caption{Optimal principal parameter values for each design speed.}
+  \label{tab:optimal-values}
+  \centering
+  \begin{tabular}{lllll}
+    \toprule
+    Parameter & 3 m/s & 5 m/s & 7 m/s & 9 m/s \\
+    \midrule
+alphaD & 0.66300  & 0.91544    & 1.1228    & 1.076700 \\
+alphaH & 0.57150  & -1.2647 &    1.5708    & 0.23760 \\
+alphaP & -0.66130 & 1.1497 &     -0.83627  & 1.5584 \\
+c      & 0.68780  & -0.0050384 & 0.0010500 & -0.48397 \\
+g      & 9.8100   & 9.8100 &     9.8100 &    9.8100 \\
+kDaa   & 1.4491   & 1.2179 &     1.3242 &    0.86084 \\
+kDbb   & 0.72360  & 1.3868 &     1.2903 &    2.0139 \\
+kDyy   & 1.4711   & 1.5585 &     1.5550 &    2.0305 \\
+kFaa   & 0.35495  & 0.12585 &    0.063500 &  1.0314 \\
+kFbb   & 0.35495  & 0.12585 &    0.063500 &  1.0314 \\
+kFyy   & 0.70990  & 0.25170 &    0.12700 &   2.0628 \\
+kHaa   & 0.18620  & 0.049106 &   0.0 &       2.8463 \\
+kHbb   & 1.2578   & 1.1451 &     2.0055 &    0.16848 \\
+kHyy   & 0.63610  & 0.38324 &    0.81835 &   0.66922 \\
+kPaa   & 0.36797  & 0.36797 &    0.36797 &   0.36797 \\
+kPbb   & 0.15276  & 0.15276 &    0.15276 &   0.15276 \\
+kPyy   & 0.36717  & 0.36717 &    0.36717 &   0.36717 \\
+kRaa   & 0.47910  & 0.25306 &    0.073095 &  1.5135 \\
+kRbb   & 0.47910  & 0.25306 &    0.073095 &  1.5135 \\
+kRyy   & 0.95820  & 0.50612 &    0.14619 &   3.0270 \\
+lP     & 1.7280   & 1.7280 &     1.7280 &    1.7280 \\
+lam    & -0.21250 & 0.27137 &    -0.027560 & 0.23941 \\
+mD     & 3.2233   & 11.534 &     10.315 &    7.0223 \\
+mF     & 1.6211   & 4.4038 &     2.6441 &    5.5058 \\
+mH     & 0.25000  & 0.25000 &    0.49309 &   3.5435 \\
+mP     & 83.500   & 83.500 &     83.500 &    83.500 \\
+mR     & 12.625   & 8.3601 &     6.3080 &    2.2534 \\
+rF     & 0.70990  & 0.25170 &    0.12700 &   2.0628 \\
+rR     & 0.95820  & 0.50612 &    0.14619 &   3.0270 \\
+v      & 3.0      & 5.0 &        7.0 &       9.0 \\
+w      & 2.8655   & 0.84711 &    1.1570 &    5.5573 \\
+wP     & 0.48300  & 0.48300 &    0.48300 &   0.48300 \\
+xD     & 0.95820  & 0.28749 &    0.42710 &   1.0204 \\
+xF     & 2.8655   & 0.84711 &    1.1570 &    5.5573 \\
+xH     & 2.3561   & 0.53227 &    1.1449 &    5.6145 \\
+xP     & 0.76460  & 0.27642 &    0.52614 &   1.0791 \\
+xR     & 0.0      & 0.0 &        0.0 &       0.0 \\
+yD     & 0.0      & 0.0 &        0.0 &       0.0 \\
+yF     & 0.0      & 0.0 &        0.0 &       0.0 \\
+yH     & 0.0      & 0.0 &        0.0 &       0.0 \\
+yP     & 0.0      & 0.0 &        0.0 &       0.0 \\
+yR     & 0.0      & 0.0 &        0.0 &       0.0 \\
+zD     & -2.6050  & -2.7190 &    -2.9764 &   -4.0653 \\
+zF     & -0.70990 & -0.25170 &   -0.12700 &  -2.0628 \\
+zH     & -1.2977  & -0.78065 &   -1.3399 &   -1.7755 \\
+zP     & -3.1937  & -0.45256 &   -0.58632 &  -2.6623 \\
+zR     & -0.95820 & -0.50612 &   -0.14619 &  -3.0270 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 

### 2019-08-22 src/geometry_plot.py
Plot each indiviaul bike for close look.

@@ -10,7 +10,9 @@ fig, axes = plt.subplots(nrows=3, ncols=2)
 with open('data/principal-browserjason.yml', 'r') as f:
     par_dict = yaml.load(f)['values']
     par_set = PrincipalParameterSet(par_dict)
-    par_set.plot_all(ax=axes[0, 0])
+    bpar_set = par_set.to_benchmark()
+    bpar_set.plot_all(ax=axes[0, 0])
+    par_set.plot_person_diamond(ax=axes[0, 0])
     axes[0, 0].set_title('Reference')
 
 for ax, speed in zip(axes.flatten()[2:], speeds):
@@ -21,11 +23,27 @@ for ax, speed in zip(axes.flatten()[2:], speeds):
         par_dict = yaml.load(f)['values']
 
     par_set = PrincipalParameterSet(par_dict)
-    par_set.plot_all(ax=ax)
-    ax.set_title('{} m/s'.format(speed))
+    bpar_set = par_set.to_benchmark()
+    bpar_set.plot_all(ax=ax)
+    par_set.plot_person_diamond(ax=ax)
+    ax.set_title('{} m/s Design Speed'.format(speed))
     ax.set_ylabel('')
 
 width = 8.0
 fig.set_size_inches((width, 4/3*width))
 fig.tight_layout()
 fig.savefig('figures/optimal-geometries.png', dpi=300)
+
+
+for speed in speeds:
+
+    fig, ax = plt.subplots()
+    tail = 'a' if speed == 3 else 'best'
+
+    with open('data/principal-extended{}{}.yml'.format(speed, tail), 'r') as f:
+        par_dict = yaml.load(f)['values']
+
+    par_set = PrincipalParameterSet(par_dict)
+    par_set.plot_all(ax=ax)
+    ax.set_title('{} m/s Design Speed'.format(speed))
+    fig.savefig('figures/optimal-geometry-{}.png'.format(speed), dpi=300)

### 2019-08-30 paper/paper.tex
Removed citations from abstract and made parameter table consistent.

--- 2019-08-21 paper/paper.tex
@@
 {\fontsize{16pt}{20pt}\selectfont%
-  Expanded Optimization for Discovering\\}
-{\fontsize{16pt}{20pt}\selectfont%
-  Optimal Lateral Handling Bicycles}
+  Expanded Optimization for Discovering Optimal Lateral Handling Bicycles}
 \end{flushleft}
@@
 %
-We introduced a method of optimizing four geometric parameters of a bicycle's
-design to minimize the so called Handling Quality Metric
-(HQM)~\cite{Moore2016}. Here we expand that method to optimize all X geometric
-and inertial parameters of the benchmark parameterimization of the linear
-Whipple-Carvallo bicycle model under constraints that guarantee a physically
-realizable bicycle. This improves over the prior work by expanding the search
-space with many more parameters and the guarntee of realizability.
+Previously, we introduced a method of optimizing four primary geometric
+parameters of a bicycle's design to minimize its Handling Quality Metric. Here
+we expand that method to optimize 23 geometric and inertial parameters in the
+benchmark parameterimization of the linear Whipple-Carvallo bicycle model. To
+ensure physically realizable bicycle designs we include 7 equality constraints,
+21 inequality constraints, and maximal and minamal bounds on each free
+opmization parameter. This improves over the prior work by expanding the search
+space with many more parameters and the guarntee of realizability. We close by
+presenting five bicycle designs that have optimal later handling qualities. The
+bicycles are not generally self-stable and have some ununusal characteristics.
 
@@
 Physical design features of ground vehicles can affect their lateral handling
-qualities. Geometry, mass, and mass distribution of the vehicle's primary
-components as well as tire characteristics are primary contributors to poor and
-good handling due to their important influence on the vehicle's dynamics. In
-past work, we have presented a theoretical and computational framework for
-assessing the lateral task-independent handling qualities of simplified single
-track vehicle designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed
-that minimizing our proposed handling quality metric (HQM) can produce theoretically
+qualities. Geometry, mass, and mass distribution of the vehicle's components as
+well as tire characteristics are primary contributors to poor and good handling
+due to their important influence on the vehicle's dynamics. In past work, we
+have presented a theoretical and computational framework for assessing the
+lateral task-independent handling qualities of simplified single track vehicle
+designs~\cite{Hess2012,Moore2012}. In subsequent work, we showed that
+minimizing our proposed handling quality metric (HQM) can produce theoretically
 optimal handling designs when only four geometric parameters are explored as
@@
 expand this optimization problem to all of the geometry, mass, and inertial
-parameters present in the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}.
-This broadens the search space considerably but we constrain it so that only
-realizable optimal bicycle designs are discovered. To do so, we formulate a
-constrained optimization problem and use derivative-free optimization to
-discover optimal, yet realizable, bicycle designs.
+parameters present in the linear Whipple-Carvallo bicycle
+model~\cite{Meijaard2007}.  This broadens the search space considerably but we
+constrain it so that only realizable optimal bicycle designs are discovered. To
+do so, we formulate a constrained optimization problem and use derivative-free
+optimization to discover optimal, yet realizable, bicycle designs.
 
@@
 the ``principal parameterizatoin'' as opposed to the ``benchmark
-parameterization''. This parameterization differs from the benchmark
-parameterization in three ways.  Firstly, the person and rear frame are treated
-as separate rigid bodies each with their on inertial parameters. Secondly, we
-express the inertial parameters of each rigid body in terms of central
-principal radii of gyration to decouple the mass from the inertia terms.
-Lastly, we introduce two simple dimensional parameters that define the
-geometric extents of the person which are used to constrain the location of the
-person's body. Table~\ref{tab:parameters} provides the parameters and the
-reference values which are derived from the measurements of a Batavus Browser
-Bicycle and the rider Jason presented in~\cite{Moore2012}. This
-parameterization can be transformed into the benchmark parameterization
-readily, but not vice versa.
+parameterization'' in~\cite{Meijaard2007}. This parameterization differs from
+the benchmark parameterization in three ways. Firstly, the person and rear
+frame are treated as separate rigid bodies each with their on inertial
+parameters. Secondly, we express the inertial parameters of each rigid body in
+terms of central principal radii of gyration to decouple the mass from the
+inertia terms.  Lastly, we introduce two simple dimensional parameters that
+define the geometric extents of the person which are used to constrain the
+location of the person's body. Table~\ref{tab:parameters} provides the
+parameters and the reference values which are derived from the measurements of
+a Batavus Browser Bicycle and the rider ``Jason'' presented
+in~\cite{Moore2012}. This parameterization can be transformed into the
+benchmark parameterization readily, but not vice versa.
 %
@@
   \caption{Full set of 47 principal parameters and their default values derived
-  from the measurements in~\cite{Moore2012}.}
+    from the measurements in~\cite{Moore2012} of the Batavus Browser bicycle
+    and rider ``Jason''.}
   \label{tab:parameters}
@@
     \midrule
+    % 5
     $c$       &  0.068581           & \si{\meter} & Trail \\
@@
     $g$       &  9.81               & \si{\meter\per\second\squared} & Acceleration due to gravity \\
-    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
+    $v$       &  3.0, 5.0, 7.0, 9.0 & \si{\meter\per\second} & Forward speed \\
     Rear Wheel [R] & & \\
     \midrule
-    $m_R$ &  3.11 & \si{\kilogram} & mass \\
-    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
-    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
-    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
-    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
-    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
-    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
-    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
+    % 8
+    $m_R$     & 3.11 & \si{\kilogram}  & Mass \\
+    $r_R$     & 0.34096 & \si{\meter}  & Radius \\
+    $x_R$     & 0.0 & \si{\meter}      & $X$ mass center coordinate \\
+    $y_R$     & 0.0 & \si{\meter}      & $Y$ mass center coordinate \\
+    $z_R$     & -0.34096 & \si{\meter} & $Z$ mass center coordinate \\
+    $k_{Raa}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
+    $k_{Rbb}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
+    $k_{Ryy}$ & 0.22136 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
     Front Wheel [F] & & \\
     \midrule
+    % 8
     $m_F$     & 2.02 & \si{\kilogram} & Mass \\
     $r_F$     & 0.34353 & \si{\meter} & Radius \\
-    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
-    $y_F$ & 0.0 & \si{\meter} & Mass center \\
-    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
-    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
+    $x_F$     & 1.1210 & \si{\meter} & $X$ mass center coordinate \\
+    $y_F$     & 0.0 & \si{\meter} & $Y$ mass center coordinate \\
+    $z_F$     & -0.34353 & \si{\meter} & $Z$ mass center coordinate \\
+    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
+    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
+    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radii of gyration about $Y$ \\
     Person [P] & & \\
     \midrule
-    $l_P$ &  1.7280 & \si{\meter} & Body length \\
-    $w_P$ &  0.48300 & \si{\meter} & Body width \\
-    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
-    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
-    $y_P$ & 0.0 & \si{\meter} & Mass center \\
-    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
-    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
-    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
+    % 10
+    $l_P$      &  1.728   & \si{\meter} & Body length \\
+    $w_P$      &  0.483   & \si{\meter} & Body width \\
+    $m_P$      &  83.5    & \si{\kilogram} & Mass \\
+    $x_P$      &  0.31577 & \si{\meter} & $X$ mass center coordinate \\
+    $y_P$      & 0.0      & \si{\meter} & $Y$ mass center coordinate \\
+    $z_P$      & -1.0990  & \si{\meter} & $Z$ mass center coordinate \\
+    $k_{Paa}$  &  0.36797 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
+    $k_{Pbb}$  &  0.15276 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
+    $k_{Pyy}$  &  0.36717 & \si{\meter} & Central principal radii of gyration about $Y$ \\
+    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $A_P$ and $X$ \\
     Front Frame [H] & & \\
     \midrule
+    % 8
     $m_H$ & 3.2200 & \si{\kilogram}     & Mass \\
-    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
-    $y_H$ & 0.0 & \si{\meter} & Mass center \\
-    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
-    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
-    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
+    $x_H$ & 0.86695 & \si{\meter}       & $X$ mass center coordinate \\
+    $y_H$ & 0.0 & \si{\meter}           & $Y$ mass center coordinate \\
+    $z_H$ & -0.74824 & \si{\meter}      & $Z$ mass center coordinate \\
+    $k_{Haa}$ & 0.29556 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
+    $k_{Hbb}$ & 0.14493 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
+    $k_{Hyy}$ & 0.27630 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
+    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $A_H$ and $X$ \\
     Rear Frame [D] & & \\
     \midrule
+    % 8
     $m_D$ &  9.8600 & \si{\kilogram}    & Mass \\
-    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
-    $y_D$ & 0 & \si{\meter} & Mass center \\
-    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
-    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
-    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
-    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
+    $x_D$ &  0.27595 & \si{\meter}      & $X$ mass center coordinate \\
+    $y_D$ & 0 & \si{\meter}             & $Y$ mass center coordinate \\
+    $z_D$ & -0.53784 & \si{\meter}      & $Z$ mass center coordinate \\
+    $k_{Daa}$ &  0.28587 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
+    $k_{Dbb}$ &  0.22079 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
+    $k_{Dyy}$ &  0.36539 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
+    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $A_D$ and $X$ \\
     \bottomrule
@@
 signficant negative trail.
+
+3
+- Rider is over 3 m in the air and leaned forward.
+- 3 m wheel base with 0.7 meter positive trail
+- wheels are over double the diameter of browser wheels
+- minimum front frame mass
+
+5 m/s
+- rider is laid back a 0.5 meter off ground
+- < 1 m wheelbase
+- essentially zero trail (5 mm negative)
+- minimum front frame mass
+
+7 m/s
+- miminal rotatoinal inertia about the steer axis
+- tall slender fork/handlebar
+- no trail (1 mm postive)
+- half sized wheels (~12 inches)
+- normal wheelbase (1.157)
+- rider tipped forward and low (0.6 m)
+- vertical steer axis
+- 3 m high frame center of mass
+
+9
+- very large 5.6 m wheelbase
 %
@@
 \end{table}
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
+  \caption{}
+\end{figure}
+%
+\begin{table}
+  \caption{Optimal principal parameter values for each design speed.}
+  \label{tab:optimal-values}
+  \centering
+  \begin{tabular}{lSSSS}
+    \toprule
+    Parameter & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
+    \midrule
+    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
+    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
+    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
+    \midrule
+    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
+    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
+    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
+    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
+    % Should aa be larger than bb?
+    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
+    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
+    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
+    \midrule
+    $m_H$  & 0.25  & 0.25 &    0.49 &   3.54 \\
+    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
+    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
+    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
+    $k_{Haa}$ & 0.186  & 0.0491 &   0.000 &       2.846 \\
+    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
+    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
+    \midrule
+    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
+    $r_F$     & 0.710 & 0.252 & 0.127 & 2.063 \\
+    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
+    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
+    \midrule
+    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
+    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
+    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
+    \midrule
+    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
+    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
+    $k_Raa$ & 0.479  & 0.253 & 0.073 & 1.514 \\
+    $k_Ryy$ & 0.958  & 0.506 & 0.146 & 3.027 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 

### 2019-08-31 paper/paper.tex
Reworked the initial explanations of the constraints.

@@
 The optimal principal parameters are subject to a set of constraints designed
-to ensure that a physically realizable bicycle is obtained form the opimization
+to ensure that a physically realizable bicycle is obtained from the opimization
 procedure. These constraints are made up of bounds on the free parameters and
@@
 \begin{description}
-  \item[Total] The resulting combination of the five rigid bodies.
+  \item[Total $T$] The resulting combination of the five rigid bodies.
     \begin{itemize}
-      \item The physical extents of the rigid bodies must exist above the
-        ground plane.
+      \item The likely physical extents of the rigid bodies must exist above
+        the ground plane.
       \item Both bicycle and rider are symmetric about the rider's sagittal
         plane.
-      \item The mass of each bicycle rigid body is positive, greater than a
-        minimum value, and the total mass is below a reasonably lift-able
-        amount.
+      \item The total mass is below a reasonably human lift-able amount.
       \item The wheels cannot overlap.
@@
         during hard acceleration.
+      \item The closed loop path tracking system~\cite{Hess2012} must be
+        stable. This is required to obtain a valid HQM value.
     \end{itemize}
 
-  \item[Wheels] Both front and rear wheels have identical constraints.
-    \begin{itemize}
-      \item Wheel radius and mass must be greater than a minimum value.
-      \item Wheels are inertially wheel-like, i.e. symmetric about each plane
-        and most of the mass is at the rim.
-    \end{itemize}
-
-  \item[Frames] Front frame (handlebar + fork) and rear frame
-    \begin{itemize}
-      \item The mass and inertia of the frames are positive and large enough to
-        be constructed from materials in a space frame of specified minimal
-        density.
-      \item The rear frame is planar in nature and the front frame's moments of
-        inertia are consitently dependent.
-    \end{itemize}
-
-  \item[Rider] A single rigid body represents the rider.
+  \item[Rider $P$] A single rigid body represents the rider.
     \begin{itemize}
@@
         derived from standard body segment estimation methods.
+      \item The rider cannot penetrate the ground.
+    \end{itemize}
+
+  \item[Frames $H$, $D$] Front frame (handlebar + fork) and rear frame
+    \begin{itemize}
+      \item The rear frame is planar in nature and the front frame's moments of
+        inertia are consistently dependent.
+      \item The mass and inertia of the frames are positive and large enough to
+        be constructed from a steel space frame.
+    \end{itemize}
+
+  \item[Wheels $F$, $R$] Both front and rear wheels have identical constraints.
+    \begin{itemize}
+      \item Wheel radius and mass must positive and be greater than a minimum
+        value.
+      \item Wheels are inertially wheel-like, i.e. symmetric about each plane
+        and most of the mass is at the rim.
     \end{itemize}
@@
     \toprule
-    Min & & Parmeter & & Max\\
+    Min & & Parameter & & Max\\
     \midrule
@@
 
-\subsubsection*{Rear Frame [D]}
+\subsection{Person [P]}
+%
+We assume that the person's joint configurations are such that they are in a
+nominal configuration for pedaling, i.e. an average normal everyday riding
+position on a typical bicycle, i.e. they stayed in the same configuration as
+they were seated on the Batavus Browser bicycle. The person is assumed to be
+symmetric about the $XZ$ plane. We allow the rider to be rotated about the $Y$
+axis and positioned anywhere within the plane of symmetry above the ground.
+
+To prevent the rider from being positioned and oriented such that their body is
+not penetrating the ground we introduce two dimensions that define a cross
+whose apex is at the center of mass of the person and the cross axes are
+parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
+along the principal axis to the tip of the toes and $w_P$ is the distance along
+the second principal axes to the tip of the hands. Constraints
+$c_2,\ldots,c_5$.
+
+\subsection{Front Frame [H]}
+%
+The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. We
+allow for any angular orientation of the $XZ$ principal directions but limit
+the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.  We prevent the
+rear frame from penetrating the ground by limiting the inertial spread with
+respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
+ensure a frame can span from the rear wheel to the mass center of the rear
+frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
+geometrical spread of a typical bicycle frame and its radius of gyration. The
+front frame is not planar but is narrow with respect to the $XZ$ plane, which
+is enforced by constraint $c_1$.
+
+\subsection{Rear Frame [D]}
 %
@@
 
-\subsection{Person [P]}
-
-We assume that the person's joint configurations are such that they are in a
-nominal configuration for pedaling, i.e. an average normal everyday riding
-position on a typical bicycle, i.e. they stayed in the same configuration as
-they were seated on the Batavus Browser bicycle. The person is assumed to be
-symmetric about the $XZ$ plane. We allow the rider to be rotated about the $Y$
-axis and positioned anywhere within the plane of symmetry above the ground.
-
-To prevent the rider from being positioned and oriented such that their body is
-not penetrating the the ground we introduce two dimensions that define a cross
-whose apex is at the center of mass of the person and the cross axes are
-parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
-along the principal axis to the tip of the toes and $w_P$ is the distance along
-the second principal axes to the tip of the hands. Constraints
-$c_2,\ldots,c_5$.
-
-\subsection{Front Frame}
-%
-The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. We
-allow for any angular orientation of the $XZ$ principal directions but limit
-the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$.  We prevent the
-rear frame from penetrating the ground by limiting the inertial spread with
-respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
-ensure a frame can span from the rear wheel to the mass center of the rear
-frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
-geometrical spread of a typical bicycle frame and its radius of gyration. The
-front frame is not planar but is narrow with respect to the $XZ$ plane, which
-is enforced by constraint $c_1$.
-
 \subsection{Front [F] and Rear [R] Wheels}
@@
 simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than a
-minimal size based on small purchable spoked wheel with tire.
+minimal size based on small purchasable spoked wheel with tire.
 
@@
 The trail and wheelbase can take on any real values. The steer axis tilt is
-limted to 180 degrees. We introduce a constraint $c_{10}$ that prevents the
-wheels from physically overlapping and require that the bicycle be liftable by
+limited to 180 degrees. We introduce a constraint $c_{10}$ that prevents the
+wheels from physically overlapping and require that the bicycle be lift-able by
 an average person, $c_{11}$. Finally, we require that the bicycle not topple
@@
 
-\section{Free parameters}
+\section{Optimization}
 %
@@
 
-\section{Optimization}
-%
 Our objective in the optimization is to minimize the peak HQM value subject to

### 2019-09-01 paper/paper.tex
Edited down to results.

@@
 of the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}. We call this
-the ``principal parameterizatoin'' as opposed to the ``benchmark
+the ``principal parameterization'' as opposed to the ``benchmark
 parameterization'' in~\cite{Meijaard2007}. This parameterization differs from
@@
 %
+\begin{figure}
+  \centering
+  \includegraphics[width=4in]{../figures/reference-geometry.png}
+  \caption{Depiction of the principal parameters of the Batavus Browser
+    with rider ``Jason''. The solid black lines represent the essential bicycle
+    geometry. The dotted black line represents the steer axis. The inertial
+    properties of the five rigid bodies: front wheel (orange), rear wheel
+    (purple), rear frame (blue), front frame (green), and person (red) are
+    shown with the mass center and the extents of the centroidal principal
+    radii of gyration of each rigid body. The primary principal angles,
+    $\alpha_{D,P,H}$, are defined as the angle about the $y$ axis from $x$ to
+    the maximum principal axes in the $XZ$ plane.}
+\end{figure}
+%
 \begin{table}
@@
     from the measurements in~\cite{Moore2012} of the Batavus Browser bicycle
-    and rider ``Jason''.}
+    and rider ``Jason''. Note that the benchmark parameters are also used in
+    the paper and are defined in~\cite{Meijaard2007}.}
   \label{tab:parameters}
+  \small
   \centering
@@
 The optimal principal parameters are subject to a set of constraints designed
-to ensure that a physically realizable bicycle is obtained from the opimization
-procedure. These constraints are made up of bounds on the free parameters and
-both equality and inequality constraints among the parameters. Below the basic
-constraint concepts presented and grouped by the associated rigid body or
-collection thereof:
+to ensure that a physically realizable bicycle is obtained from the
+optimization procedure. These constraints are made up of bounds on the free
+parameters and both equality and inequality constraints among the parameters.
+Below the basic constraint concepts presented and grouped by the associated
+rigid body or collection thereof:
 %
 \begin{description}
-  \item[Total $T$] The resulting combination of the five rigid bodies.
+  \item[Total $T$] The combination of the five rigid bodies.
     \begin{itemize}
@@
         during hard acceleration.
-      \item The closed loop path tracking system~\cite{Hess2012} must be
-        stable. This is required to obtain a valid HQM value.
+      \item The closed-loop path tracking controlled system~\cite{Hess2012}
+        must be stable. This is required to obtain a valid HQM value.
     \end{itemize}
@@
     \begin{itemize}
-      \item Wheel radius and mass must positive and be greater than a minimum
-        value.
+      \item Wheel radius and mass must be positive and be greater than a
+        minimum value.
       \item Wheels are inertially wheel-like, i.e. symmetric about each plane
@@
 \begin{table}
-  \caption{Free parameter upper and lower bounds}
+  \caption{Parameter lower and upper bounds.}
   \label{tab:bounds}
@@
     $-\pi/2 $ & $\leq$ & $\lambda $  &  $\leq$  &  $\pi/2$ \\
-    $1.0    $ & $\leq$ & $m_D     $  &  $\leq$  &  $\infty$ \\
+    $1.0\si{\kilogram}$ & $\leq$ & $m_D     $  &  $\leq$  &  $\infty$ \\
     $-\infty$ & $\leq$ & $x_D     $  &  $\leq$  &  $\infty$ \\
@@
     $-\pi/2 $ & $\leq$ & $\alpha_H$  &  $\leq$  &  $\pi/2$ \\
-    $0.127  $ & $\leq$ & $r_R     $  &  $\leq$  &  $\infty$ \\
-    $1.0    $ & $\leq$ & $m_R     $  &  $\leq$  &  $\infty$ \\
-    $0.127  $ & $\leq$ & $r_F     $  &  $\leq$  &  $\infty$ \\
-    $1.0    $ & $\leq$ & $m_F     $  &  $\leq$  &  $\infty$ \\
+    $0.127\si{\meter}$ & $\leq$ & $r_R     $  &  $\leq$  &  $\infty$ \\
+    $1.0\si{\kilogram}$ & $\leq$ & $m_R     $  &  $\leq$  &  $\infty$ \\
+    $0.127\si{\meter}$ & $\leq$ & $r_F     $  &  $\leq$  &  $\infty$ \\
+    $1.0\si{\kilogram}$ & $\leq$ & $m_F     $  &  $\leq$  &  $\infty$ \\
     \bottomrule
@@
 \begin{table}
-  \caption{Equality constraints}
+  \caption{Equality constraints.}
   \label{tab:equality-constraints}
@@
 \begin{table}
-  \caption{Inequality constraints}
+  \caption{Inequality constraints.}
   \label{tab:inequality-constraints}
@@
 nominal configuration for pedaling, i.e. an average normal everyday riding
-position on a typical bicycle, i.e. they stayed in the same configuration as
-they were seated on the Batavus Browser bicycle. The person is assumed to be
-symmetric about the $XZ$ plane. We allow the rider to be rotated about the $Y$
-axis and positioned anywhere within the plane of symmetry above the ground.
+position on a typical bicycle. We retain the same configuration as they were
+seated on the Batavus Browser bicycle. The person is assumed to be symmetric
+about the $XZ$ plane. We allow the rider to be rotated about the $Y$ axis and
+positioned anywhere within the plane of symmetry above the ground.
 
@@
 parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
-along the principal axis to the tip of the toes and $w_P$ is the distance along
-the second principal axes to the tip of the hands. Constraints
-$c_2,\ldots,c_5$.
+along the principal axis to the tip of the toes and $w_P / 2$ is the distance
+along the second principal axes to the tip of the hands. The constraints
+$c_2,\ldots,c_5$ are derived from these rules.
 
@@
 The trail and wheelbase can take on any real values. The steer axis tilt is
-limited to 180 degrees. We introduce a constraint $c_{10}$ that prevents the
-wheels from physically overlapping and require that the bicycle be lift-able by
-an average person, $c_{11}$. Finally, we require that the bicycle not topple
-forward during hard breaking or backward during hard acceleration.
+limited to $\pm90$ degrees. We introduce a constraint $c_{10}$ that prevents
+the wheels from physically overlapping and require that the bicycle be
+lift-able by an average person, $c_{11}$. Finally, we require that the bicycle
+not topple forward during hard breaking or backward during hard acceleration
+with:
 
 \begin{align}
-  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}
+  -\frac{3g}{4} < \textrm{acceleration} < \frac{g}{4}\textrm{.}
 \end{align}
@@
 mass $(x_T,z_T)$ in a triangle in the $XZ$ plane. Lastly, we constrain the
-eight closed loop eigenvalues associated with the controller in
-~\cite{Hess2012} to be stable, i.e. have negative real parts.  These are
-expressed in constraints $c_{14},\ldots,c_{21}$. Closed loop stability is
-required for the HQM to provide a meaningful result.
+eight closed loop eigenvalues associated with the controller in~\cite{Hess2012}
+to be stable, i.e. have negative real parts. These are expressed in
+constraints $c_{14},\ldots,c_{21}$. Closed loop stability is required for the
+HQM to provide a meaningful result.
 
@@
 %
-The above constraints leaves 23 of the 47 parameters free for optimizing which
+The above constraints leave 23 of the 47 parameters free for optimizing which
 we collect in the vector $\mathbf{p}\in\mathbb{R}^{23}$ and define as:
@@
 this limitation we move the inequality constraints into the objective function
-an penalize the objective if the constraints are violated with the following
+and penalize the objective if the constraints are violated with the following
 rules:
@@
   \begin{cases}
-    max(\textrm{HQM}(\mathbf{p})) & \textrm{if} \quad any(\mathbf{g}(\mathbf{p})) < 0 \\
+    max(\textrm{HQM}(\mathbf{p})) & \textrm{if} \quad all(\mathbf{g}(\mathbf{p})) \leq 0 \\
     30 + ||\mathbf{g}_{+}(\mathbf{p})||/10 & \textrm{if} \quad
-      any(\mathbf{g}(\mathbf{p})) \geq 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| < 30 \\
+      any(\mathbf{g}(\mathbf{p})) > 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| < 30 \\
     ||\mathbf{g}_{+}(\mathbf{p})|| & \textrm{if} \quad
-      any(\mathbf{g}(\mathbf{p})) \geq 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| \geq 30
+      any(\mathbf{g}(\mathbf{p})) > 0 \textrm{ and } ||\mathbf{g}_{+}(\mathbf{p})|| \geq 30
   \end{cases}
 \end{align}
-where $||\mathbf{g}_{+}||$ is the norm of the positive elements of
-$\mathbf{g}$.
+where $||\mathbf{g}_{+}||$ is the norm of the vector of positive elements of
+$\mathbf{g}$ and $any,all$ are ``any elements of'' and ``all elements of'',
+respectively.
 
-This creates a discontinous objective function but in practice the CMS-ES
-algorthim is able to move into the parameter space where all the constraints
+This creates a discontinuous objective function but in practice the CMA-ES
+algorithm is able to move into the parameter space where all the constraints
 are satisfied and find a (local) minima. For our purposes, this sufficiently
-finds parameter values that produce an optimally handling design.
+finds parameter values that produce optimally handling bicycle designs.
 
@@
 %
-We discover four bicycles for four different design speeds (3, 5, 7, and 9 m/s)
-that have an optimally low HQM, see Table~\ref{tab:hqm} and satisify all
-constraints and parameter boundaries. We belive these bicycles to be physically
-realizable with minor differences. The most striking feature is that all of the
-bicycle are larger that the reference bicycle in some way.  The 3 m/s bicycle
-places both the rider's mass center and the rear frame's mass center close to 3
-meters above the ground plane. The 5 and 7 m/s bicycles have wheelbase values
-that are simlar to the reference bicycle, but 3 and 9 m/s have very large wheel
-bases. The 9 m/s bicycle is very large overall. Only the 5 m/s bicycle is of a
-scale close to a typical bicycle. It is noteable that the 9 m/s bicycle has
-signficant negative trail.
+We discover four bicycles for different design speeds (3, 5, 7, and
+9~\si{\meter\per\second}) that have an optimally low HQM, see
+Table~\ref{tab:hqm}, and satisfy all constraints and parameter bounds. The
+optimal parameters for these four bicycles are presented in
+Table~\ref{tab:optimal-values}.
+%
+\begin{table}
+  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
+    at each design speed.}
+  \label{tab:hqm}
+  \centering
+  \begin{tabular}{llll}
+    \toprule
+    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
+    \midrule
+    3 & 13.075 & 2.012 & 85\% \\
+    5 & 4.521  & 0.012 & 100\% \\
+    7 & 3.043  & 0.022 & 99\% \\
+    9 & 2.338  & 0.839 & 64\% \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+%
+\begin{table}
+  \caption{Optimal principal parameter values for each design speed. Values
+    with an asterisk are at a bound.}
+  \label{tab:optimal-values}
+  \centering
+  \begin{tabular}{lSSSS}
+    \toprule
+    $v$ & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
+    \midrule
+    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
+    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
+    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
+    \midrule
+    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
+    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
+    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
+    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
+    % Should aa be larger than bb?
+    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
+    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
+    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
+    \midrule
+    $m_H$  & 0.25*  & 0.25* &    0.49 &   3.54 \\
+    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
+    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
+    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
+    $k_{Haa}$ & 0.186  & 0.0491 &   0.000* &       2.846 \\
+    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
+    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
+    \midrule
+    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
+    $r_F$     & 0.710 & 0.252 & 0.127* & 2.063 \\
+    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
+    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
+    \midrule
+    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
+    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
+    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
+    \midrule
+    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
+    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
+    $k_{Raa}$ & 0.479  & 0.253 & 0.073 & 1.514 \\
+    $k_{Ryy}$ & 0.958  & 0.506 & 0.146 & 3.027 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+The pictorial representation of the bicycles are presented in
+Figure~\ref{fig:optimal-geometries}. The most striking feature is that all of
+the bicycle are larger that the reference bicycle in some way.
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=6in]{../figures/optimal-geometries.png}
+  \label{fig:optimal-geometries}
+  \caption{Depictions of the bicycle geometry and geometric representations of
+    the inertial quantities for the reference bicycle and four optimal solutions
+    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
+    wheel (orange), rear wheel (purple), rear frame (blue), front frame
+    (green), and person (red). The solid black lines represent the essential
+    bicycle geometry. The dotted black line represents the steer axis. The
+    solid colored curves represent the contours of solid ellipsoids with
+    equivalent inertia as the principal inertia of the associated rigid body.
+    The dotted colored lines represent the extents of the centroidal radii of
+    gyration of each rigid body.}
+\end{figure}
+
+We
+belive these bicycles to be physically realizable with minor differences.
+
+  The 3 m/s bicycle places both the rider's mass center and
+the rear frame's mass center close to 3 meters above the ground plane. The 5
+and 7 m/s bicycles have wheelbase values that are simlar to the reference
+bicycle, but 3 and 9 m/s have very large wheel bases. The 9 m/s bicycle is very
+large overall. Only the 5 m/s bicycle is of a scale close to a typical bicycle.
+It is noteable that the 9 m/s bicycle has signficant negative trail.
 
@@
   \centering
-  \includegraphics[width=6in]{../figures/optimal-geometries.png}
-  \caption{Depcitions of the bicycle geometry and geometric representations of
-    the inertial quanties for the reference bicycle and four optimal solutions
-    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
-    wheel (oragne), rear wheel (purple), rear frame (blue), front frame
-    (green), and person (red). The solid black lines represent the essential
-    bicycle geometry. The dotted black line represents the steer axis. The
-    solid colored curves represent the contours of solid ellipsoids with
-    equivalent inertia as the principal inertia of the associated rigid body.
-    The dotted colored lines represent the extents of the centerial radii of
-    gyration of each rigid body.}
-\end{figure}
-%
-\begin{table}
-  \caption{Peak HQM values for the reference bicycle and the optimal bicycles
-    at each speed.}
-  \label{tab:hqm}
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-    Speed [m/s] & Reference Peak HQM & Optimal Peak HQM & Percent Improvement \\
-    \midrule
-    3 & 13.0753 & 2.0118 & 85\% \\
-    5 & 4.5213  & 0.0115 & 100\% \\
-    7 & 3.0434  & 0.0220 &  99\% \\
-    9 & 2.3377  & 0.8386 &  64\% \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-%
-\begin{figure}
-  \centering
   \includegraphics[width=6in]{../figures/optimal-eigenvalues.png}
+  \label{fig:optimal-eigenvalues}
   \caption{}
 \end{figure}
-%
-\begin{table}
-  \caption{Optimal principal parameter values for each design speed.}
-  \label{tab:optimal-values}
-  \centering
-  \begin{tabular}{lSSSS}
-    \toprule
-    Parameter & 3~\si{\meter\per\second} & 5~\si{\meter\per\second} & 7~\si{\meter\per\second} & 9~\si{\meter\per\second} \\
-    \midrule
-    $c$       & 0.688  & -0.005 & 0.001  & -0.484 \\
-    $w$       & 2.866  & 0.847  & 1.157  & 5.557  \\
-    $\lambda$ & -0.213 & 0.271  & -0.028 & 0.239  \\
-    \midrule
-    $m_D$ & 3.22   &  11.53 & 10.32  & 7.02 \\
-    $x_D$ & 0.958  &  0.287 &  0.427 & 1.020 \\
-    $z_D$ & -2.605 & -2.719 & -2.976 & -4.065 \\
-    $\alpha_D$ & 0.663  & 0.915  & 1.123  & 1.077 \\
-    % Should aa be larger than bb?
-    $k_{Daa}$  & 1.449  & 1.218  &  1.324 & 0.861 \\
-    $k_{Dbb}$  & 0.724  & 1.387  &  1.290 & 2.014 \\
-    $k_{Dyy}$  & 1.471  & 1.559  &  1.555 & 2.031 \\
-    \midrule
-    $m_H$  & 0.25  & 0.25 &    0.49 &   3.54 \\
-    $x_H$     & 2.356   & 0.532 &    1.145 &    5.615 \\
-    $z_H$     & -1.298  & -0.781 &   -1.340 &   -1.776 \\
-    $\alpha_H$ & 0.572  & -1.265 & 1.571  & 0.238 \\
-    $k_{Haa}$ & 0.186  & 0.0491 &   0.000 &       2.846 \\
-    $k_{Hbb}$ & 1.259   & 1.145 &     2.006 &    0.168 \\
-    $k_{Hyy}$ & 0.636  & 0.383 &    0.818 &   0.669 \\
-    \midrule
-    $m_F$     & 1.62  & 4.40  & 2.64  & 5.51  \\
-    $r_F$     & 0.710 & 0.252 & 0.127 & 2.063 \\
-    $k_{Faa}$ & 0.355 & 0.126 & 0.064 & 1.031 \\
-    $k_{Fyy}$ & 0.710 & 0.252 & 0.127 & 2.063 \\
-    \midrule
-    $x_P$      & 0.765  & 0.276  &  0.526 & 1.079 \\
-    $z_P$      & -3.194 & -0.453 & -0.586 &-2.662 \\
-    $\alpha_P$ & -0.661 & 1.150  & -0.836 & 1.558 \\
-    \midrule
-    $m_R$   & 12.63  & 8.36  &  6.31 & 2.25 \\
-    $r_R$   & 0.958  & 0.506 & 0.146 & 3.027 \\
-    $k_Raa$ & 0.479  & 0.253 & 0.073 & 1.514 \\
-    $k_Ryy$ & 0.958  & 0.506 & 0.146 & 3.027 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
 
@@
 %
+- need constraints to keep rider out of wheels
+- need more realistic constraints for the frame mass and inertia based on space
+frame
+\section{Reproducibility}
+%
+All of the source code, data, and documents needed to reproduce the presented
+results and this paper can be found at the repository hosted at
+\url{https://github.com/moorepants/BMD2019}.
 

### 2019-09-01 paper/paper.tex
Some results in place.

@@
 Table~\ref{tab:hqm}, and satisfy all constraints and parameter bounds. The
-optimal parameters for these four bicycles are presented in
-Table~\ref{tab:optimal-values}.
+optimal parameter values for these four bicycles are presented in
+Table~\ref{tab:optimal-values}. We belive these bicycles to be reasoably
+physically realizable. The pictorial representation of the bicycles are
+presented in Figure~\ref{fig:optimal-geometries}.
 %
@@
 \end{table}
-
-The pictorial representation of the bicycles are presented in
-Figure~\ref{fig:optimal-geometries}. The most striking feature is that all of
-the bicycle are larger that the reference bicycle in some way.
 %
@@
 
-We
-belive these bicycles to be physically realizable with minor differences.
+The bicycle optimized for 3~\si{\meter\per\second} is about three times larger
+than the reference bicycle. The wheelbase is about 3~\si{\meter} and it has a
+relatively large postive trail (0.7~\si{\meter}). The person is rearward on the
+bike and over 3~\si{\meter} above the ground. The mass of the front frame is
+the minimum posible value and the mass of the rear frame is also small.
 
-  The 3 m/s bicycle places both the rider's mass center and
-the rear frame's mass center close to 3 meters above the ground plane. The 5
-and 7 m/s bicycles have wheelbase values that are simlar to the reference
-bicycle, but 3 and 9 m/s have very large wheel bases. The 9 m/s bicycle is very
-large overall. Only the 5 m/s bicycle is of a scale close to a typical bicycle.
-It is noteable that the 9 m/s bicycle has signficant negative trail.
+The bicycle optimized for 5~\si{\meter\per\second} has a similar geometric
+scale as the reference bicycle except that the front wheel is smaller and the
+rear wheel is larger. The person is located low to the ground and tipped back,
+like one might see on a recumbent bicycle.  This significantly increases the
+vehicles yaw moment of inertia. The trail is miminal, but suprisingly slight
+negative (5~\si{\milli\meter}). The steer axis tilt is slightly shallower than
+the reference bicycle. The mass of the rear frame is similar to the reference
+bicycle but the front frame is at the bound and much lower. The minor principal
+axis of the front frame is almost aligned with the steer axis.
 
-3
-- Rider is over 3 m in the air and leaned forward.
-- 3 m wheel base with 0.7 meter positive trail
-- wheels are over double the diameter of browser wheels
-- minimum front frame mass
+The bicycle optimized for 5~\si{\meter\per\second} is one of the three that is
+geometrically smaller than the reference bicycle. The wheelbase is slightly
+larger but the wheels are approximately half the diameter. The rider is low to
+the ground and tipped forward. There is effiectively no trail, the steer axis
+is vertical, and the (very) minimal principal axis is aligned with the steer
+axis making the steering inertia-less. The roll moment of inertia is larger.
 
-5 m/s
-- rider is laid back a 0.5 meter off ground
-- < 1 m wheelbase
-- essentially zero trail (5 mm negative)
-- minimum front frame mass
+The bicycle optimized for 5~\si{\meter\per\second} is much larger than the
+reference bicycle: wheelbase and rear wheel diameter of 6~\si{\meter}, front
+wheel diameter of 4~\si{\meter}. The trail is large and negative, surely making
+the bicycle open-loop unstable.
 
-7 m/s
-- miminal rotatoinal inertia about the steer axis
-- tall slender fork/handlebar
-- no trail (1 mm postive)
-- half sized wheels (~12 inches)
-- normal wheelbase (1.157)
-- rider tipped forward and low (0.6 m)
-- vertical steer axis
-- 3 m high frame center of mass
+The parameters of the bicycles can be distilled into open loop eigenvalues for
+a clearer understanding of the dynamics, shown in
+Figure~\ref{fig:optimal-eigenvalues}. The 3 and 9 design speed have a simlar
+eigenvalue pattern as well as the 5 and 7.
 
-9
-- very large 5.6 m wheelbase
 %
@@
   \label{fig:optimal-eigenvalues}
-  \caption{}
+  \caption{Real (solid) and imaginary (dashed) parts of the open-loop eigenvlues plotted versus
+    travel speed. The vertical dashed black line indicates the design speed.}
 \end{figure}
@@
 frame
+- need constraints of rear frame that force it to span from rear wheel to
+person to head tube
 \section{Reproducibility}

### 2019-09-01 src/geometry_plot.py
Changed geometry plots to show separate frame and rider.

@@ -7,12 +7,20 @@ speeds = [3, 5, 7, 9]
 
 fig, axes = plt.subplots(nrows=3, ncols=2)
 
+axes[0, 1].axis('off')
+
+
+def plot(par_set, ax):
+    par_set.plot_mass_centers(ax=ax)
+    par_set.plot_principal_radii_of_gyration(ax=ax)
+    par_set.plot_person_diamond(ax=ax)
+    par_set.plot_geometry(ax=ax)
+    return ax
+
 with open('data/principal-browserjason.yml', 'r') as f:
     par_dict = yaml.load(f)['values']
     par_set = PrincipalParameterSet(par_dict)
-    bpar_set = par_set.to_benchmark()
-    bpar_set.plot_all(ax=axes[0, 0])
-    par_set.plot_person_diamond(ax=axes[0, 0])
+    plot(par_set, axes[0, 0])
     axes[0, 0].set_title('Reference')
 
 for ax, speed in zip(axes.flatten()[2:], speeds):
@@ -23,13 +31,10 @@ for ax, speed in zip(axes.flatten()[2:], speeds):
         par_dict = yaml.load(f)['values']
 
     par_set = PrincipalParameterSet(par_dict)
-    bpar_set = par_set.to_benchmark()
-    bpar_set.plot_all(ax=ax)
-    par_set.plot_person_diamond(ax=ax)
+    plot(par_set, ax)
     ax.set_title('{} m/s Design Speed'.format(speed))
-    ax.set_ylabel('')
 
-width = 8.0
+width = 12.0
 fig.set_size_inches((width, 4/3*width))
 fig.tight_layout()
 fig.savefig('figures/optimal-geometries.png', dpi=300)

### 2019-09-01 paper/paper.tex
First draft of discussion and conclusion.

@@
   \centering
-  \includegraphics[width=4in]{../figures/reference-geometry.png}
+  \includegraphics[width=3in]{../figures/reference-geometry.png}
   \caption{Depiction of the principal parameters of the Batavus Browser
@@
 Figure~\ref{fig:optimal-eigenvalues}. The 3 and 9 design speed have a simlar
-eigenvalue pattern as well as the 5 and 7.
-
+eigenvalue pattern as well as the 5 and 7. The 7~\si{\meter\per\second} design
+is self-stable at speeds greater than 1.2~\si{\meter\per\second} but the other
+three designs are not self-stable at any travel speed from
+\SIrange{0}{10}{\meter\per\second}. Both the 5 and 7 designs have a weave mode
+that increases in frequency rapidly with speed. The weave frequency of the 3
+m/s design is higher than the reference design and present at all speeds.
 %
@@
 %
-- need constraints to keep rider out of wheels
-- need more realistic constraints for the frame mass and inertia based on space
-frame
-- need constraints of rear frame that force it to span from rear wheel to
-person to head tube
+We have demonstrated the ability to find optimal parameter values of the linear
+Whipple-Carvallo bicycle model under constraints that enforce a physically
+realizable bicycle with an objective of improving lateral handling qualities.
+We showcase four optimal bicycle designs for a range of target travel speeds.
+The resulting bicycles are similar to the familiar and popular bicycle design
+but have some oddities: very large size relative to the person, large negative
+trail, large positive trail, minimal steer inertia, recumbent rider
+orientation, and both very low and high rider mass center locations. Three of
+the four bicycles are not open-loop self stable for any speed up to
+10~\si{\meter\per\second} and the weave model frequencies are significantly
+larger than the reference bicycle but all bicycles are controllable and have a
+minimal HQM as defined in~\cite{Hess2012}.
+
+The bicycles are reasonably physically realizable but some additional
+constraints could be introduced to further improve this. In the
+5~\si{\meter\per\second} design the person's body overlaps the wheels, which
+would be almost impossible to realize. A constraint that ensures the torso
+cannot occupy the same space as the wheels would solve this. The masses of the
+front and rear frames are low with respect to the geometrical spread a steel
+space frame would have to occupy. The frames also need to be able to span the
+space between the respecitve wheel, mass center of the frame, and the steer
+axis and for the rear frame span to the rider's support locations. We currently
+only expliclty deal with spanning the space between the mass center of the
+frame and the respective wheel. Another more restricting appraoch would be to
+specify a generic geometric structure for the front and rear frames. If done,
+maximal stress and deflections of the frames under load could also be
+minimized.
+
+Each optimizal solution requires approximately 12 hours of computation time on
+a high end consumer desktop machine running on a single thread. The number of
+iterations are typically in the hundreds of thousands for convergence. There
+are many avenues for speeding this up that could likely result in solutions in
+one or two hours on the same machine but this is still slow to be particualrly
+pleasant to utlizie in the process of designing a bicycle. Furthermore, the
+time required to translate the resulting parameters in to an actual bicycle
+design complete with structural destails is extremely time consuming.
+Never-the-less, the method shows promise for optimizing an entire vehicle for
+optimal dynamics. This method can be applied to a whole host of human operated
+vehicles opening up many new designs, but as with any optimization it only
+captures a very small set of the variables that a designer has to take into
+account for a vehicle.
+
 \section{Reproducibility}

### 2019-09-01 src/eigenvalue_plots.py
Slight improvements to main results figures.

@@ -10,6 +10,8 @@ speed_range = np.linspace(0, 10, num=500)
 
 fig, axes = plt.subplots(nrows=3, ncols=2)
 
+axes[0, 1].axis('off')
+
 with open('data/principal-browserjason.yml', 'r') as f:
     par_dict = yaml.load(f)['values']
     par_set = PrincipalParameterSet(par_dict)
@@ -18,6 +20,7 @@ with open('data/principal-browserjason.yml', 'r') as f:
     model.plot_eigenvalue_parts(v=speed_range, ax=axes[0, 0])
     axes[0, 0].set_ylim((-10, 10))
     axes[0, 0].set_title('Reference')
+    axes[0, 0].set_ylabel('')
 
 for ax, speed in zip(axes.flatten()[2:], speeds):
 
@@ -31,11 +34,13 @@ for ax, speed in zip(axes.flatten()[2:], speeds):
     model = MinimalLinearWhippleCarvalloModel(ben_par_set)
     model.plot_eigenvalue_parts(v=speed_range, ax=ax)
     ax.axvline(speed, color='black', linestyle='--')
-    ax.set_title('{} m/s'.format(speed))
+    ax.set_title('{} m/s Design Speed'.format(speed))
     ax.set_ylim((-10, 10))
     ax.set_ylabel('')
 
-width = 8.0
+axes[1, 0].set_ylabel('Real and Imaginary Parts of the Eigenvalues')
+
+width = 6.0
 fig.set_size_inches((width, 4/3*width))
 fig.tight_layout()
 fig.savefig('figures/optimal-eigenvalues.png', dpi=300)

### 2019-09-01 src/geometry_plot.py

@@ -34,7 +34,7 @@ for ax, speed in zip(axes.flatten()[2:], speeds):
     plot(par_set, ax)
     ax.set_title('{} m/s Design Speed'.format(speed))
 
-width = 12.0
+width = 6.0
 fig.set_size_inches((width, 4/3*width))
 fig.tight_layout()
 fig.savefig('figures/optimal-geometries.png', dpi=300)

### 2019-09-01 paper/paper.tex
Switch to newer bmd cls file.
Updated to correct bmd2019 format.

@@
-\documentclass{bmd2019a}
+\documentclass{bmd2019p}
 
@@
 
-\begin{flushleft}
-{\fontsize{16pt}{20pt}\selectfont%
-  Expanded Optimization for Discovering Optimal Lateral Handling Bicycles}
-\end{flushleft}
+\begin{center}
+\fontsize{14}{20}{\bf Expanded Optimization for Discovering \\[3pt]
+                      Optimal Lateral Handling Bicycles}
+\end{center}
 
 %%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
-\begin{flushleft}
-  {\fontsize{12}{14}{Jason K. Moore and Mont Hubbard}\\}
-  \textit{Mechanical and Aerospace Engineering\\
+\begin{center}
+  \normalsize{\bf{Jason K. Moore$^{*}$, Mont Hubbard$^\dag$}}
+\end{center}
+
+\begin{center}
+  \begin{tabular}{c}
+    Department of Mechanical and Aerospace Engineering\\
     University of California, Davis\\
     One Shields Avenue, Davis, CA, USA 95817\\
-          e-mail: jkm@ucdavis.edu, mhubbard@ucdavis.edu}
-\end{flushleft}
+    e-mail: jkm@ucdavis.edu$^{*}$, mhubbard@ucdavis.edu$^\dag$
+  \end{tabular}
+\end{center}
 
-% - Abstract
-% - Introduction
-% - Parameterization
-% - Bounds and Constraints
-% - Free Parameters
-% - Optimization
-% - Solutions
-%   - Parameters
-%   - Eigenvalues
-%   - HQM
-% - Discussion and Conclusion
-
-\section*{Abstract}
+\section*{ABSTRACT}
 %
@@
 bicycles are not generally self-stable and have some ununusal characteristics.
+%
+\begin{keywords}
+  bicycle,
+  dynamics,
+  handling qualities,
+  control,
+  optimization,
+  design
+\end{keywords}
 
-\section{Introduction}
+\section{INTRODUCTION}
 %
@@
 
-\section{Bicycle Model Parameterization}
+\section{BICYCLE MODEL PARAMETERIZATION}
 %
@@
 
-\section{Bounds and Constraints}
+\section*{BOUNDS AND CONSTRAINTS}
 %
@@
 
-\section{Optimization}
+\section{OPTIMIZATION}
 %
@@
 
-\section{Results}
+\section{RESULTS}
 %
@@
 
-\section{Discussion and Conclusion}
+\section{DISCUSSION AND CONCLUSION}
 %
@@
 
-\section{Reproducibility}
+\section{REPRODUCIBILITY}
 %

### 2019-09-01 paper/paper.tex
Editing.

@@
 Previously, we introduced a method of optimizing four primary geometric
-parameters of a bicycle's design to minimize its Handling Quality Metric. Here
-we expand that method to optimize 23 geometric and inertial parameters in the
-benchmark parameterimization of the linear Whipple-Carvallo bicycle model. To
-ensure physically realizable bicycle designs we include 7 equality constraints,
-21 inequality constraints, and maximal and minamal bounds on each free
-opmization parameter. This improves over the prior work by expanding the search
-space with many more parameters and the guarntee of realizability. We close by
-presenting five bicycle designs that have optimal later handling qualities. The
-bicycles are not generally self-stable and have some ununusal characteristics.
+parameters of a bicycle's design to maximize its lateral handling qualities.
+Here we expand that method to optimize over all of the geometric and inertial
+parameters in the linear Whipple-Carvallo bicycle model. To ensure physically
+realizable bicycle designs we include 7 equality constraints, 21 inequality
+constraints, and maximal and minimal bounds on each free optimization
+parameter. This improves over the prior work by expanding the search space with
+many more parameters and the guarantee of physical realizability. We present
+four bicycle designs discovered by the optimization procedure that have optimal
+later handling qualities. The bicycles are similar in design to familiar
+bicycle designs but are not generally self-stable and exhibit unusual
+characteristics such as large positive and negative trail, large size relative
+to the rider, and minimal steering inertia. The method is a useful tool for
+generating atypical bicycle designs that exhibit desired dynamical qualities
+and could be broadly applied to other vehicle designs.
 %
@@
 do so, we formulate a constrained optimization problem and use derivative-free
-optimization to discover optimal, yet realizable, bicycle designs.
+optimization to discover optimal, yet realizable, bicycle designs. We close by
+discussing the characteristics of the discovered designs.
 
@@
 Our problem formulation relies on a new bicycle model parameterization that
-reflects both a reformulation of and addition to the benchmark parameterization
-of the linear Whipple-Carvallo bicycle model~\cite{Meijaard2007}. We call this
-the ``principal parameterization'' as opposed to the ``benchmark
-parameterization'' in~\cite{Meijaard2007}. This parameterization differs from
-the benchmark parameterization in three ways. Firstly, the person and rear
-frame are treated as separate rigid bodies each with their on inertial
-parameters. Secondly, we express the inertial parameters of each rigid body in
-terms of central principal radii of gyration to decouple the mass from the
-inertia terms.  Lastly, we introduce two simple dimensional parameters that
-define the geometric extents of the person which are used to constrain the
-location of the person's body. Table~\ref{tab:parameters} provides the
-parameters and the reference values which are derived from the measurements of
-a Batavus Browser Bicycle and the rider ``Jason'' presented
-in~\cite{Moore2012}. This parameterization can be transformed into the
-benchmark parameterization readily, but not vice versa.
+reflects both a reformulation of and extension to the benchmark
+parameterization of the linear Whipple-Carvallo bicycle
+model~\cite{Meijaard2007}. We call this the ``principal parameterization'' as
+opposed to the ``benchmark parameterization'' in~\cite{Meijaard2007}. This
+parameterization differs from the benchmark parameterization in three ways.
+Firstly, the person and rear frame are treated as separate rigid bodies each
+with their own inertial parameters. Secondly, we express the inertial
+parameters of each rigid body in terms of central principal radii of gyration
+to decouple the mass from the inertia terms. Lastly, we introduce two simple
+dimensional parameters that define the geometric extents of the person which
+are used to constrain the location of the person's body in the optimization
+problem. Table~\ref{tab:parameters} provides the parameters and the reference
+values which are derived from the measurements of a Batavus Browser Bicycle and
+the rider ``Jason'' presented in~\cite{Moore2012}. This represents a typical
+bicycle and rider. The principal parameterization can be transformed into the
+benchmark parameterization readily, but not vice versa. Note that the benchmark
+parameter variables names are also used in the paper for convenience and are
+defined in~\cite{Meijaard2007}.
 %
@@
     geometry. The dotted black line represents the steer axis. The inertial
-    properties of the five rigid bodies: front wheel (orange), rear wheel
+    properties of the five rigid bodies, front wheel (orange), rear wheel
     (purple), rear frame (blue), front frame (green), and person (red) are
     shown with the mass center and the extents of the centroidal principal
-    radii of gyration of each rigid body. The primary principal angles,
-    $\alpha_{D,P,H}$, are defined as the angle about the $y$ axis from $x$ to
-    the maximum principal axes in the $XZ$ plane.}
+    radii of gyration of each rigid body as colored solid lines. The primary
+    principal angles, $\alpha_{D,P,H}$, are defined as the angle about the $y$
+    axis from $x$ to the maximum principal axes in the $XZ$ plane.}
 \end{figure}
@@
     from the measurements in~\cite{Moore2012} of the Batavus Browser bicycle
-    and rider ``Jason''. Note that the benchmark parameters are also used in
-    the paper and are defined in~\cite{Meijaard2007}.}
+    and rider ``Jason''. The $A$ and $B$ axes are the maximal and minimal
+    principal axes in the $XZ$ plane, respectively.}
   \label{tab:parameters}
@@
     % 5
-    $c$       &  0.068581           & \si{\meter} & Trail \\
-    $w$       &  1.1210             & \si{\meter} & Wheelbase \\
+    $c$       &  0.069           & \si{\meter} & Trail \\
+    $w$       &  1.121             & \si{\meter} & Wheelbase \\
     $\lambda$ & 0.39968             & \si{\radian} & Steer axis tilt \\
@@
     $m_R$     & 3.11 & \si{\kilogram}  & Mass \\
-    $r_R$     & 0.34096 & \si{\meter}  & Radius \\
-    $x_R$     & 0.0 & \si{\meter}      & $X$ mass center coordinate \\
-    $y_R$     & 0.0 & \si{\meter}      & $Y$ mass center coordinate \\
-    $z_R$     & -0.34096 & \si{\meter} & $Z$ mass center coordinate \\
-    $k_{Raa}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
-    $k_{Rbb}$ & 0.17050 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
-    $k_{Ryy}$ & 0.22136 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
+    $r_R$     & 0.341 & \si{\meter}  & Radius \\
+    $x_R$     & 0.000 & \si{\meter}      & $X$ mass center coordinate \\
+    $y_R$     & 0.000 & \si{\meter}      & $Y$ mass center coordinate \\
+    $z_R$     & -0.341 & \si{\meter} & $Z$ mass center coordinate \\
+    $k_{Raa}$ & 0.171 & \si{\meter}  & Central principal radii of gyration about $A_R$ \\
+    $k_{Rbb}$ & 0.171 & \si{\meter}  & Central principal radii of gyration about $B_R$ \\
+    $k_{Ryy}$ & 0.221 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
     Front Wheel [F] & & \\
@@
     $m_F$     & 2.02 & \si{\kilogram} & Mass \\
-    $r_F$     & 0.34353 & \si{\meter} & Radius \\
-    $x_F$     & 1.1210 & \si{\meter} & $X$ mass center coordinate \\
-    $y_F$     & 0.0 & \si{\meter} & $Y$ mass center coordinate \\
-    $z_F$     & -0.34353 & \si{\meter} & $Z$ mass center coordinate \\
-    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
-    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
-    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radii of gyration about $Y$ \\
+    $r_F$     & 0.344 & \si{\meter} & Radius \\
+    $x_F$     & 1.121 & \si{\meter} & $X$ mass center coordinate \\
+    $y_F$     & 0.000 & \si{\meter} & $Y$ mass center coordinate \\
+    $z_F$     & -0.344 & \si{\meter} & $Z$ mass center coordinate \\
+    $k_{Faa}$ &  0.209 & \si{\meter} & Central principal radii of gyration about $A_F$ \\
+    $k_{Fbb}$ &  0.209 & \si{\meter} & Central principal radii of gyration about $B_F$ \\
+    $k_{Fyy}$ &  0.272 & \si{\meter} & Central principal radii of gyration about $Y$ \\
     Person [P] & & \\
@@
     $w_P$      &  0.483   & \si{\meter} & Body width \\
-    $m_P$      &  83.5    & \si{\kilogram} & Mass \\
-    $x_P$      &  0.31577 & \si{\meter} & $X$ mass center coordinate \\
-    $y_P$      & 0.0      & \si{\meter} & $Y$ mass center coordinate \\
-    $z_P$      & -1.0990  & \si{\meter} & $Z$ mass center coordinate \\
-    $k_{Paa}$  &  0.36797 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
-    $k_{Pbb}$  &  0.15276 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
-    $k_{Pyy}$  &  0.36717 & \si{\meter} & Central principal radii of gyration about $Y$ \\
-    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $A_P$ and $X$ \\
+    $m_P$      &  83.50    & \si{\kilogram} & Mass \\
+    $x_P$      &  0.316 & \si{\meter} & $X$ mass center coordinate \\
+    $y_P$      & 0.000      & \si{\meter} & $Y$ mass center coordinate \\
+    $z_P$      & -1.099  & \si{\meter} & $Z$ mass center coordinate \\
+    $k_{Paa}$  &  0.368 & \si{\meter} & Central principal radii of gyration about $A_P$ \\
+    $k_{Pbb}$  &  0.153 & \si{\meter} & Central principal radii of gyration about $B_P$ \\
+    $k_{Pyy}$  &  0.367 & \si{\meter} & Central principal radii of gyration about $Y$ \\
+    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $X$ and $A_P$ \\
     Front Frame [H] & & \\
@@
     % 8
-    $m_H$ & 3.2200 & \si{\kilogram}     & Mass \\
-    $x_H$ & 0.86695 & \si{\meter}       & $X$ mass center coordinate \\
-    $y_H$ & 0.0 & \si{\meter}           & $Y$ mass center coordinate \\
-    $z_H$ & -0.74824 & \si{\meter}      & $Z$ mass center coordinate \\
-    $k_{Haa}$ & 0.29556 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
-    $k_{Hbb}$ & 0.14493 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
-    $k_{Hyy}$ & 0.27630 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
-    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $A_H$ and $X$ \\
+    $m_H$ & 3.22 & \si{\kilogram}     & Mass \\
+    $x_H$ & 0.867 & \si{\meter}       & $X$ mass center coordinate \\
+    $y_H$ & 0.000 & \si{\meter}           & $Y$ mass center coordinate \\
+    $z_H$ & -0.748 & \si{\meter}      & $Z$ mass center coordinate \\
+    $k_{Haa}$ & 0.296 & \si{\meter}   & Central principal radii of gyration about $A_H$ \\
+    $k_{Hbb}$ & 0.145 & \si{\meter}   & Central principal radii of gyration about $B_H$ \\
+    $k_{Hyy}$ & 0.276 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
+    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $X$ and $A_H$ \\
     Rear Frame [D] & & \\
@@
     % 8
-    $m_D$ &  9.8600 & \si{\kilogram}    & Mass \\
-    $x_D$ &  0.27595 & \si{\meter}      & $X$ mass center coordinate \\
-    $y_D$ & 0 & \si{\meter}             & $Y$ mass center coordinate \\
-    $z_D$ & -0.53784 & \si{\meter}      & $Z$ mass center coordinate \\
-    $k_{Daa}$ &  0.28587 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
-    $k_{Dbb}$ &  0.22079 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
-    $k_{Dyy}$ &  0.36539 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
-    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $A_D$ and $X$ \\
+    $m_D$ &  9.86 & \si{\kilogram}    & Mass \\
+    $x_D$ &  0.276 & \si{\meter}      & $X$ mass center coordinate \\
+    $y_D$ & 0.000 & \si{\meter}             & $Y$ mass center coordinate \\
+    $z_D$ & -0.538 & \si{\meter}      & $Z$ mass center coordinate \\
+    $k_{Daa}$ &  0.286 & \si{\meter}  & Central principal radii of gyration about $A_D$ \\
+    $k_{Dbb}$ &  0.221 & \si{\meter}  & Central principal radii of gyration about $B_D$ \\
+    $k_{Dyy}$ &  0.365 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
+    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $X$ and $A_D$ \\
     \bottomrule
@@
 
-\section*{BOUNDS AND CONSTRAINTS}
+\section{BOUNDS AND CONSTRAINTS}
 %
@@
 
-  \item[Rider $P$] A single rigid body represents the rider.
+  \item[Person $P$] A single rigid body represents the rider.
     \begin{itemize}
-      \item Rider mass is that of an average person.
+      \item Rider mass is that of a typical person.
       \item The rider's joint angles are fixed in a nominal configuration
@@
     \toprule
-    Min & & Parameter & & Max\\
+    Lower & & Parameter & & Upper \\
     \midrule
@@
     $-\pi/2 $ & $\leq$ & $\alpha_P$  &  $\leq$  &  $\pi/2$ \\
-    $0.25   $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\
+    $0.25\si{\kilogram} $ & $\leq$ & $m_H     $  &  $\leq$  &  $\infty$ \\
     $-\infty$ & $\leq$ & $x_H     $  &  $\leq$  &  $\infty$ \\
@@
 To prevent the rider from being positioned and oriented such that their body is
-not penetrating the ground we introduce two dimensions that define a cross
-whose apex is at the center of mass of the person and the cross axes are
-parallel to the principal axes in the $XZ$ plane. $l_P / 2$ is the distance
-along the principal axis to the tip of the toes and $w_P / 2$ is the distance
-along the second principal axes to the tip of the hands. The constraints
-$c_2,\ldots,c_5$ are derived from these rules.
+penetrating the ground we introduce two dimensions that define a cross whose
+apex is at the center of mass of the person and the cross axes are parallel to
+the principal axes in the $XZ$ plane. $l_P / 2$ is the distance along the
+principal axis to the tip of the toes and $w_P / 2$ is the distance along the
+second principal axes to the tip of the hands. The constraints $c_2,\ldots,c_5$
+are derived from these rules.
 

### 2019-09-01 paper/paper.tex
Edited down to results.

@@
 The front frame is symmetric about the $XZ$ plane so $I_{Hxy}, I_{Hyz} = 0$. We
-allow for any angular orientation of the $XZ$ principal directions but limit
-the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$. We prevent the
-rear frame from penetrating the ground by limiting the inertial spread with
-respect to its mass center, $c_{13}$, but also set a minimum inertial spread to
-ensure a frame can span from the rear wheel to the mass center of the rear
-frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based on the ratio of
-geometrical spread of a typical bicycle frame and its radius of gyration. The
-front frame is not planar but is narrow with respect to the $XZ$ plane, which
-is enforced by constraint $c_1$.
+allow for any angular orientation of the principal directions in the $XZ$ plane
+but limit the angle to $-\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2}$. We
+prevent the rear frame from penetrating the ground by limiting the inertial
+spread with respect to its mass center, $c_{13}$, but also set a minimum
+inertial spread to ensure a frame can span from the rear wheel to the mass
+center of the rear frame, $c_8$. The spread factor in $c_{13}$ of 1.4 is based
+on the ratio of geometrical spread of a typical bicycle frame and its radius of
+gyration. The front frame is not planar due to the need for handlebars
+protruding in the $Y$ direction so we only ensure consistent moments of inertia
+with $c_1$.
 
@@
 geometrical spread of a typical bicycle frame and its radius of gyration.
-Several parameters are bounded. We require the rear frame mass to be positive,
-the center of mass not penetrate the ground, and we allow for any angular
-orientation of the $XZ$ principal directions but limit the angle to
-$-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$ as angle beyond that are
-redundant.
+Several rear frame parameters are bounded. We require the rear frame mass to be
+positive, the center of mass not penetrate the ground, and we allow for any
+angular orientation of the principal directions in the $XZ$ plane but limit the
+angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$ as angle beyond that
+are redundant.
 
@@
 We enforce the assumption that both wheels have moments of inertia of that of a
-simple ring, $g_2 \ldots g_7$ and that mass and radius should be greater than a
-minimal size based on small purchasable spoked wheel with tire.
+simple ring, $g_2 \ldots g_7$ and that the mass and radius should be greater
+than a minimal size based on small purchasable spoked wheel with tire.
 
@@
 eight closed loop eigenvalues associated with the controller in~\cite{Hess2012}
-to be stable, i.e. have negative real parts. These are expressed in
-constraints $c_{14},\ldots,c_{21}$. Closed loop stability is required for the
-HQM to provide a meaningful result.
+to be stable, i.e. have negative real parts. These are expressed in constraints
+$c_{14},\ldots,c_{21}$. Closed-loop stability is required for the HQM to
+provide a meaningful result, see~\cite{Hess2012} for an explanation.
 
@@
 parameter values we generate a bandwidth limited human-like controller using
-the methods in ~\cite{Moore2012}. Once the closed loop stable controller is
-constructed, the HQM can be computed as per the definition in \cite{Hess2012}
+the methods in~\cite{Moore2012}. Once the closed-loop stable controller is
+constructed, the HQM can be computed as per the definition in~\cite{Hess2012}
 and the scaler peak value returned as the objective $J$. This problem is

### 2019-09-01 paper/paper.tex
Finished my first pass at editing.

@@
     $w$       &  1.121             & \si{\meter} & Wheelbase \\
-    $\lambda$ & 0.39968             & \si{\radian} & Steer axis tilt \\
+    $\lambda$ & 0.400             & \si{\radian} & Steer axis tilt \\
     $g$       &  9.81               & \si{\meter\per\second\squared} & Acceleration due to gravity \\
@@
     $k_{Pyy}$  &  0.367 & \si{\meter} & Central principal radii of gyration about $Y$ \\
-    $\alpha_P$ & 0.18618 & \si{\radian} & Angle about $Y$ between $X$ and $A_P$ \\
+    $\alpha_P$ & 0.186 & \si{\radian} & Angle about $Y$ between $X$ and $A_P$ \\
     Front Frame [H] & & \\
@@
     $k_{Hyy}$ & 0.276 & \si{\meter}   & Central principal radii of gyration about $Y$ \\
-    $\alpha_H$ & 0.36995 & \si{\radian} & Angle about $Y$ between $X$ and $A_H$ \\
+    $\alpha_H$ & 0.370 & \si{\radian} & Angle about $Y$ between $X$ and $A_H$ \\
     Rear Frame [D] & & \\
@@
     $k_{Dyy}$ &  0.365 & \si{\meter}  & Central principal radii of gyration about $Y$ \\
-    $\alpha_D$ &  1.1722 & \si{\radian} & Angle about $Y$ between $X$ and $A_D$ \\
+    $\alpha_D$ &  1.172 & \si{\radian} & Angle about $Y$ between $X$ and $A_D$ \\
     \bottomrule
@@
 optimal parameter values for these four bicycles are presented in
-Table~\ref{tab:optimal-values}. We belive these bicycles to be reasoably
+Table~\ref{tab:optimal-values}. We believe these bicycles to be reasonably
 physically realizable. The pictorial representation of the bicycles are
@@
   \caption{Optimal principal parameter values for each design speed. Values
-    with an asterisk are at a bound.}
+    with an asterisk are at a parameter bound.}
   \label{tab:optimal-values}
@@
   \caption{Depictions of the bicycle geometry and geometric representations of
-    the inertial quantities for the reference bicycle and four optimal solutions
-    at 3, 5, 7, and 9 m/s. Five rigid bodies are shown for each bicycle: front
-    wheel (orange), rear wheel (purple), rear frame (blue), front frame
-    (green), and person (red). The solid black lines represent the essential
-    bicycle geometry. The dotted black line represents the steer axis. The
-    solid colored curves represent the contours of solid ellipsoids with
-    equivalent inertia as the principal inertia of the associated rigid body.
-    The dotted colored lines represent the extents of the centroidal radii of
-    gyration of each rigid body.}
+    the inertial quantities for the reference bicycle and four optimal
+    solutions at 3, 5, 7, and 9~\si{\meter\per\second}. Five rigid bodies are
+    shown for each bicycle: front wheel (orange), rear wheel (purple), rear
+    frame (blue), front frame (green), and person (red). The solid black lines
+    represent the essential bicycle geometry. The dashed black line represents
+    the steer axis. The dashed colored lines represent the extents of the
+    centroidal radii of gyration of each rigid body and the red diamond shows
+    the physical extents of the persons geometry.}
 \end{figure}
@@
 than the reference bicycle. The wheelbase is about 3~\si{\meter} and it has a
-relatively large postive trail (0.7~\si{\meter}). The person is rearward on the
-bike and over 3~\si{\meter} above the ground. The mass of the front frame is
-the minimum posible value and the mass of the rear frame is also small.
+relatively large positive trail (0.7~\si{\meter}). The person is rearward on
+the bike and over 3~\si{\meter} above the ground. The mass of the front frame
+is the minimum possible value and the mass of the rear frame is also small.
 
@@
 like one might see on a recumbent bicycle. This significantly increases the
-vehicles yaw moment of inertia. The trail is miminal, but suprisingly slight
+vehicles yaw moment of inertia. The trail is minimal, but surprisingly slightly
 negative (5~\si{\milli\meter}). The steer axis tilt is slightly shallower than
@@
 
-The bicycle optimized for 5~\si{\meter\per\second} is one of the three that is
+The bicycle optimized for 7~\si{\meter\per\second} is one of the three that is
 geometrically smaller than the reference bicycle. The wheelbase is slightly
 larger but the wheels are approximately half the diameter. The rider is low to
-the ground and tipped forward. There is effiectively no trail, the steer axis
-is vertical, and the (very) minimal principal axis is aligned with the steer
-axis making the steering inertia-less. The roll moment of inertia is larger.
+the ground and tipped forward. There is effectively no trail, the steer axis is
+vertical, and the (very) minimal principal axis is aligned with the steer axis
+making the steering effectively inertia-less. The roll moment of inertia is
+larger than the reference bicycle.
 
-The bicycle optimized for 5~\si{\meter\per\second} is much larger than the
+The bicycle optimized for 9~\si{\meter\per\second} is much larger than the
 reference bicycle: wheelbase and rear wheel diameter of 6~\si{\meter}, front
 wheel diameter of 4~\si{\meter}. The trail is large and negative, surely making
-the bicycle open-loop unstable.
+the bicycle open-loop unstable. The person is in a recumbent position and very
+rearward. The mass centers of the person, front and rear frames are very close
+to the respective wheel mass centers. The minor principal axis of the front
+frame is aligned with the steer axis, reducing the steer inertia. The masses of
+the two frames do not seem to be sufficient for such a large steel space frame.
 
-The parameters of the bicycles can be distilled into open loop eigenvalues for
-a clearer understanding of the dynamics, shown in
-Figure~\ref{fig:optimal-eigenvalues}. The 3 and 9 design speed have a simlar
-eigenvalue pattern as well as the 5 and 7. The 7~\si{\meter\per\second} design
-is self-stable at speeds greater than 1.2~\si{\meter\per\second} but the other
+The parameters of the bicycles can also be distilled into open loop eigenvalues
+for a clearer picture of the dynamics, shown in
+Figure~\ref{fig:optimal-eigenvalues}. The 3 and 9~\si{\meter\per\second}
+designs have a similar eigenvalue pattern as does the 5 and
+7~\si{\meter\per\second} designs. The 7~\si{\meter\per\second} design is
+self-stable at speeds greater than 1.2~\si{\meter\per\second} but the other
 three designs are not self-stable at any travel speed from
-\SIrange{0}{10}{\meter\per\second}. Both the 5 and 7 designs have a weave mode
-that increases in frequency rapidly with speed. The weave frequency of the 3
-m/s design is higher than the reference design and present at all speeds.
+\SIrange{0}{10}{\meter\per\second}. Both the 5 and 7~\si{\meter\per\second}
+designs have a weave mode that increases in frequency rapidly with speed. The
+weave frequency of the 3~\si{\meter\per\second} design is higher than the
+reference design and present at all speeds.
 %
@@
   \label{fig:optimal-eigenvalues}
-  \caption{Real (solid) and imaginary (dashed) parts of the open-loop eigenvlues plotted versus
-    travel speed. The vertical dashed black line indicates the design speed.}
+  \caption{Real (solid) and imaginary (dashed) parts of the open-loop
+    eigenvalues plotted versus travel speed. The vertical dashed black line
+    indicates the design speed.}
 \end{figure}
@@
 space frame would have to occupy. The frames also need to be able to span the
-space between the respecitve wheel, mass center of the frame, and the steer
+space between the respective wheel, mass center of the frame, and the steer
 axis and for the rear frame span to the rider's support locations. We currently
-only expliclty deal with spanning the space between the mass center of the
-frame and the respective wheel. Another more restricting appraoch would be to
+only explicitly deal with spanning the space between the mass center of the
+frame and the respective wheel. Another more restricting approach would be to
 specify a generic geometric structure for the front and rear frames. If done,
@@
 
-Each optimizal solution requires approximately 12 hours of computation time on
-a high end consumer desktop machine running on a single thread. The number of
-iterations are typically in the hundreds of thousands for convergence. There
-are many avenues for speeding this up that could likely result in solutions in
-one or two hours on the same machine but this is still slow to be particualrly
-pleasant to utlizie in the process of designing a bicycle. Furthermore, the
-time required to translate the resulting parameters in to an actual bicycle
-design complete with structural destails is extremely time consuming.
-Never-the-less, the method shows promise for optimizing an entire vehicle for
-optimal dynamics. This method can be applied to a whole host of human operated
-vehicles opening up many new designs, but as with any optimization it only
-captures a very small set of the variables that a designer has to take into
-account for a vehicle.
+Each optimized solution requires approximately 12 hours of computation time on
+a high-end consumer desktop computer running the optimization procedure
+sequentially. The number of iterations are typically in the hundreds of
+thousands for convergence. There are many avenues for speeding this up that
+could likely result in solutions in one or two hours on the same computer
+hardware but this is still slow to be particularly convenient to utilize in the
+process of designing a bicycle. Furthermore, the time required to translate the
+resulting parameters into an actual bicycle design complete with structural
+details is extremely time consuming.  Never-the-less, the method shows promise
+for optimizing an entire vehicle for optimal dynamics. This method can be
+applied to a whole host of human operated vehicles opening up many new designs,
+but as with any optimization it only captures a very small set of the variables
+that a designer has to take into account for a vehicle.
 

### 2019-09-01 paper/paper.tex
Change to lower/upper.

@@
 realizable bicycle designs we include 7 equality constraints, 21 inequality
-constraints, and maximal and minimal bounds on each free optimization
+constraints, and lower and upper bounds on each free optimization
 parameter. This improves over the prior work by expanding the search space with

### 2019-09-01 paper/paper.tex
Fixed Mont's comments.

--- 2019-09-01 paper/paper.tex
@@
     shown with the mass center and the extents of the centroidal principal
-    radii of gyration of each rigid body as colored solid lines. The primary
-    principal angles, $\alpha_{D,P,H}$, are defined as the angle about the $y$
-    axis from $x$ to the maximum principal axes in the $XZ$ plane.}
+    radii of gyration of each rigid body as colored solid straight lines and
+    circles. The primary principal angles, $\alpha_{D,P,H}$, are defined as the
+    angle about the $y$ axis from $x$ to the maximum principal axes in the $XZ$
+    plane.}
 \end{figure}
@@
 angular orientation of the principal directions in the $XZ$ plane but limit the
-angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$ as angle beyond that
-are redundant.
+angle to $-\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2}$ as angles beyond
+that are redundant.
 
@@
 constructed, the HQM can be computed as per the definition in~\cite{Hess2012}
-and the scaler peak value returned as the objective $J$. This problem is
-presented as a non-linear programming problem in the following equation.
+and the scaler peak value returned as the objective function $J$. This problem
+is presented as a non-linear programming problem in the following equation.
 %
@@
 algorithm is able to move into the parameter space where all the constraints
-are satisfied and find a (local) minima. For our purposes, this sufficiently
+are satisfied and find a local minima. For our purposes, this sufficiently
 finds parameter values that produce optimally handling bicycle designs.
@@
     represent the essential bicycle geometry. The dashed black line represents
-    the steer axis. The dashed colored lines represent the extents of the
-    centroidal radii of gyration of each rigid body and the red diamond shows
-    the physical extents of the persons geometry.}
+    the steer axis. The dashed colored straight lines and circles represent the
+    extents of the centroidal radii of gyration of each rigid body and the red
+    diamond shows the physical extents of the person's length $l_p$ and width
+    $w_p$.}
 \end{figure}
@@
 like one might see on a recumbent bicycle. This significantly increases the
-vehicles yaw moment of inertia. The trail is minimal, but surprisingly slightly
+vehicle's yaw moment of inertia. The trail is minimal, but surprisingly slightly
 negative (5~\si{\milli\meter}). The steer axis tilt is slightly shallower than
@@
 resulting parameters into an actual bicycle design complete with structural
-details is extremely time consuming.  Never-the-less, the method shows promise
-for optimizing an entire vehicle for optimal dynamics. This method can be
-applied to a whole host of human operated vehicles opening up many new designs,
-but as with any optimization it only captures a very small set of the variables
-that a designer has to take into account for a vehicle.
+details is extremely time consuming. Nevertheless, the method shows promise for
+optimizing an entire vehicle for optimal dynamics. This method can be applied
+to a whole host of human operated vehicles opening up many new designs, but as
+with any optimization it only captures a very small set of the variables that a
+designer has to take into account for a vehicle.
 
